---
title: "Statistical analyzes : Simon Couillard <br><br>"
subtitle:
  "Project : ACQ5 multicohort <br>"
author: "Reference number : 1206 <br><br> By : Pascal Charland, Samuel Lemaire-Paquette, and Simon Couillard,"
date: " Last update : `r Sys.Date()`<br><br>"
output: 
  html_document:
    theme : united
    df_print: paged
    toc: true
    toc_depth: 1
    number_sections: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
# Define variable containing url
url <- "https://www.crchus.ca/clients/SanteEstrie/Sous-sites/Centres_de_recherche/CRCHUS/Medias-publications/signature_CR-CHUS_rvb-s2.jpg"
```

```{=html}
<script>
   $(document).ready(function() {
     $head = $('#header');
     $head.prepend('<img src="`r url`" style=\"float: right;width: 200px;\"/>')
   });
</script>
```

```{=html}
<style>
body {
text-align: justify}
</style>
```

```{=html}
<style type="text/css">
  body{
  font-size: 12pt;
}
</style>
```

```{=html}
<style type="text/css">
.main-container {
  max-width: 1600px;
  margin-left: auto;
  margin-right: auto;
}
</style>
```

```{r Packages setup, include = FALSE}

options(scipen=999)

load_packages <- function(packages) {
  # Check which packages are not installed
  not_installed <- packages[!(packages %in% installed.packages()[,"Package"])]
  
  # Install missing packages if any
  if(length(not_installed) > 0) {
    install.packages(not_installed, dependencies = TRUE)
  }
  
  # Suppress package loading messages
  suppressPackageStartupMessages({
    invisible(lapply(packages, library, character.only = TRUE))
  })
}

# List of packages
Packages <- c(
  "readxl","haven","ggplot2","dplyr","MASS","summarytools","car","lme4","lmerTest","nlme","multcomp","emmeans","ggeffects","ggiraphExtra","sjPlot","sjmisc","qwraps2","extrafont","openxlsx","igraph","gimme","purrr","ggthemes","plotly","knitr","gmodels","tidyverse","infer","skimr","pander", "RColorBrewer","expss","ggrepel","fmsb","lubridate","mgcv","nlme","glmmTMB","lmerTest","mice","table1","splines", "forplo", "cowplot","forestplot",'gridExtra','grid','metafor'
)

# Load packages
load_packages(Packages)

```

```{r p-value function, include = FALSE}

# P-value function for tables

pvalue <- function(x, ...) {
    y <- unlist(x)
    g <- factor(rep(1:length(x), times=sapply(x, length)))
    if (is.numeric(y)) {
        p <- wilcox.test(y ~ g)$p.value
    } else {
        p <- fisher.test(table(y, g))$p.value
    }
    c("", sub("<", "&lt;", format.pval(p, digits=3, eps=0.001)))
}

pvalue2 <- function(x, ...) {
  y <- unlist(x)
  g <- factor(rep(1:length(x), times=sapply(x, length)))
  
  if (is.numeric(y)) {
    p <- kruskal.test(y ~ g)$p.value
  } else {
    p <- fisher.test(table(y, g), simulate.p.value = TRUE, B = 20000)$p.value
  }
  
  c("", sub("<", "&lt;", format.pval(p, digits=3, eps=0.001)))
}


pvalue3 <- function(x, ...) {
    y <- unlist(x)
    g <- factor(rep(1:length(x), times=sapply(x, length)))
    if (is.numeric(y)) {
        p <- t.test(y ~ g)$p.value
    } else {
        p <- fisher.test(table(y, g))$p.value
    }
    c("", sub("<", "&lt;", format.pval(p, digits=3, eps=0.001)))
}

pvalue4 <- function(x, ...) {
  y <- unlist(x)
  g <- factor(rep(1:length(x), times = sapply(x, length)))
  
  if (is.numeric(y)) {
    # Use one-way ANOVA instead of Kruskal-Wallis
    p <- summary(aov(y ~ g))[[1]][["Pr(>F)"]][1]
  } else {
    p <- fisher.test(table(y, g), simulate.p.value = TRUE, B = 20000)$p.value
  }
  
  c("", sub("<", "&lt;", format.pval(p, digits = 3, eps = 0.001)))
}

```

```{r Data PRISMA, include = FALSE}

# PRISMA import

Data.Prisma <- read_excel("2025_03_24_Master.complete_WithSuppression.xlsx") %>% 
  filter(
    redcap_event_name == "visite_1_arm_1",
    HEALTHY_OR_ASTHMA == 1,
    EXCLUDE == 0
  ) %>% 
  mutate(
    Age_at_study_enrollment_rescale10 = Age_at_study_enrollment*0.1,
    BMI_rescale5 = BMI/5,
    Pre_FEV1_PORC_rescale = Pre_FEV1_PORC*-0.1,
    Pre_FEV1_FVC_ratio_rescale = Pre_FEV1_FVC_ratio*-0.1,
    Severe_asthma_attacks_past_12m_cat = ifelse(is.na(Total_exacerbations_12m),0,1),
    BloodEOS_Abso_log10 = log10(BloodEOS_Abso),
    FeNO_log10 = log10(FeNO),
    ACQ5_rescale = acq5_moy * 2,
    SpuEOS_PCT_log10 = log10(SpuEOS_PCT + 0.25),
    PBS_EOTAX_log10 = log10(PBS_EOTAX),
    PBS_IFNg_log10 = log10(PBS_IFNg),
    PBS_IL13_log10 = log10(PBS_IL13),
    PBS_IL4_log10 = log10(PBS_IL4),
    PBS_IL5_log10 = log10(PBS_IL5),
    PBS_IL33_log10 = log10(PBS_IL33),
    PBS_TARC_log10 = log10(PBS_TARC),
    PBS_TNFa_log10 = log10(PBS_TNFa),
    PBS_TSLP_log10 = log10(PBS_TSLP),
    PBS_PGD2_log10 = log10(PBS_PGD2),
    
    Rhinitis_cat = ifelse( Rhinitis == 0,"No","Yes"        ),
    Chronic_sinusitis_cat = ifelse( Chronic_sinusitis == 0,0,1        ),
    Nasal_Polyps_cat = ifelse( Nasal_Polyps == 0,0,1        ),
    
    CRSsNP = ifelse(Chronic_sinusitis_cat == 1 & Nasal_Polyps_cat == 0, "Yes", "No"),
    CRSwNP = ifelse(Chronic_sinusitis_cat == 1 & Nasal_Polyps_cat == 1, "Yes", "No")
  )

val_lab(Data.Prisma$Severe_asthma_attacks_past_12m_cat) <- make_labels("
                                                                       0 No
                                                                       1 Yes   
                                                                       ")

Data.Prisma$Severe_asthma_attacks_past_12m_cat <-  as.factor(Data.Prisma$Severe_asthma_attacks_past_12m_cat )

#Data.Prisma <- within(Data.Prisma, Severe_asthma_attacks_past_12m_cat <- relevel(Severe_asthma_attacks_past_12m_cat, ref = "No"))

val_lab(Data.Prisma$Sex) <- make_labels("
                                        1 Female
                                        2 Male
                                        ")
Data.Prisma$Sex <- as.factor(Data.Prisma$Sex)

Data.Prisma <- within(Data.Prisma, Sex <- relevel(Sex, ref = "Male"))


val_lab(Data.Prisma$Tobacco) <- make_labels("
                                        2 Never
                                        3 Ex
                                        ")
Data.Prisma$Tobacco <- as.factor(Data.Prisma$Tobacco)

Data.Prisma$T2_hothot_med_lowlow_ordinal <- Data.Prisma$T2_hothot_med_lowlow

val_lab(Data.Prisma$T2_hothot_med_lowlow) <- make_labels("
                                                                                 0 T2-Low/Low
                                                                                 1 T2-Mid
                                                                                 2 T2-High/High
                                                                                 ")
Data.Prisma$T2_hothot_med_lowlow <- as.factor(Data.Prisma$T2_hothot_med_lowlow)

Data.Prisma$GINA_Step <- as.factor(Data.Prisma$GINA_Step)

```

```{r Data Oracle, include = FALSE, warning = FALSE}

# ORACLE import

Data.oracle <- read_excel("ORACLE_DATA_not_fleur_V23_2024-07-11_captainadherence and stratos fev1rev multiplied by 100.xlsx") 

Data.oracle[Data.oracle == 9999] <- NA

Data.oracle <- Data.oracle %>% 
  mutate(
    Age_rescale10 = Age / 10,
    BMI_rescale5 = BMI / 5,
    .after = Treatment_arm
  ) %>% 
  mutate(
    BEC_log10 = log10(Blood_Eos_baseline_x10_9_cells_per_L + 0.25),
    FeNO_log10 = log10(FeNO_baseline_ppb + 0.25),
    FEV1_preBD_PCT_Baseline_rescale10 = -FEV1_preBD_PCT_Baseline / 10,
    Pre_FEV1_FVC_ratio = FEV1_preBD_L_Baseline / FVC_preBD_L_Baseline * 100,
    Pre_FEV1_FVC_ratio_rescale10 = -Pre_FEV1_FVC_ratio / 10,
    ACQ5_rescale = ACQ_baseline_score_mean * 2,
    CRSsNP = ifelse(Chronic_Rhinosinusitis_0no_1yes_9999notknown == 1 & Nasal_polyposis_0no_1yes_9999notknown == 0, "Yes", "No"),
    CRSwNP = ifelse(Chronic_Rhinosinusitis_0no_1yes_9999notknown == 1 & Nasal_polyposis_0no_1yes_9999notknown == 1, "Yes", "No"),
  )

Data.oracle$Treatment_step <- as.factor(Data.oracle$Treatment_step)
Data.oracle$CRSsNP <- as.factor(Data.oracle$CRSsNP)
Data.oracle$CRSwNP <- as.factor(Data.oracle$CRSwNP)

val_lab(Data.oracle$Any_severe_attack_previous_12m_0no_1yes) <- make_labels("
                                                                            0 No
                                                                            1 Yes
                                                                            ")
Data.oracle$Any_severe_attack_previous_12m_0no_1yes <- as.factor(Data.oracle$Any_severe_attack_previous_12m_0no_1yes)

val_lab(Data.oracle$Gender_0Female_1Male) <- make_labels("
                                                         0 Female
                                                         1 Male
                                                         ")
Data.oracle$Gender_0Female_1Male <- as.factor(Data.oracle$Gender_0Female_1Male)
Data.oracle$Gender_0Female_1Male <- relevel(Data.oracle$Gender_0Female_1Male, ref = "Male")

val_lab(Data.oracle$Smoking_0never_1ex_2current) <- make_labels("
                                                                0 Never
                                                                1 Ex
                                                                2 Current
                                                                ")
Data.oracle$Smoking_0never_1ex_2current <- as.factor(Data.oracle$Smoking_0never_1ex_2current)

val_lab(Data.oracle$AllergicRhinitis__0no_1yes_9999notknown) <- make_labels("
                                                                            0 No
                                                                            1 Yes
                                                                            ")
Data.oracle$AllergicRhinitis__0no_1yes_9999notknown <- as.factor(Data.oracle$AllergicRhinitis__0no_1yes_9999notknown)

val_lab(Data.oracle$Chronic_Rhinosinusitis_0no_1yes_9999notknown) <- make_labels("
                                                                                 0 No
                                                                                 1 Yes
                                                                                 ")
Data.oracle$Chronic_Rhinosinusitis_0no_1yes_9999notknown <- as.factor(Data.oracle$Chronic_Rhinosinusitis_0no_1yes_9999notknown)

Data.oracle$T2_composite <- ifelse( Data.oracle$FeNO_baseline_ppb < 25 & Data.oracle$Blood_Eos_baseline_x10_9_cells_per_L < 0.15,0, 
                                            ifelse(Data.oracle$FeNO_baseline_ppb >= 35 & Data.oracle$Blood_Eos_baseline_x10_9_cells_per_L >= 0.3,2,1  ))


Data.oracle$T2_composite_ordinal <- Data.oracle$T2_composite


val_lab(Data.oracle$T2_composite) <- make_labels("
                                                                                 0 T2-Low/Low
                                                                                 1 T2-Mid
                                                                                 2 T2-High/High
                                                                                 ")
Data.oracle$T2_composite <- as.factor(Data.oracle$T2_composite)

# Data Dream
Data.oracle.dream <- Data.oracle %>% 
  filter(Enrolled_Trial_name == "DREAM") %>% 
  # Create a column for Placebo or Mepo
  mutate(
    Mepo_placebo = ifelse(Treatment_arm == "Placebo", "Placebo", "Mepo"),
    .after = Treatment_arm
  )

Data.oracle.dream$Mepo_placebo <- as.factor(Data.oracle.dream$Mepo_placebo)
Data.oracle.dream$Mepo_placebo <- relevel(Data.oracle.dream$Mepo_placebo, ref = "Placebo")

```

```{r Data Oracle imputed, include = FALSE, warning = FALSE}

# ORACLE imputed import

Data.oracle.imputed <- read_excel("data_ORACLE_imp_20250102_joined.xlsx")

Data.oracle.imputed[Data.oracle.imputed == "NA"] <- NA

Data.oracle.imputed <- Data.oracle.imputed %>% 
  mutate(
    Age_rescale10 = Age / 10,
    BMI_rescale5 = BMI / 5,
    .after = Treatment_arm
  ) %>% 
  mutate(
    FEV1_preBD_PCT_Baseline_rescale10 = -FEV1_preBD_PCT_Baseline / 10,
    FEV1_FVC_ratio_rescale10 = -FEV1_FVC_ratio / 10,
    ACQ5_rescale = ACQ_baseline_score_mean * 2,
    Severe_asthma_attacks_during_followup_yesno = ifelse(Number_severe_asthma_attacks_during_followup == 0, "No", "Yes"),
    CRSsNP = ifelse(Chronic_Rhinosinusitis_0no_1yes_9999notknown == 1 & Nasal_polyposis_0no_1yes_9999notknown == 0, "Yes", "No"),
    CRSwNP = ifelse(Chronic_Rhinosinusitis_0no_1yes_9999notknown == 1 & Nasal_polyposis_0no_1yes_9999notknown == 1, "Yes", "No")
  )


Data.oracle.imputed$Treatment_step <- as.factor(Data.oracle.imputed$Treatment_step)
Data.oracle.imputed$CRSsNP <- as.factor(Data.oracle.imputed$CRSsNP)
Data.oracle.imputed$CRSwNP <- as.factor(Data.oracle.imputed$CRSwNP)

Data.oracle.imputed$Severe_asthma_attacks_during_followup_yesno <- as.factor(Data.oracle.imputed$Severe_asthma_attacks_during_followup_yesno)

val_lab(Data.oracle.imputed$Gender_0Female_1Male) <- make_labels("
                                                                 0 Female
                                                                 1 Male
                                                                 ")
Data.oracle.imputed$Gender_0Female_1Male <- as.factor(Data.oracle.imputed$Gender_0Female_1Male)
Data.oracle.imputed$Gender_0Female_1Male <- relevel(Data.oracle.imputed$Gender_0Female_1Male, ref = "Male")

Data.oracle.imputed$.imp <- Data.oracle.imputed$.imp -1

Data.imp <- as.mids(Data.oracle.imputed)

# Oracle2 (imputed version) Any ACQ dataset (n = 750)
exclude_trials <- c("AZISAST", "Novel_START", "PRACTICAL", "PACT", "DREAM")
Data.oracle.any.acq <- Data.oracle.imputed %>% filter(Enrolled_Trial_name %in% exclude_trials)

Data.oracle.any.acq$.imp <- Data.oracle.any.acq$.imp -1

Data.imp.anyacq <- as.mids(Data.oracle.any.acq)

Data.oracle.supp <- Data.oracle %>%
  dplyr::semi_join(
    Test1 %>%
      dplyr::distinct(Subject_ID, Enrolled_Trial_name),
    by = c("Subject_ID", "Enrolled_Trial_name")
  ) 

Data.oracle.supp.any.acq <- Data.oracle.supp %>% filter(Enrolled_Trial_name %in% exclude_trials)

```

```{r , include = FALSE}

#RASP-UK import

Data.raw.raspuk <- read_excel("RAW for SPSS v2 INCLUDES RASP-UK.xlsx") %>% 
  filter(Healthy0Asthma1 == 1)

Data.raw.raspuk[Data.raw.raspuk == 99999] <- NA

Data.raw.raspuk <- Data.raw.raspuk %>% 
  mutate(age_rescale10 = AgeD0 / 10,
    BMI_rescale5 = BMI / 5,
    BEC_log10 = log10(Blood_Eos_optimised + 0.25),
    FeNO_log10 = log10(FeNO_optimised + 0.25),
    FEV1_PCTpredicted_rescale10 = - FEV1_PCTpredicted_optimised / 10,
    FEV1_FVC_ratio_rescale10 = - FEV1_FVC_ratio_optimised / 10,
    Previous_asthma_attack_yesno = ifelse(Number_of_Rescue_Steriod_Courses_In_12_Mths == 0, "No", "Yes"),
    Sp_Eos_optimised_log10 = log10(Sp_Eos_optimised + 0.25),
    Eotaxin_3_Optimised_sputm_log10 = log10(`Eotaxin-3_Optimised_sputm`),
    IFNgamma_Optimised_sputm_log10 = log10(IFNgamma_Optimised_sputm),
    IL13_Optimised_sputm_log10 = log10(IL13_Optimised_sputm),
    IL33_Optimised_sputm_log10 = log10(IL33_Optimised_sputm),
    IL4_Optimised_sputm_log10 = log10(IL4_Optimised_sputm),
    IL5_Optimised_sputm_log10 = log10(IL5_Optimised_sputm),
    TARC_Optimised_sputm_log10 = log10(TARC_Optimised_sputm),
    TNFalpha_Optimised_sputm_log10 = log10(TNFalpha_Optimised_sputm),
    TSLP_Optimised_sputm_log10 = log10(TSLP_Optimised_sputm),
    LTE4_Optimised_sputm_log10 = log10(LTE4_Optimised_sputm),
    PGD2_Optimised_sputm_log10 = log10(PGD2_Optimised_sputm))

Data.raw.raspuk$Previous_asthma_attack_yesno <- as.factor(Data.raw.raspuk$Previous_asthma_attack_yesno)

val_lab(Data.raw.raspuk$Sex0M1F) <- make_labels("
                                                0 Male
                                                1 Female
                                                ")
Data.raw.raspuk$Sex0M1F <- as.factor(Data.raw.raspuk$Sex0M1F)

Data.raw.raspuk$Gina_step <- NA

val_lab(Data.raw.raspuk$Smoking_status_0nil_1ex_2active) <- make_labels("
                                                  0 Never
                                                  1 Ex
                                                  2 Current
                                                  ")
Data.raw.raspuk$Smoking_status_0nil_1ex_2active <- as.factor(Data.raw.raspuk$Smoking_status_0nil_1ex_2active)


Data.raw.raspuk$AllergicRhinitis__0no_1yes_9999notknown <- NA

Data.raw.raspuk$CRSsNP <- NA

Data.raw.raspuk$CRSwNP <- NA

```

```{r Data FeNO, echo = FALSE}

Data.feno <- Data.raw.raspuk %>% 
  filter(CohortOxf0Rasp1 == 0)

```

```{r Data RASP-UK, include = FALSE}

Data.raspuk <- Data.raw.raspuk %>% 
  filter(CohortOxf0Rasp1 == 1)

```

```{r Data Boost, include = FALSE}

# BOOST import

Data.boost <- read.csv("Data_Simon_final.csv", header = TRUE, sep = ",", dec = ".") %>% 
  mutate(
    delta = fev1_D7 - fev1_attack,
    Previous_asthma_attack_yesno = ifelse(exac_12m == 0, "No", "Yes"),
    BEC_log10 = log10(blood_eos_attack + 0.25),
    FeNO_log10 = log10(feno_attack),
    Age_rescale10 = Age / 10,
    BMI_rescale5 =  BMI/5,
    Pre_FEV1_PORC_rescale = fev1_pct_attack*-0.1,
    Pre_FEV1_FVC_ratio_rescale = fev1_fvc_attack*-0.1,
    ACQ5_rescale = acq_attack * 2,
    SpuEOS_PCT_log10 = log10(Sputum_eos_attack + 0.25)  )

Data.boost$Previous_asthma_attack_yesno <- as.factor(Data.boost$Previous_asthma_attack_yesno)

val_lab(Data.boost$Sex) <- make_labels("
                                       1 Male
                                       2 Female
                                       ")
Data.boost$Sex <- as.factor(Data.boost$Sex)

val_lab(Data.boost$Biologic) <- make_labels("
                                            1 mepo
                                            2 benra
                                            3 reslizumab
                                            ")
Data.boost$Biologic <- as.factor(Data.boost$Biologic)

val_lab(Data.boost$smoking_status) <- make_labels("
                                                  1 Never
                                                  2 Current
                                                  3 Ex smoker
                                                  ")
Data.boost$smoking_status <- as.factor(Data.boost$smoking_status)

Data.boost$Gina_step <- 5

Data.boost$Gina_step <- as.factor(Data.boost$Gina_step)

```

# Table 1 {.tabset .tabset-fade}

Frequencies and percentages will be presented for nominal variables. Means (standard deviation) and medians [interquartile range] will be presented for continuous variables. For comparisons between two groups, Mann-Whitney U tests will be used for continuous variables and Fisher's exact tests for categorical variables. For comparisons between three or more groups, Kruskal-Wallis tests will be used for continuous variables and Fisher's exact tests for categorical variables.

## Master {.tabset .tabset-fade}

### Non-parametric - Not Imputed

```{r, echo = FALSE}
# ORACLE2 - Full (n=6513)

table1(~ BMI, 
       data = Data.oracle.supp, 
       render.continuous = c( .="Median [Q1 - Q3]"), 
       render.categorical = "FREQ (PCTnoNA%)", 
       topclass = "Rtable1-grid Rtable1-shade Rtable1-times")

Data.oracle2.full.master <- Data.oracle.supp %>% 
  dplyr::select(c("Sequential_number","Gender_0Female_1Male",'Age',"BMI","ACQ_baseline_score_mean","Any_severe_attack_previous_12m_0no_1yes","Treatment_step","FEV1_preBD_PCT_Baseline","Pre_FEV1_FVC_ratio","Blood_Eos_baseline_x10_9_cells_per_L","FeNO_baseline_ppb","Smoking_0never_1ex_2current","AllergicRhinitis__0no_1yes_9999notknown","CRSsNP","CRSwNP")) %>% 
  mutate(Study = "ORACLE2 - Full",
         ACQ_below_1.5 = ifelse(is.na(ACQ_baseline_score_mean), NA,
                                ifelse(ACQ_baseline_score_mean < 1.5, "Yes", "No")))

# ORACLE2 - Any ACQ (n=750)
Data.oracle2.anyacq.master <- Data.oracle.supp.any.acq %>% 
  dplyr::select(c("Sequential_number","Gender_0Female_1Male",'Age',"BMI","ACQ_baseline_score_mean","Any_severe_attack_previous_12m_0no_1yes","Treatment_step","FEV1_preBD_PCT_Baseline","Pre_FEV1_FVC_ratio","Blood_Eos_baseline_x10_9_cells_per_L","FeNO_baseline_ppb","Smoking_0never_1ex_2current","AllergicRhinitis__0no_1yes_9999notknown","CRSsNP","CRSwNP")) %>% 
  mutate(Study = "ORACLE2 - Any ACQ",
         ACQ_below_1.5 = ifelse(is.na(ACQ_baseline_score_mean), NA,
                                ifelse(ACQ_baseline_score_mean < 1.5, "Yes", "No")))

# DREAM - Placebo
Data.oracle.dream.placebo.master <- Data.oracle.dream %>% 
  filter(Mepo_placebo == "Placebo") %>%
  dplyr::select(c("Sequential_number","Gender_0Female_1Male",'Age',"BMI","ACQ_baseline_score_mean","Any_severe_attack_previous_12m_0no_1yes","Treatment_step","FEV1_preBD_PCT_Baseline","Pre_FEV1_FVC_ratio","Blood_Eos_baseline_x10_9_cells_per_L","FeNO_baseline_ppb","Smoking_0never_1ex_2current","AllergicRhinitis__0no_1yes_9999notknown","CRSsNP","CRSwNP")) %>% 
  mutate(Study = "DREAM - Placebo",
         ACQ_below_1.5 = ifelse(is.na(ACQ_baseline_score_mean), NA,
                                ifelse(ACQ_baseline_score_mean < 1.5, "Yes", "No")))

# DREAM - Mepo
Data.oracle.dream.mepo.master <- Data.oracle.dream %>% 
  filter(Mepo_placebo == "Mepo") %>%
  dplyr::select(c("Sequential_number","Gender_0Female_1Male",'Age',"BMI","ACQ_baseline_score_mean","Any_severe_attack_previous_12m_0no_1yes","Treatment_step","FEV1_preBD_PCT_Baseline","Pre_FEV1_FVC_ratio","Blood_Eos_baseline_x10_9_cells_per_L","FeNO_baseline_ppb","Smoking_0never_1ex_2current","AllergicRhinitis__0no_1yes_9999notknown","CRSsNP","CRSwNP")) %>% 
  mutate(Study = "DREAM - Mepo",
         ACQ_below_1.5 = ifelse(is.na(ACQ_baseline_score_mean), NA,
                                ifelse(ACQ_baseline_score_mean < 1.5, "Yes", "No")))

# SA-OIT (FeNO + RASP-UK)
Data.raw.raspuk.master <- Data.raw.raspuk %>% 
  dplyr::select(c("FENO_DiarySeqNumber","Sex0M1F",'AgeD0',"BMI","ACQ_optimised","Previous_asthma_attack_yesno","Gina_step","FEV1_PCTpredicted_optimised","FEV1_FVC_ratio_optimised","Blood_Eos_optimised","FeNO_optimised","Smoking_status_0nil_1ex_2active","AllergicRhinitis__0no_1yes_9999notknown","CRSsNP","CRSwNP")) %>% 
  mutate(Study = "SA-OIT",
         ACQ_below_1.5 = ifelse(is.na(ACQ_optimised), NA,
                                ifelse(ACQ_optimised < 1.5, "Yes", "No"))) %>% 
  rename(Sequential_number = FENO_DiarySeqNumber,
         Gender_0Female_1Male = Sex0M1F,
         Age = AgeD0,
         ACQ_baseline_score_mean = ACQ_optimised,
         Any_severe_attack_previous_12m_0no_1yes = Previous_asthma_attack_yesno,
         Treatment_step = Gina_step,
         FEV1_preBD_PCT_Baseline = FEV1_PCTpredicted_optimised,
         Pre_FEV1_FVC_ratio = FEV1_FVC_ratio_optimised,
         Blood_Eos_baseline_x10_9_cells_per_L = Blood_Eos_optimised,
         FeNO_baseline_ppb = FeNO_optimised,
         Smoking_0never_1ex_2current = Smoking_status_0nil_1ex_2active)

# PRISMA
Data.Prisma.master <- Data.Prisma %>% 
  dplyr::select(c("ID","Sex",'Age_at_study_enrollment',"BMI","acq5_moy","Severe_asthma_attacks_past_12m_cat","GINA_Step","Pre_FEV1_PORC","Pre_FEV1_FVC_ratio","BloodEOS_Abso","FeNO","Tobacco","Rhinitis_cat","CRSsNP","CRSwNP")) %>% 
  mutate(Study = "PRISMA",
         ACQ_below_1.5 = ifelse(is.na(acq5_moy), NA,
                                ifelse(acq5_moy < 1.5, "Yes", "No"))) %>% 
  rename(Sequential_number = ID,
         Gender_0Female_1Male = Sex,
         Age = Age_at_study_enrollment,
         ACQ_baseline_score_mean = acq5_moy,
         Any_severe_attack_previous_12m_0no_1yes = Severe_asthma_attacks_past_12m_cat,
         Treatment_step = GINA_Step,
         FEV1_preBD_PCT_Baseline = Pre_FEV1_PORC,
         Blood_Eos_baseline_x10_9_cells_per_L = BloodEOS_Abso,
         FeNO_baseline_ppb = FeNO,
         Smoking_0never_1ex_2current = Tobacco,
         AllergicRhinitis__0no_1yes_9999notknown = Rhinitis_cat)

# BOOST
Data.boost.master <- Data.boost %>% 
  dplyr::select(c("study_id","Sex",'Age',"BMI","acq_attack","Previous_asthma_attack_yesno","Gina_step","fev1_pct_attack","fev1_fvc_attack","blood_eos_attack","feno_attack","smoking_status","Allergic_rhinitis","CRSsNP","CRSwNP")) %>% 
  mutate(Study = "BOOST",
         ACQ_below_1.5 = ifelse(is.na(acq_attack), NA,
                                ifelse(acq_attack < 1.5, "Yes", "No"))) %>% 
  rename(Sequential_number = study_id,
         Gender_0Female_1Male = Sex,
         ACQ_baseline_score_mean = acq_attack,
         Any_severe_attack_previous_12m_0no_1yes = Previous_asthma_attack_yesno,
         Treatment_step = Gina_step,
         FEV1_preBD_PCT_Baseline = fev1_pct_attack,
         Pre_FEV1_FVC_ratio = fev1_fvc_attack,
         Blood_Eos_baseline_x10_9_cells_per_L = blood_eos_attack,
         FeNO_baseline_ppb = feno_attack,
         Smoking_0never_1ex_2current = smoking_status,
         AllergicRhinitis__0no_1yes_9999notknown = Allergic_rhinitis)

val_lab(Data.oracle2.full.master$Any_severe_attack_previous_12m_0no_1yes) <- make_labels("
                                           0 No
                                           1 Yes
                                            ")
Data.oracle2.full.master$Any_severe_attack_previous_12m_0no_1yes <- as.factor(Data.oracle2.full.master$Any_severe_attack_previous_12m_0no_1yes)

val_lab(Data.oracle2.full.master$Smoking_0never_1ex_2current) <- make_labels("
                                           0 Never
                                           1 Ex
                                           2 Current
                                            ")
Data.oracle2.full.master$Smoking_0never_1ex_2current <- as.factor(Data.oracle2.full.master$Smoking_0never_1ex_2current)

val_lab(Data.oracle2.full.master$AllergicRhinitis__0no_1yes_9999notknown) <- make_labels("
                                           0 No
                                           1 Yes
                                            ")
Data.oracle2.full.master$AllergicRhinitis__0no_1yes_9999notknown <- as.factor(Data.oracle2.full.master$AllergicRhinitis__0no_1yes_9999notknown)


val_lab(Data.oracle2.anyacq.master$Any_severe_attack_previous_12m_0no_1yes) <- make_labels("
                                           0 No
                                           1 Yes
                                            ")
Data.oracle2.anyacq.master$Any_severe_attack_previous_12m_0no_1yes <- as.factor(Data.oracle2.anyacq.master$Any_severe_attack_previous_12m_0no_1yes)

val_lab(Data.oracle2.anyacq.master$Smoking_0never_1ex_2current) <- make_labels("
                                           0 Never
                                           1 Ex
                                           2 Current
                                            ")
Data.oracle2.anyacq.master$Smoking_0never_1ex_2current <- as.factor(Data.oracle2.anyacq.master$Smoking_0never_1ex_2current)

val_lab(Data.oracle2.anyacq.master$AllergicRhinitis__0no_1yes_9999notknown) <- make_labels("
                                           0 No
                                           1 Yes
                                            ")
Data.oracle2.anyacq.master$AllergicRhinitis__0no_1yes_9999notknown <- as.factor(Data.oracle2.anyacq.master$AllergicRhinitis__0no_1yes_9999notknown)

Data.boost.master$Smoking_0never_1ex_2current <- ifelse( Data.boost.master$Smoking_0never_1ex_2current == "Never","Never", "Ex"          )



# Combine all datasets
Data.master <- rbind(
  Data.oracle2.full.master,
  Data.oracle2.anyacq.master,
  Data.oracle.dream.placebo.master,
  Data.oracle.dream.mepo.master,
  Data.raw.raspuk.master,
  Data.Prisma.master,
  Data.boost.master
)

# Set factor levels to control column order
Data.master$Study <- factor(Data.master$Study, 
                             levels = c("ORACLE2 - Full", "ORACLE2 - Any ACQ",
                                        "DREAM - Placebo", "DREAM - Mepo",
                                        "SA-OIT", "PRISMA", "BOOST"))

Data.master$Treatment_step <- as.factor(Data.master$Treatment_step)

# Convert ACQ_below_1.5 to factor
Data.master$ACQ_below_1.5 <- as.factor(Data.master$ACQ_below_1.5)

table1(~ Gender_0Female_1Male + Age + BMI + ACQ_baseline_score_mean + ACQ_below_1.5 + Any_severe_attack_previous_12m_0no_1yes + Treatment_step + FEV1_preBD_PCT_Baseline + Pre_FEV1_FVC_ratio + Blood_Eos_baseline_x10_9_cells_per_L + FeNO_baseline_ppb + Smoking_0never_1ex_2current + AllergicRhinitis__0no_1yes_9999notknown + CRSsNP + CRSwNP | Study, 
       data = Data.master, overall = F, 
       extra.col = list("P-value" = pvalue2), 
       render.continuous = c( .="Median [Q1 - Q3]"), 
       render.categorical = "FREQ (PCTnoNA%)", 
       topclass = "Rtable1-grid Rtable1-shade Rtable1-times")
```

### Parametric - Not Imputed

```{r, echo = FALSE}

table1(~ Gender_0Female_1Male + Age + BMI + ACQ_baseline_score_mean + Any_severe_attack_previous_12m_0no_1yes + Treatment_step + FEV1_preBD_PCT_Baseline + Pre_FEV1_FVC_ratio + Blood_Eos_baseline_x10_9_cells_per_L + FeNO_baseline_ppb + Smoking_0never_1ex_2current + AllergicRhinitis__0no_1yes_9999notknown + CRSsNP + CRSwNP | Study, 
       data = Data.master, overall = F, 
       extra.col = list("P-value" = pvalue4), 
       render.continuous = c( .="Median [Q1 - Q3]"), 
       render.categorical = "FREQ (PCTnoNA%)", 
       topclass = "Rtable1-grid Rtable1-shade Rtable1-times")
```
