---
title: "Statistical analyzes : Simon Couillard <br><br>"
subtitle:
  "Project : ACQ5 multicohort <br>"
author: "Reference number : 1206 <br><br> By : Pascal Charland, Samuel Lemaire-Paquette, and Simon Couillard,"
date: " Last update : `r Sys.Date()`<br><br>"
output: 
  html_document:
    theme : united
    df_print: paged
    toc: true
    toc_depth: 1
    number_sections: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
# Define variable containing url
url <- "https://www.crchus.ca/clients/SanteEstrie/Sous-sites/Centres_de_recherche/CRCHUS/Medias-publications/signature_CR-CHUS_rvb-s2.jpg"
```

```{=html}
<script>
   $(document).ready(function() {
     $head = $('#header');
     $head.prepend('<img src="`r url`" style=\"float: right;width: 200px;\"/>')
   });
</script>
```

```{=html}
<style>
body {
text-align: justify}
</style>
```

```{=html}
<style type="text/css">
  body{
  font-size: 12pt;
}
</style>
```

```{=html}
<style type="text/css">
.main-container {
  max-width: 1600px;
  margin-left: auto;
  margin-right: auto;
}
</style>
```

```{r Packages setup, include = FALSE}

options(scipen=999)

load_packages <- function(packages) {
  # Check which packages are not installed
  not_installed <- packages[!(packages %in% installed.packages()[,"Package"])]
  
  # Install missing packages if any
  if(length(not_installed) > 0) {
    install.packages(not_installed, dependencies = TRUE)
  }
  
  # Suppress package loading messages
  suppressPackageStartupMessages({
    invisible(lapply(packages, library, character.only = TRUE))
  })
}

# List of packages
Packages <- c(
  "readxl","haven","ggplot2","dplyr","MASS","summarytools","car","lme4","lmerTest","nlme","multcomp","emmeans","ggeffects","ggiraphExtra","sjPlot","sjmisc","qwraps2","extrafont","openxlsx","igraph","gimme","purrr","ggthemes","plotly","knitr","gmodels","tidyverse","infer","skimr","pander", "RColorBrewer","expss","ggrepel","fmsb","lubridate","mgcv","nlme","glmmTMB","lmerTest","mice","table1","splines", "forplo", "cowplot","forestplot",'gridExtra','grid'
)

# Load packages
load_packages(Packages)

```

```{r p-value function, include = FALSE}

# P-value function for tables

pvalue <- function(x, ...) {
    y <- unlist(x)
    g <- factor(rep(1:length(x), times=sapply(x, length)))
    if (is.numeric(y)) {
        p <- wilcox.test(y ~ g)$p.value
    } else {
        p <- fisher.test(table(y, g))$p.value
    }
    c("", sub("<", "&lt;", format.pval(p, digits=3, eps=0.001)))
}

pvalue2 <- function(x, ...) {
  y <- unlist(x)
  g <- factor(rep(1:length(x), times=sapply(x, length)))
  
  if (is.numeric(y)) {
    p <- kruskal.test(y ~ g)$p.value
  } else {
    p <- fisher.test(table(y, g), simulate.p.value = TRUE, B = 20000)$p.value
  }
  
  c("", sub("<", "&lt;", format.pval(p, digits=3, eps=0.001)))
}


pvalue3 <- function(x, ...) {
    y <- unlist(x)
    g <- factor(rep(1:length(x), times=sapply(x, length)))
    if (is.numeric(y)) {
        p <- t.test(y ~ g)$p.value
    } else {
        p <- fisher.test(table(y, g))$p.value
    }
    c("", sub("<", "&lt;", format.pval(p, digits=3, eps=0.001)))
}

pvalue4 <- function(x, ...) {
  y <- unlist(x)
  g <- factor(rep(1:length(x), times = sapply(x, length)))
  
  if (is.numeric(y)) {
    # Use one-way ANOVA instead of Kruskal-Wallis
    p <- summary(aov(y ~ g))[[1]][["Pr(>F)"]][1]
  } else {
    p <- fisher.test(table(y, g), simulate.p.value = TRUE, B = 20000)$p.value
  }
  
  c("", sub("<", "&lt;", format.pval(p, digits = 3, eps = 0.001)))
}

```

```{r Data PRISMA, include = FALSE}

# PRISMA import

Data.Prisma <- read_excel("2025_03_24_Master.complete_WithSuppression.xlsx") %>% 
  filter(
    redcap_event_name == "visite_1_arm_1",
    HEALTHY_OR_ASTHMA == 1,
    EXCLUDE == 0
  ) %>% 
  mutate(
    Age_at_study_enrollment_rescale10 = Age_at_study_enrollment*0.1,
    BMI_rescale5 = BMI/5,
    Pre_FEV1_PORC_rescale = Pre_FEV1_PORC*-0.1,
    Pre_FEV1_FVC_ratio_rescale = Pre_FEV1_FVC_ratio*-0.1,
    Severe_asthma_attacks_past_12m_cat = ifelse(is.na(Total_exacerbations_12m),0,1),
    BloodEOS_Abso_log10 = log10(BloodEOS_Abso),
    FeNO_log10 = log10(FeNO),
    ACQ5_rescale = acq5_moy * 2,
    SpuEOS_PCT_log10 = log10(SpuEOS_PCT + 0.25),
    PBS_EOTAX_log10 = log10(PBS_EOTAX),
    PBS_IFNg_log10 = log10(PBS_IFNg),
    PBS_IL13_log10 = log10(PBS_IL13),
    PBS_IL4_log10 = log10(PBS_IL4),
    PBS_IL5_log10 = log10(PBS_IL5),
    PBS_IL33_log10 = log10(PBS_IL33),
    PBS_TARC_log10 = log10(PBS_TARC),
    PBS_TNFa_log10 = log10(PBS_TNFa),
    PBS_TSLP_log10 = log10(PBS_TSLP),
    PBS_PGD2_log10 = log10(PBS_PGD2),
    
    Rhinitis_cat = ifelse( Rhinitis == 0,"No","Yes"        ),
    Chronic_sinusitis_cat = ifelse( Chronic_sinusitis == 0,0,1        ),
    Nasal_Polyps_cat = ifelse( Nasal_Polyps == 0,0,1        ),
    
    CRSsNP = ifelse(Chronic_sinusitis_cat == 1 & Nasal_Polyps_cat == 0, "Yes", "No"),
    CRSwNP = ifelse(Chronic_sinusitis_cat == 1 & Nasal_Polyps_cat == 1, "Yes", "No")
  )

val_lab(Data.Prisma$Severe_asthma_attacks_past_12m_cat) <- make_labels("
                                                                       0 No
                                                                       1 Yes   
                                                                       ")

Data.Prisma$Severe_asthma_attacks_past_12m_cat <-  as.factor(Data.Prisma$Severe_asthma_attacks_past_12m_cat )

#Data.Prisma <- within(Data.Prisma, Severe_asthma_attacks_past_12m_cat <- relevel(Severe_asthma_attacks_past_12m_cat, ref = "No"))

val_lab(Data.Prisma$Sex) <- make_labels("
                                        1 Female
                                        2 Male
                                        ")
Data.Prisma$Sex <- as.factor(Data.Prisma$Sex)

Data.Prisma <- within(Data.Prisma, Sex <- relevel(Sex, ref = "Male"))


val_lab(Data.Prisma$Tobacco) <- make_labels("
                                        2 Never
                                        3 Ex
                                        ")
Data.Prisma$Tobacco <- as.factor(Data.Prisma$Tobacco)

Data.Prisma$T2_hothot_med_lowlow_ordinal <- Data.Prisma$T2_hothot_med_lowlow

val_lab(Data.Prisma$T2_hothot_med_lowlow) <- make_labels("
                                                                                 0 T2-Low/Low
                                                                                 1 T2-Mid
                                                                                 2 T2-High/High
                                                                                 ")
Data.Prisma$T2_hothot_med_lowlow <- as.factor(Data.Prisma$T2_hothot_med_lowlow)

Data.Prisma$GINA_Step <- as.factor(Data.Prisma$GINA_Step)

```

```{r Data Oracle, include = FALSE, warning = FALSE}

# ORACLE import

Data.oracle <- read_excel("ORACLE_DATA_not_fleur_V23_2024-07-11_captainadherence and stratos fev1rev multiplied by 100.xlsx") 

Data.oracle[Data.oracle == 9999] <- NA

Data.oracle <- Data.oracle %>% 
  mutate(
    Age_rescale10 = Age / 10,
    BMI_rescale5 = BMI / 5,
    .after = Treatment_arm
  ) %>% 
  mutate(
    BEC_log10 = log10(Blood_Eos_baseline_x10_9_cells_per_L + 0.25),
    FeNO_log10 = log10(FeNO_baseline_ppb + 0.25),
    FEV1_preBD_PCT_Baseline_rescale10 = -FEV1_preBD_PCT_Baseline / 10,
    Pre_FEV1_FVC_ratio = FEV1_preBD_L_Baseline / FVC_preBD_L_Baseline * 100,
    Pre_FEV1_FVC_ratio_rescale10 = -Pre_FEV1_FVC_ratio / 10,
    ACQ5_rescale = ACQ_baseline_score_mean * 2,
    CRSsNP = ifelse(Chronic_Rhinosinusitis_0no_1yes_9999notknown == 1 & Nasal_polyposis_0no_1yes_9999notknown == 0, "Yes", "No"),
    CRSwNP = ifelse(Chronic_Rhinosinusitis_0no_1yes_9999notknown == 1 & Nasal_polyposis_0no_1yes_9999notknown == 1, "Yes", "No"),
  )

Data.oracle$Treatment_step <- as.factor(Data.oracle$Treatment_step)
Data.oracle$CRSsNP <- as.factor(Data.oracle$CRSsNP)
Data.oracle$CRSwNP <- as.factor(Data.oracle$CRSwNP)

val_lab(Data.oracle$Any_severe_attack_previous_12m_0no_1yes) <- make_labels("
                                                                            0 No
                                                                            1 Yes
                                                                            ")
Data.oracle$Any_severe_attack_previous_12m_0no_1yes <- as.factor(Data.oracle$Any_severe_attack_previous_12m_0no_1yes)

val_lab(Data.oracle$Gender_0Female_1Male) <- make_labels("
                                                         0 Female
                                                         1 Male
                                                         ")
Data.oracle$Gender_0Female_1Male <- as.factor(Data.oracle$Gender_0Female_1Male)
Data.oracle$Gender_0Female_1Male <- relevel(Data.oracle$Gender_0Female_1Male, ref = "Male")

val_lab(Data.oracle$Smoking_0never_1ex_2current) <- make_labels("
                                                                0 Never
                                                                1 Ex
                                                                2 Current
                                                                ")
Data.oracle$Smoking_0never_1ex_2current <- as.factor(Data.oracle$Smoking_0never_1ex_2current)

val_lab(Data.oracle$AllergicRhinitis__0no_1yes_9999notknown) <- make_labels("
                                                                            0 No
                                                                            1 Yes
                                                                            ")
Data.oracle$AllergicRhinitis__0no_1yes_9999notknown <- as.factor(Data.oracle$AllergicRhinitis__0no_1yes_9999notknown)

val_lab(Data.oracle$Chronic_Rhinosinusitis_0no_1yes_9999notknown) <- make_labels("
                                                                                 0 No
                                                                                 1 Yes
                                                                                 ")
Data.oracle$Chronic_Rhinosinusitis_0no_1yes_9999notknown <- as.factor(Data.oracle$Chronic_Rhinosinusitis_0no_1yes_9999notknown)

Data.oracle$T2_composite <- ifelse( Data.oracle$FeNO_baseline_ppb < 25 & Data.oracle$Blood_Eos_baseline_x10_9_cells_per_L < 0.15,0, 
                                            ifelse(Data.oracle$FeNO_baseline_ppb >= 35 & Data.oracle$Blood_Eos_baseline_x10_9_cells_per_L >= 0.3,2,1  ))


Data.oracle$T2_composite_ordinal <- Data.oracle$T2_composite


val_lab(Data.oracle$T2_composite) <- make_labels("
                                                                                 0 T2-Low/Low
                                                                                 1 T2-Mid
                                                                                 2 T2-High/High
                                                                                 ")
Data.oracle$T2_composite <- as.factor(Data.oracle$T2_composite)

# Data Dream
Data.oracle.dream <- Data.oracle %>% 
  filter(Enrolled_Trial_name == "DREAM") %>% 
  # Create a column for Placebo or Mepo
  mutate(
    Mepo_placebo = ifelse(Treatment_arm == "Placebo", "Placebo", "Mepo"),
    .after = Treatment_arm
  )

Data.oracle.dream$Mepo_placebo <- as.factor(Data.oracle.dream$Mepo_placebo)
Data.oracle.dream$Mepo_placebo <- relevel(Data.oracle.dream$Mepo_placebo, ref = "Placebo")

```

```{r Data Oracle imputed, include = FALSE, warning = FALSE}

# ORACLE imputed import

Data.oracle.imputed <- read_excel("data_ORACLE_imp_20250102_joined.xlsx")

Data.oracle.imputed[Data.oracle.imputed == "NA"] <- NA

Data.oracle.imputed <- Data.oracle.imputed %>% 
  mutate(
    Age_rescale10 = Age / 10,
    BMI_rescale5 = BMI / 5,
    .after = Treatment_arm
  ) %>% 
  mutate(
    FEV1_preBD_PCT_Baseline_rescale10 = -FEV1_preBD_PCT_Baseline / 10,
    FEV1_FVC_ratio_rescale10 = -FEV1_FVC_ratio / 10,
    ACQ5_rescale = ACQ_baseline_score_mean * 2,
    Severe_asthma_attacks_during_followup_yesno = ifelse(Number_severe_asthma_attacks_during_followup == 0, "No", "Yes"),
    CRSsNP = ifelse(Chronic_Rhinosinusitis_0no_1yes_9999notknown == 1 & Nasal_polyposis_0no_1yes_9999notknown == 0, "Yes", "No"),
    CRSwNP = ifelse(Chronic_Rhinosinusitis_0no_1yes_9999notknown == 1 & Nasal_polyposis_0no_1yes_9999notknown == 1, "Yes", "No")
  )


Data.oracle.imputed$Treatment_step <- as.factor(Data.oracle.imputed$Treatment_step)
Data.oracle.imputed$CRSsNP <- as.factor(Data.oracle.imputed$CRSsNP)
Data.oracle.imputed$CRSwNP <- as.factor(Data.oracle.imputed$CRSwNP)

Data.oracle.imputed$Severe_asthma_attacks_during_followup_yesno <- as.factor(Data.oracle.imputed$Severe_asthma_attacks_during_followup_yesno)

val_lab(Data.oracle.imputed$Gender_0Female_1Male) <- make_labels("
                                                                 0 Female
                                                                 1 Male
                                                                 ")
Data.oracle.imputed$Gender_0Female_1Male <- as.factor(Data.oracle.imputed$Gender_0Female_1Male)
Data.oracle.imputed$Gender_0Female_1Male <- relevel(Data.oracle.imputed$Gender_0Female_1Male, ref = "Male")

Data.oracle.imputed$.imp <- Data.oracle.imputed$.imp -1

Data.imp <- as.mids(Data.oracle.imputed)

# Oracle2 (imputed version) Any ACQ dataset (n = 750)
exclude_trials <- c("AZISAST", "Novel_START", "PRACTICAL", "PACT", "DREAM")
Data.oracle.any.acq <- Data.oracle.imputed %>% filter(Enrolled_Trial_name %in% exclude_trials)

Data.oracle.any.acq$.imp <- Data.oracle.any.acq$.imp -1

Data.imp.anyacq <- as.mids(Data.oracle.any.acq)

```

```{r , include = FALSE}

#RASP-UK import

Data.raw.raspuk <- read_excel("RAW for SPSS v2 INCLUDES RASP-UK.xlsx") %>% 
  filter(Healthy0Asthma1 == 1)

Data.raw.raspuk[Data.raw.raspuk == 99999] <- NA

Data.raw.raspuk <- Data.raw.raspuk %>% 
  mutate(age_rescale10 = AgeD0 / 10,
    BMI_rescale5 = BMI / 5,
    BEC_log10 = log10(Blood_Eos_optimised + 0.25),
    FeNO_log10 = log10(FeNO_optimised + 0.25),
    FEV1_PCTpredicted_rescale10 = - FEV1_PCTpredicted_optimised / 10,
    FEV1_FVC_ratio_rescale10 = - FEV1_FVC_ratio_optimised / 10,
    Previous_asthma_attack_yesno = ifelse(Number_of_Rescue_Steriod_Courses_In_12_Mths == 0, "No", "Yes"),
    Sp_Eos_optimised_log10 = log10(Sp_Eos_optimised + 0.25),
    Eotaxin_3_Optimised_sputm_log10 = log10(`Eotaxin-3_Optimised_sputm`),
    IFNgamma_Optimised_sputm_log10 = log10(IFNgamma_Optimised_sputm),
    IL13_Optimised_sputm_log10 = log10(IL13_Optimised_sputm),
    IL33_Optimised_sputm_log10 = log10(IL33_Optimised_sputm),
    IL4_Optimised_sputm_log10 = log10(IL4_Optimised_sputm),
    IL5_Optimised_sputm_log10 = log10(IL5_Optimised_sputm),
    TARC_Optimised_sputm_log10 = log10(TARC_Optimised_sputm),
    TNFalpha_Optimised_sputm_log10 = log10(TNFalpha_Optimised_sputm),
    TSLP_Optimised_sputm_log10 = log10(TSLP_Optimised_sputm),
    LTE4_Optimised_sputm_log10 = log10(LTE4_Optimised_sputm),
    PGD2_Optimised_sputm_log10 = log10(PGD2_Optimised_sputm))

Data.raw.raspuk$Previous_asthma_attack_yesno <- as.factor(Data.raw.raspuk$Previous_asthma_attack_yesno)

val_lab(Data.raw.raspuk$Sex0M1F) <- make_labels("
                                                0 Male
                                                1 Female
                                                ")
Data.raw.raspuk$Sex0M1F <- as.factor(Data.raw.raspuk$Sex0M1F)

Data.raw.raspuk$Gina_step <- NA

val_lab(Data.raw.raspuk$Smoking_status_0nil_1ex_2active) <- make_labels("
                                                  0 Never
                                                  1 Ex
                                                  2 Current
                                                  ")
Data.raw.raspuk$Smoking_status_0nil_1ex_2active <- as.factor(Data.raw.raspuk$Smoking_status_0nil_1ex_2active)


Data.raw.raspuk$AllergicRhinitis__0no_1yes_9999notknown <- NA

Data.raw.raspuk$CRSsNP <- NA

Data.raw.raspuk$CRSwNP <- NA

```

```{r Data FeNO, echo = FALSE}

Data.feno <- Data.raw.raspuk %>% 
  filter(CohortOxf0Rasp1 == 0)

```

```{r Data RASP-UK, include = FALSE}

Data.raspuk <- Data.raw.raspuk %>% 
  filter(CohortOxf0Rasp1 == 1)

```

```{r Data Boost, include = FALSE}

# BOOST import

Data.boost <- read.csv("Data_Simon_final.csv", header = TRUE, sep = ",", dec = ".") %>% 
  mutate(
    delta = fev1_D7 - fev1_attack,
    Previous_asthma_attack_yesno = ifelse(exac_12m == 0, "No", "Yes"),
    BEC_log10 = log10(blood_eos_attack + 0.25),
    FeNO_log10 = log10(feno_attack),
    Age_rescale10 = Age / 10,
    BMI_rescale5 =  BMI/5,
    Pre_FEV1_PORC_rescale = fev1_pct_attack*-0.1,
    Pre_FEV1_FVC_ratio_rescale = fev1_fvc_attack*-0.1,
    ACQ5_rescale = acq_attack * 2,
    SpuEOS_PCT_log10 = log10(Sputum_eos_attack + 0.25)  )

Data.boost$Previous_asthma_attack_yesno <- as.factor(Data.boost$Previous_asthma_attack_yesno)

val_lab(Data.boost$Sex) <- make_labels("
                                       1 Male
                                       2 Female
                                       ")
Data.boost$Sex <- as.factor(Data.boost$Sex)

val_lab(Data.boost$Biologic) <- make_labels("
                                            1 mepo
                                            2 benra
                                            3 reslizumab
                                            ")
Data.boost$Biologic <- as.factor(Data.boost$Biologic)

val_lab(Data.boost$smoking_status) <- make_labels("
                                                  1 Never
                                                  2 Current
                                                  3 Ex smoker
                                                  ")
Data.boost$smoking_status <- as.factor(Data.boost$smoking_status)

Data.boost$Gina_step <- 5

Data.boost$Gina_step <- as.factor(Data.boost$Gina_step)

```

# Table 1 {.tabset .tabset-fade}

Frequencies and percentages will be presented for nominal variables. Means (standard deviation) and medians [interquartile range] will be presented for continuous variables. For comparisons between two groups, Mann-Whitney U tests will be used for continuous variables and Fisher's exact tests for categorical variables. For comparisons between three or more groups, Kruskal-Wallis tests will be used for continuous variables and Fisher's exact tests for categorical variables.

## Cytokine  {.tabset .tabset-fade}

### Non-paramétrique

```{r Data Cytokine,echo=FALSE}

Data.Prisma.cyto <- Data.Prisma %>% 
  dplyr::select(c("ID","SpuEOS_PCT","PBS_EOTAX","PBS_IFNg","PBS_IL13","PBS_IL4","PBS_IL5","PBS_IL33","PBS_TARC","PBS_TNFa","PBS_TSLP","PBS_PGD2")) %>% 
  mutate( Study = "Prisma")

Data.raw.raspuk.cyto <- Data.raw.raspuk %>% 
  dplyr::select(c("FENO_DiarySeqNumber","Sp_Eos_optimised","Eotaxin-3_Optimised_sputm","IFNgamma_Optimised_sputm","IL13_Optimised_sputm","IL4_Optimised_sputm","IL5_Optimised_sputm","IL33_Optimised_sputm","TARC_Optimised_sputm","TNFalpha_Optimised_sputm","TSLP_Optimised_sputm","PGD2_Optimised_sputm")) %>% 
  mutate( Study = "FeNO+RASP-UK") %>% 
  rename(   ID = FENO_DiarySeqNumber,
            SpuEOS_PCT = Sp_Eos_optimised,
            PBS_EOTAX = `Eotaxin-3_Optimised_sputm`,
            PBS_IFNg = IFNgamma_Optimised_sputm,
            PBS_IL13 = IL13_Optimised_sputm,
            PBS_IL4 = IL4_Optimised_sputm,
            PBS_IL5 = IL5_Optimised_sputm,
            PBS_IL33 = IL33_Optimised_sputm,
            PBS_TARC = TARC_Optimised_sputm,
            PBS_TNFa = TNFalpha_Optimised_sputm,
            PBS_TSLP = TSLP_Optimised_sputm,
            PBS_PGD2 = PGD2_Optimised_sputm)

Data.boost.cyto <- Data.boost %>% 
  dplyr::select( c("study_id","Sputum_eos_attack"))%>% 
  mutate( PBS_EOTAX = NA,
            PBS_IFNg = NA,
            PBS_IL13 = NA,
            PBS_IL4 = NA,
            PBS_IL5 = NA,
            PBS_IL33 = NA,
            PBS_TARC = NA,
            PBS_TNFa = NA,
            PBS_TSLP = NA,
            PBS_PGD2 = NA,
            Study = "BOOST") %>% 
  rename(   ID = study_id,
            SpuEOS_PCT = Sputum_eos_attack )

Data.cyto <- rbind( Data.Prisma.cyto,Data.raw.raspuk.cyto,Data.boost.cyto            )

table1( ~ SpuEOS_PCT + PBS_EOTAX + PBS_IFNg + PBS_IL13 + PBS_IL4 + PBS_IL5 + PBS_IL33 + PBS_TARC + PBS_TNFa + PBS_TSLP + PBS_PGD2   | Study, data = Data.cyto, overall = F, extra.col = list("P-value" = pvalue2), render.continuous=c(.="Median [Q1 - Q3]"), render.categorical="FREQ (PCTnoNA%)", topclass="Rtable1-grid Rtable1-shade Rtable1-times")

```

### Paramétrique

```{r ,echo=FALSE}

table1( ~ SpuEOS_PCT + PBS_EOTAX + PBS_IFNg + PBS_IL13 + PBS_IL4 + PBS_IL5 + PBS_IL33 + PBS_TARC + PBS_TNFa + PBS_TSLP + PBS_PGD2   | Study, data = Data.cyto, overall = F, extra.col = list("P-value" = pvalue4), render.continuous=c(.="Median [Q1 - Q3]"), render.categorical="FREQ (PCTnoNA%)", topclass="Rtable1-grid Rtable1-shade Rtable1-times")

```

## Master   {.tabset .tabset-fade}

### Non-paramétrique

```{r Data Master,echo=FALSE}

Data.oracle.dream.master <- Data.oracle.dream %>% 
  dplyr::select(c("Sequential_number","Gender_0Female_1Male","BMI","ACQ_baseline_score_mean","Any_severe_attack_previous_12m_0no_1yes","Treatment_step","FEV1_preBD_PCT_Baseline","Pre_FEV1_FVC_ratio","Blood_Eos_baseline_x10_9_cells_per_L","FeNO_baseline_ppb","Smoking_0never_1ex_2current","AllergicRhinitis__0no_1yes_9999notknown","CRSsNP","CRSwNP")) %>% 
  mutate( Study = "Dream")

Data.Prisma.master <- Data.Prisma %>% 
  dplyr::select(c("ID","Sex","BMI","acq5_moy","Severe_asthma_attacks_past_12m_cat","GINA_Step","Pre_FEV1_PORC","Pre_FEV1_FVC_ratio","BloodEOS_Abso","FeNO","Tobacco","Rhinitis_cat","CRSsNP","CRSwNP")) %>% 
  mutate( Study = "Prisma") %>% 
  rename(   Sequential_number = ID,
            Gender_0Female_1Male = Sex,
            BMI = BMI,
            ACQ_baseline_score_mean = acq5_moy,
            Any_severe_attack_previous_12m_0no_1yes = Severe_asthma_attacks_past_12m_cat,
            Treatment_step = GINA_Step,
            FEV1_preBD_PCT_Baseline = Pre_FEV1_PORC,
            Pre_FEV1_FVC_ratio = Pre_FEV1_FVC_ratio,
            Blood_Eos_baseline_x10_9_cells_per_L = BloodEOS_Abso,
            FeNO_baseline_ppb = FeNO,
            Smoking_0never_1ex_2current = Tobacco,
            AllergicRhinitis__0no_1yes_9999notknown = Rhinitis_cat,
            CRSsNP = CRSsNP ,
            CRSwNP =  CRSwNP)

Data.boost.master <- Data.boost %>% 
  dplyr::select(c("study_id","Sex","BMI","acq_attack","Previous_asthma_attack_yesno","Gina_step","fev1_pct_attack","fev1_fvc_attack","blood_eos_attack","feno_attack","smoking_status","Allergic_rhinitis","CRSsNP","CRSwNP")) %>% 
  mutate( Study = "BOOST")%>% 
  rename(   Sequential_number = study_id,
            Gender_0Female_1Male = Sex,
            BMI = BMI,
            ACQ_baseline_score_mean = acq_attack,
            Any_severe_attack_previous_12m_0no_1yes = Previous_asthma_attack_yesno,
            Treatment_step = Gina_step,
            FEV1_preBD_PCT_Baseline = fev1_pct_attack,
            Pre_FEV1_FVC_ratio = fev1_fvc_attack,
            Blood_Eos_baseline_x10_9_cells_per_L = blood_eos_attack,
            FeNO_baseline_ppb = feno_attack,
            Smoking_0never_1ex_2current = smoking_status,
            AllergicRhinitis__0no_1yes_9999notknown = Allergic_rhinitis,
            CRSsNP = CRSsNP ,
            CRSwNP =  CRSwNP)

Data.raw.raspuk.master <- Data.raw.raspuk %>% 
  dplyr::select(c("FENO_DiarySeqNumber","Sex0M1F","BMI","ACQ_optimised","Previous_asthma_attack_yesno","Gina_step","FEV1_PCTpredicted_optimised","FEV1_FVC_ratio_optimised","Blood_Eos_optimised","FeNO_optimised","Smoking_status_0nil_1ex_2active","AllergicRhinitis__0no_1yes_9999notknown","CRSsNP","CRSwNP")) %>% 
  mutate( Study = "FeNO+RASP-UK")%>% 
  rename(   Sequential_number = FENO_DiarySeqNumber,
            Gender_0Female_1Male = Sex0M1F,
            BMI = BMI,
            ACQ_baseline_score_mean = ACQ_optimised,
            Any_severe_attack_previous_12m_0no_1yes = Previous_asthma_attack_yesno,
            Treatment_step = Gina_step,
            FEV1_preBD_PCT_Baseline = FEV1_PCTpredicted_optimised,
            Pre_FEV1_FVC_ratio = FEV1_FVC_ratio_optimised,
            Blood_Eos_baseline_x10_9_cells_per_L = Blood_Eos_optimised,
            FeNO_baseline_ppb = FeNO_optimised,
            Smoking_0never_1ex_2current = Smoking_status_0nil_1ex_2active,
            AllergicRhinitis__0no_1yes_9999notknown = AllergicRhinitis__0no_1yes_9999notknown,
            CRSsNP = CRSsNP ,
            CRSwNP =  CRSwNP)

Data.master <- rbind( Data.oracle.dream.master ,Data.Prisma.master, Data.boost.master,Data.raw.raspuk.master   )

Data.master$Treatment_step <- as.factor(Data.master$Treatment_step)

table1( ~ Gender_0Female_1Male + BMI + ACQ_baseline_score_mean + Any_severe_attack_previous_12m_0no_1yes + Treatment_step + FEV1_preBD_PCT_Baseline + Pre_FEV1_FVC_ratio + Blood_Eos_baseline_x10_9_cells_per_L + FeNO_baseline_ppb + Smoking_0never_1ex_2current + AllergicRhinitis__0no_1yes_9999notknown + CRSsNP + CRSwNP   | Study, data = Data.master, overall = F, extra.col = list("P-value" = pvalue2), render.continuous=c( .="Median [Q1 - Q3]"), render.categorical="FREQ (PCTnoNA%)", topclass="Rtable1-grid Rtable1-shade Rtable1-times")

```

### Paramétrique

```{r ,echo=FALSE}

table1( ~ Gender_0Female_1Male + BMI + ACQ_baseline_score_mean + Any_severe_attack_previous_12m_0no_1yes + Treatment_step + FEV1_preBD_PCT_Baseline + Pre_FEV1_FVC_ratio + Blood_Eos_baseline_x10_9_cells_per_L + FeNO_baseline_ppb + Smoking_0never_1ex_2current + AllergicRhinitis__0no_1yes_9999notknown + CRSsNP + CRSwNP   | Study, data = Data.master, overall = F, extra.col = list("P-value" = pvalue4), render.continuous=c( .="Median [Q1 - Q3]"), render.categorical="FREQ (PCTnoNA%)", topclass="Rtable1-grid Rtable1-shade Rtable1-times")

```

## PRISMA  {.tabset .tabset-fade}

```{r ,echo=FALSE}

table1( ~ GINA_Step + Age_at_study_enrollment + Sex + BMI + BloodEOS_Abso + BloodEOS_Abso_log10 + FeNO + FeNO_log10 + Pre_FEV1_PORC + Pre_FEV1_FVC_ratio + Severe_asthma_attacks_past_12m_cat  + acq5_moy + V2_V1_Post_FEV1_L, data= Data.Prisma, render.continuous=c( .="Median [Q1 - Q3]"), render.categorical="FREQ (PCTnoNA%)", topclass="Rtable1-grid Rtable1-shade Rtable1-times")

```

## Oracle DREAM {.tabset .tabset-fade}

### Complete

```{r, echo = FALSE}

table1( ~ Mepo_placebo + Gender_0Female_1Male + BMI + ACQ_baseline_score_mean + Any_severe_attack_previous_12m_0no_1yes + Treatment_step + FEV1_preBD_PCT_Baseline + Pre_FEV1_FVC_ratio + Blood_Eos_baseline_x10_9_cells_per_L + FeNO_baseline_ppb + Smoking_0never_1ex_2current + AllergicRhinitis__0no_1yes_9999notknown + CRSsNP + CRSwNP  + T2_composite, data = Data.oracle.dream, render.continuous=c( .="Median [Q1 - Q3]"), render.categorical="FREQ (PCTnoNA%)", topclass="Rtable1-grid Rtable1-shade Rtable1-times")

```

### Mepo vs placebo  {.tabset .tabset-fade}

#### Non-paramétrique

```{r, echo = FALSE}

Data.dream.nona <- Data.oracle.dream %>% 
  filter(!is.na(Mepo_placebo))

table1( ~ Gender_0Female_1Male + BMI + ACQ_baseline_score_mean + Any_severe_attack_previous_12m_0no_1yes + Treatment_step + FEV1_preBD_PCT_Baseline + Pre_FEV1_FVC_ratio + Blood_Eos_baseline_x10_9_cells_per_L + FeNO_baseline_ppb + Smoking_0never_1ex_2current + AllergicRhinitis__0no_1yes_9999notknown + CRSsNP + CRSwNP  + T2_composite | Mepo_placebo, data = Data.dream.nona, overall = F, extra.col = list("P-value" = pvalue), render.continuous=c( .="Median [Q1 - Q3]"), render.categorical="FREQ (PCTnoNA%)", topclass="Rtable1-grid Rtable1-shade Rtable1-times")

```

#### Paramétrique

```{r, echo = FALSE}

Data.dream.nona <- Data.oracle.dream %>% 
  filter(!is.na(Mepo_placebo))

table1( ~ Gender_0Female_1Male + BMI + ACQ_baseline_score_mean + Any_severe_attack_previous_12m_0no_1yes + Treatment_step + FEV1_preBD_PCT_Baseline + Pre_FEV1_FVC_ratio + Blood_Eos_baseline_x10_9_cells_per_L + FeNO_baseline_ppb + Smoking_0never_1ex_2current + AllergicRhinitis__0no_1yes_9999notknown + CRSsNP + CRSwNP  + T2_composite | Mepo_placebo, data = Data.dream.nona, overall = F, extra.col = list("P-value" = pvalue3), render.continuous=c( .="Median [Q1 - Q3]"), render.categorical="FREQ (PCTnoNA%)", topclass="Rtable1-grid Rtable1-shade Rtable1-times")

```

## ORACLE2 {.tabset .tabset-fade}

### Full

```{r, echo = FALSE}

Data.oracle.imputed.imp1 <- Data.oracle.imputed %>% 
  filter(.imp == 1 )

table1( ~ Enrolled_Trial_name, data = Data.oracle.imputed.imp1, render.continuous=c( .="Median [Q1 - Q3]"), render.categorical="FREQ (PCTnoNA%)", topclass="Rtable1-grid Rtable1-shade Rtable1-times")

```

### Any ACQ

```{r, echo = FALSE}

Data.oracle.any.acq.imp1 <- Data.oracle.any.acq %>% 
  filter(.imp == 1 )

table1( ~ Enrolled_Trial_name, data = Data.oracle.any.acq.imp1, render.continuous=c( .="Median [Q1 - Q3]"), render.categorical="FREQ (PCTnoNA%)", topclass="Rtable1-grid Rtable1-shade Rtable1-times")

```

## FeNO

```{r, echo = FALSE}

table1( ~ AgeD0 + Sex0M1F + BMI + Blood_Eos_optimised + FeNO_optimised + FEV1_PCTpredicted_optimised + FEV1_FVC_ratio_optimised + ACQ_optimised + Previous_asthma_attack_yesno + Sp_Eos_optimised + `Eotaxin-3_Optimised_sputm` + IFNgamma_Optimised_sputm + IL13_Optimised_sputm + IL33_Optimised_sputm + IL4_Optimised_sputm + IL5_Optimised_sputm + TARC_Optimised_sputm + TNFalpha_Optimised_sputm + TSLP_Optimised_sputm + LTE4_Optimised_sputm + PGD2_Optimised_sputm, data = Data.feno, render.continuous=c( .="Median [Q1 - Q3]"), render.categorical="FREQ (PCTnoNA%)", topclass="Rtable1-grid Rtable1-shade Rtable1-times")

```

## RASP-UK

```{r, echo = FALSE}

table1( ~ AgeD0 + Sex0M1F + BMI + Blood_Eos_optimised + FeNO_optimised + FEV1_PCTpredicted_optimised + FEV1_FVC_ratio_optimised + ACQ_optimised + Previous_asthma_attack_yesno + Sp_Eos_optimised + `Eotaxin-3_Optimised_sputm` + IFNgamma_Optimised_sputm + IL13_Optimised_sputm + IL33_Optimised_sputm + IL4_Optimised_sputm + IL5_Optimised_sputm + TARC_Optimised_sputm + TNFalpha_Optimised_sputm + TSLP_Optimised_sputm + LTE4_Optimised_sputm + PGD2_Optimised_sputm, data = Data.raspuk, render.continuous=c( .="Median [Q1 - Q3]"), render.categorical="FREQ (PCTnoNA%)", topclass="Rtable1-grid Rtable1-shade Rtable1-times")

```

## FeNo + RASP-UK

```{r, echo = FALSE}

table1( ~ AgeD0 + Sex0M1F + BMI + Blood_Eos_optimised + FeNO_optimised + FEV1_PCTpredicted_optimised + FEV1_FVC_ratio_optimised + ACQ_optimised + Previous_asthma_attack_yesno + Sp_Eos_optimised + `Eotaxin-3_Optimised_sputm` + IFNgamma_Optimised_sputm + IL13_Optimised_sputm + IL33_Optimised_sputm + IL4_Optimised_sputm + IL5_Optimised_sputm + TARC_Optimised_sputm + TNFalpha_Optimised_sputm + TSLP_Optimised_sputm + LTE4_Optimised_sputm + PGD2_Optimised_sputm, data = Data.raw.raspuk, render.continuous=c( .="Median [Q1 - Q3]"), render.categorical="FREQ (PCTnoNA%)", topclass="Rtable1-grid Rtable1-shade Rtable1-times")

```

## BOOST

```{r, echo = FALSE}

table1( ~ Age + Sex + BMI + Gina_step + acq_attack + CRSsNP + CRSwNP + Allergic_rhinitis + smoking_status + fev1_pct_attack + fev1_fvc_attack + blood_eos_attack + BEC_log10 + feno_attack + FeNO_log10 + delta, data = Data.boost, render.continuous=c( .="Median [Q1 - Q3]"), render.categorical="FREQ (PCTnoNA%)", topclass="Rtable1-grid Rtable1-shade Rtable1-times")

```

## MasterV2 {.tabset .tabset-fade}

### Non-paramétrique

```{r, echo = FALSE}
# ORACLE2 - Full (n=6513)

Data.oracle.imputed_abs <- Data.oracle.imputed
Data.oracle.any.acq_abs <- Data.oracle.any.acq

Data.oracle.imputed_abs$Blood_Eos_baseline_x10_9_cells_per_L <-
  10 ^ Data.oracle.imputed$Blood_Eos_baseline_x10_9_cells_per_L_zeroreplaced

Data.oracle.imputed_abs$FeNO_baseline_ppb <-
  10 ^ Data.oracle.imputed$FeNO_baseline_ppb

Data.oracle.any.acq_abs$Blood_Eos_baseline_x10_9_cells_per_L <-
  10 ^ Data.oracle.any.acq$Blood_Eos_baseline_x10_9_cells_per_L_zeroreplaced

Data.oracle.any.acq_abs$FeNO_baseline_ppb <-
  10 ^ Data.oracle.any.acq$FeNO_baseline_ppb

Data.oracle2.full.master <- Data.oracle.imputed_abs %>% 
  filter(.imp == 1) %>%
  dplyr::select(c("Sequential_number","Gender_0Female_1Male","BMI","ACQ_baseline_score_mean","Any_severe_attack_previous_12m_0no_1yes","Treatment_step","FEV1_preBD_PCT_Baseline","FEV1_FVC_ratio","Blood_Eos_baseline_x10_9_cells_per_L","FeNO_baseline_ppb","Smoking_0never_1ex_2current","AllergicRhinitis__0no_1yes_9999notknown","CRSsNP","CRSwNP")) %>% 
  mutate(Study = "ORACLE2 - Full",
         Pre_FEV1_FVC_ratio = FEV1_FVC_ratio,
         ACQ_below_1.5 = ifelse(is.na(ACQ_baseline_score_mean), NA,
                                ifelse(ACQ_baseline_score_mean < 1.5, "Yes", "No"))) %>%
  dplyr::select(-FEV1_FVC_ratio)

# ORACLE2 - Any ACQ (n=750)
Data.oracle2.anyacq.master <- Data.oracle.any.acq_abs %>% 
  filter(.imp == 1) %>%
  dplyr::select(c("Sequential_number","Gender_0Female_1Male","BMI","ACQ_baseline_score_mean","Any_severe_attack_previous_12m_0no_1yes","Treatment_step","FEV1_preBD_PCT_Baseline","FEV1_FVC_ratio","Blood_Eos_baseline_x10_9_cells_per_L","FeNO_baseline_ppb","Smoking_0never_1ex_2current","AllergicRhinitis__0no_1yes_9999notknown","CRSsNP","CRSwNP")) %>% 
  mutate(Study = "ORACLE2 - Any ACQ",
         Pre_FEV1_FVC_ratio = FEV1_FVC_ratio,
         ACQ_below_1.5 = ifelse(is.na(ACQ_baseline_score_mean), NA,
                                ifelse(ACQ_baseline_score_mean < 1.5, "Yes", "No"))) %>%
  dplyr::select(-FEV1_FVC_ratio)

# DREAM - Placebo
Data.oracle.dream.placebo.master <- Data.oracle.dream %>% 
  filter(Mepo_placebo == "Placebo") %>%
  dplyr::select(c("Sequential_number","Gender_0Female_1Male","BMI","ACQ_baseline_score_mean","Any_severe_attack_previous_12m_0no_1yes","Treatment_step","FEV1_preBD_PCT_Baseline","Pre_FEV1_FVC_ratio","Blood_Eos_baseline_x10_9_cells_per_L","FeNO_baseline_ppb","Smoking_0never_1ex_2current","AllergicRhinitis__0no_1yes_9999notknown","CRSsNP","CRSwNP")) %>% 
  mutate(Study = "DREAM - Placebo",
         ACQ_below_1.5 = ifelse(is.na(ACQ_baseline_score_mean), NA,
                                ifelse(ACQ_baseline_score_mean < 1.5, "Yes", "No")))

# DREAM - Mepo
Data.oracle.dream.mepo.master <- Data.oracle.dream %>% 
  filter(Mepo_placebo == "Mepo") %>%
  dplyr::select(c("Sequential_number","Gender_0Female_1Male","BMI","ACQ_baseline_score_mean","Any_severe_attack_previous_12m_0no_1yes","Treatment_step","FEV1_preBD_PCT_Baseline","Pre_FEV1_FVC_ratio","Blood_Eos_baseline_x10_9_cells_per_L","FeNO_baseline_ppb","Smoking_0never_1ex_2current","AllergicRhinitis__0no_1yes_9999notknown","CRSsNP","CRSwNP")) %>% 
  mutate(Study = "DREAM - Mepo",
         ACQ_below_1.5 = ifelse(is.na(ACQ_baseline_score_mean), NA,
                                ifelse(ACQ_baseline_score_mean < 1.5, "Yes", "No")))

# SA-OIT (FeNO + RASP-UK)
Data.raw.raspuk.master <- Data.raw.raspuk %>% 
  dplyr::select(c("FENO_DiarySeqNumber","Sex0M1F","BMI","ACQ_optimised","Previous_asthma_attack_yesno","Gina_step","FEV1_PCTpredicted_optimised","FEV1_FVC_ratio_optimised","Blood_Eos_optimised","FeNO_optimised","Smoking_status_0nil_1ex_2active","AllergicRhinitis__0no_1yes_9999notknown","CRSsNP","CRSwNP")) %>% 
  mutate(Study = "SA-OIT",
         ACQ_below_1.5 = ifelse(is.na(ACQ_optimised), NA,
                                ifelse(ACQ_optimised < 1.5, "Yes", "No"))) %>% 
  rename(Sequential_number = FENO_DiarySeqNumber,
         Gender_0Female_1Male = Sex0M1F,
         ACQ_baseline_score_mean = ACQ_optimised,
         Any_severe_attack_previous_12m_0no_1yes = Previous_asthma_attack_yesno,
         Treatment_step = Gina_step,
         FEV1_preBD_PCT_Baseline = FEV1_PCTpredicted_optimised,
         Pre_FEV1_FVC_ratio = FEV1_FVC_ratio_optimised,
         Blood_Eos_baseline_x10_9_cells_per_L = Blood_Eos_optimised,
         FeNO_baseline_ppb = FeNO_optimised,
         Smoking_0never_1ex_2current = Smoking_status_0nil_1ex_2active)

# PRISMA
Data.Prisma.master <- Data.Prisma %>% 
  dplyr::select(c("ID","Sex","BMI","acq5_moy","Severe_asthma_attacks_past_12m_cat","GINA_Step","Pre_FEV1_PORC","Pre_FEV1_FVC_ratio","BloodEOS_Abso","FeNO","Tobacco","Rhinitis_cat","CRSsNP","CRSwNP")) %>% 
  mutate(Study = "PRISMA",
         ACQ_below_1.5 = ifelse(is.na(acq5_moy), NA,
                                ifelse(acq5_moy < 1.5, "Yes", "No"))) %>% 
  rename(Sequential_number = ID,
         Gender_0Female_1Male = Sex,
         ACQ_baseline_score_mean = acq5_moy,
         Any_severe_attack_previous_12m_0no_1yes = Severe_asthma_attacks_past_12m_cat,
         Treatment_step = GINA_Step,
         FEV1_preBD_PCT_Baseline = Pre_FEV1_PORC,
         Blood_Eos_baseline_x10_9_cells_per_L = BloodEOS_Abso,
         FeNO_baseline_ppb = FeNO,
         Smoking_0never_1ex_2current = Tobacco,
         AllergicRhinitis__0no_1yes_9999notknown = Rhinitis_cat)

# BOOST
Data.boost.master <- Data.boost %>% 
  dplyr::select(c("study_id","Sex","BMI","acq_attack","Previous_asthma_attack_yesno","Gina_step","fev1_pct_attack","fev1_fvc_attack","blood_eos_attack","feno_attack","smoking_status","Allergic_rhinitis","CRSsNP","CRSwNP")) %>% 
  mutate(Study = "BOOST",
         ACQ_below_1.5 = ifelse(is.na(acq_attack), NA,
                                ifelse(acq_attack < 1.5, "Yes", "No"))) %>% 
  rename(Sequential_number = study_id,
         Gender_0Female_1Male = Sex,
         ACQ_baseline_score_mean = acq_attack,
         Any_severe_attack_previous_12m_0no_1yes = Previous_asthma_attack_yesno,
         Treatment_step = Gina_step,
         FEV1_preBD_PCT_Baseline = fev1_pct_attack,
         Pre_FEV1_FVC_ratio = fev1_fvc_attack,
         Blood_Eos_baseline_x10_9_cells_per_L = blood_eos_attack,
         FeNO_baseline_ppb = feno_attack,
         Smoking_0never_1ex_2current = smoking_status,
         AllergicRhinitis__0no_1yes_9999notknown = Allergic_rhinitis)



val_lab(Data.oracle2.full.master$Any_severe_attack_previous_12m_0no_1yes) <- make_labels("
                                           0 No
                                           1 Yes
                                            ")
Data.oracle2.full.master$Any_severe_attack_previous_12m_0no_1yes <- as.factor(Data.oracle2.full.master$Any_severe_attack_previous_12m_0no_1yes)

val_lab(Data.oracle2.full.master$Smoking_0never_1ex_2current) <- make_labels("
                                           0 Never
                                           1 Ex
                                           2 Current
                                            ")
Data.oracle2.full.master$Smoking_0never_1ex_2current <- as.factor(Data.oracle2.full.master$Smoking_0never_1ex_2current)

val_lab(Data.oracle2.full.master$AllergicRhinitis__0no_1yes_9999notknown) <- make_labels("
                                           0 No
                                           1 Yes
                                            ")
Data.oracle2.full.master$AllergicRhinitis__0no_1yes_9999notknown <- as.factor(Data.oracle2.full.master$AllergicRhinitis__0no_1yes_9999notknown)


val_lab(Data.oracle2.anyacq.master$Any_severe_attack_previous_12m_0no_1yes) <- make_labels("
                                           0 No
                                           1 Yes
                                            ")
Data.oracle2.anyacq.master$Any_severe_attack_previous_12m_0no_1yes <- as.factor(Data.oracle2.anyacq.master$Any_severe_attack_previous_12m_0no_1yes)

val_lab(Data.oracle2.anyacq.master$Smoking_0never_1ex_2current) <- make_labels("
                                           0 Never
                                           1 Ex
                                           2 Current
                                            ")
Data.oracle2.anyacq.master$Smoking_0never_1ex_2current <- as.factor(Data.oracle2.anyacq.master$Smoking_0never_1ex_2current)

val_lab(Data.oracle2.anyacq.master$AllergicRhinitis__0no_1yes_9999notknown) <- make_labels("
                                           0 No
                                           1 Yes
                                            ")
Data.oracle2.anyacq.master$AllergicRhinitis__0no_1yes_9999notknown <- as.factor(Data.oracle2.anyacq.master$AllergicRhinitis__0no_1yes_9999notknown)

Data.boost.master$Smoking_0never_1ex_2current <- ifelse( Data.boost.master$Smoking_0never_1ex_2current == "Never","Never", "Ex"          )



# Combine all datasets
Data.master <- rbind(
  Data.oracle2.full.master,
  Data.oracle2.anyacq.master,
  Data.oracle.dream.placebo.master,
  Data.oracle.dream.mepo.master,
  Data.raw.raspuk.master,
  Data.Prisma.master,
  Data.boost.master
)

# Set factor levels to control column order
Data.master$Study <- factor(Data.master$Study, 
                             levels = c("ORACLE2 - Full", "ORACLE2 - Any ACQ",
                                        "DREAM - Placebo", "DREAM - Mepo",
                                        "SA-OIT", "PRISMA", "BOOST"))

Data.master$Treatment_step <- as.factor(Data.master$Treatment_step)

# Convert ACQ_below_1.5 to factor
Data.master$ACQ_below_1.5 <- as.factor(Data.master$ACQ_below_1.5)

table1(~ Gender_0Female_1Male + BMI + ACQ_baseline_score_mean + ACQ_below_1.5 + Any_severe_attack_previous_12m_0no_1yes + Treatment_step + FEV1_preBD_PCT_Baseline + Pre_FEV1_FVC_ratio + Blood_Eos_baseline_x10_9_cells_per_L + FeNO_baseline_ppb + Smoking_0never_1ex_2current + AllergicRhinitis__0no_1yes_9999notknown + CRSsNP + CRSwNP | Study, 
       data = Data.master, overall = F, 
       extra.col = list("P-value" = pvalue2), 
       render.continuous = c( .="Median [Q1 - Q3]"), 
       render.categorical = "FREQ (PCTnoNA%)", 
       topclass = "Rtable1-grid Rtable1-shade Rtable1-times")
```

### Paramétrique

```{r, echo = FALSE}

table1(~ Gender_0Female_1Male + BMI + ACQ_baseline_score_mean + Any_severe_attack_previous_12m_0no_1yes + Treatment_step + FEV1_preBD_PCT_Baseline + Pre_FEV1_FVC_ratio + Blood_Eos_baseline_x10_9_cells_per_L + FeNO_baseline_ppb + Smoking_0never_1ex_2current + AllergicRhinitis__0no_1yes_9999notknown + CRSsNP + CRSwNP | Study, 
       data = Data.master, overall = F, 
       extra.col = list("P-value" = pvalue4), 
       render.continuous = c( .="Median [Q1 - Q3]"), 
       render.categorical = "FREQ (PCTnoNA%)", 
       topclass = "Rtable1-grid Rtable1-shade Rtable1-times")
```

# Spline Curves Analysis: ACQ-5 Mean Score vs Treatment Response {.tabset .tabset-fade}

```{r Theme graphique, include = FALSE}

Data.oracle.dream$Treatment_step <- as.numeric(Data.oracle.dream$Treatment_step)
Data.oracle.imputed$Treatment_step <- as.numeric(Data.oracle.imputed$Treatment_step)
Data.oracle.any.acq$Treatment_step <- as.numeric(Data.oracle.any.acq$Treatment_step)

Data.imp <- as.mids(Data.oracle.imputed)
Data.imp.anyacq <- as.mids(Data.oracle.any.acq)

# Theme for the spline graphs
theme_graph <- theme(
  axis.text.x = element_text(size = 10),
  axis.text.y = element_text(face = "bold", size = 12),
  axis.title = element_text(face = "bold", size = 12),
  plot.title = element_text(face = "bold", size = 20),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.background = element_blank(),
  axis.line = element_line(colour = "black")
)

```

## PRISMA {.tabset}

### Unadjusted

```{r ,echo=FALSE, fig.width= 8 , fig.height= 5}

# Linear regression with spline 3 degree of freedom
model_spline <- lm(V2_V1_Post_FEV1_L ~  ns(acq5_moy, df = 3), data = Data.Prisma)

# Create data frame for the x axis
acq_seq <- data.frame(acq5_moy = seq(min(Data.Prisma$acq5_moy, na.rm = TRUE),
                                     max(Data.Prisma$acq5_moy, na.rm = TRUE), length.out = 100))

# Prediction based on the linear regression
pred <- predict(model_spline, newdata = acq_seq, se.fit = TRUE)

# Add prediction and confidence interval in the data frame
acq_seq$fit <- pred$fit
acq_seq$se.fit <- pred$se.fit
acq_seq$lower <- acq_seq$fit - 1.96 * acq_seq$se.fit
acq_seq$upper <- acq_seq$fit + 1.96 * acq_seq$se.fit

# Graph of the spline with confidence interval
spline_prisma.uni <- ggplot(acq_seq, aes(x = acq5_moy, y = fit)) +
  geom_line(linewidth = 1.2)+
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.1, fill = "black") +
  xlab("\nMean ACQ-5") +
  ylab("Difference in FEV1 postBD (L)\n") +
  scale_y_continuous(limits = c(-0.7, 0.7), breaks = seq(-1, 1, by = 0.1)) +
  theme_sjplot2() +
  theme_graph

spline_prisma.uni

```

### Adjusted v1

Adjusted for BEC, FeNO and FEV1 V1 (Post_FEV1_L).

```{r ,echo=FALSE, fig.width= 8 , fig.height= 5}

# Linear regression adjusted with clinical parameters
mod_prisma <- lm(V2_V1_Post_FEV1_L ~  ns(acq5_moy, df = 3) + BloodEOS_Abso_log10 + FeNO_log10 + Post_FEV1_L, data = Data.Prisma)

# Create data frame for the x axis
acq_seq <- data.frame(acq5_moy = seq(min(Data.Prisma$acq5_moy, na.rm = TRUE),
                                     max(Data.Prisma$acq5_moy, na.rm = TRUE), length.out = 100))

# Add median of all clinical parameters as a baseline
acq_seq$Age_at_study_enrollment <- median(Data.Prisma$Age_at_study_enrollment, na.rm = TRUE)
acq_seq$BMI <- median(Data.Prisma$BMI, na.rm = TRUE)
acq_seq$Sex <- "Female"
acq_seq$GINA_Step <- 3
acq_seq$BloodEOS_Abso_log10 <- median(Data.Prisma$BloodEOS_Abso_log10, na.rm = TRUE)
acq_seq$FeNO_log10 <- median(Data.Prisma$FeNO_log10, na.rm = TRUE)
acq_seq$Pre_FEV1_PORC <- median(Data.Prisma$Pre_FEV1_PORC, na.rm = TRUE)
acq_seq$Pre_FEV1_FVC_ratio <- median(Data.Prisma$Pre_FEV1_FVC_ratio, na.rm = TRUE)
acq_seq$Severe_asthma_attacks_past_12m_cat <- "No"
acq_seq$Post_FEV1_L <- median(Data.Prisma$Post_FEV1_L, na.rm = TRUE)

# Prediction based on the linear regression adjusted
pred <- predict(mod_prisma, newdata = acq_seq, se.fit = TRUE)

# Add prediction and confidence interval in the data frame
acq_seq$fit <- pred$fit
acq_seq$se.fit <- pred$se.fit
acq_seq$lower <- acq_seq$fit - 1.96 * acq_seq$se.fit
acq_seq$upper <- acq_seq$fit + 1.96 * acq_seq$se.fit

# Graph of the spline with confidence interval
spline_prisma <- ggplot(acq_seq, aes(x = acq5_moy, y = fit)) +
  geom_line(linewidth = 1.2)+
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.1, fill = "black") +
  xlab("\nMean ACQ-5") +
  ylab("Adjusted Difference \nin FEV1 postBD (L)\n") +
  scale_y_continuous(limits = c(-0.4, 0.8), breaks = seq(-0.4, 0.8, by = 0.2)) +
  theme_sjplot2() +
  theme_graph

spline_prisma

```

```{r, echo = FALSE, fig.height = 5, fig.width = 10}

# Linear regression adjusted without the spline
mod <- lm(V2_V1_Post_FEV1_L ~  ACQ5_rescale + BloodEOS_Abso_log10 + FeNO_log10 + Post_FEV1_L, data = Data.Prisma)

# Data frame with rate ratio, confidence interval and p-value
Rate_ratio_prisma <- as.data.frame(cbind(round(Confint(mod), 2), "p_value" = round(summary(mod)$coefficients[,4], 3)))[-1,]

# Forest plot
forplo(
  Rate_ratio_prisma[,1:3],
  em = "RR",
  row.labels = c("ACQ5 (per 0.5)", "Blood eos Count (per 10-fold)", "FeNO (per 10-fold)", 'Post FEV1 V1 (L)'),
  xlim = c(-1, 1),
  title = "",
  left.align = TRUE,
  margin.left = 16,
  add.columns = as.data.frame(cbind(rep(NA, 4), format(Rate_ratio_prisma$p_value, nsmall = 3))),
  add.colnames = c("", "P-Value"),
  margin.right = 11
)

```

### T2 composite

Adjusted only for FEV1 V1 (Post_FEV1_L) and T2 composite.

```{r ,echo=FALSE, fig.width= 8 , fig.height= 5}

# Linear regression adjusted with clinical parameters
mod_prisma <- lm(V2_V1_Post_FEV1_L ~  ns(acq5_moy, df = 3) + Post_FEV1_L + T2_hothot_med_lowlow_ordinal, data = Data.Prisma)

# Create data frame for the x axis
acq_seq <- data.frame(acq5_moy = seq(min(Data.Prisma$acq5_moy, na.rm = TRUE),
                                     max(Data.Prisma$acq5_moy, na.rm = TRUE), length.out = 100))

# Add median of all clinical parameters as a baseline
acq_seq$Age_at_study_enrollment <- median(Data.Prisma$Age_at_study_enrollment, na.rm = TRUE)
acq_seq$BMI <- median(Data.Prisma$BMI, na.rm = TRUE)
acq_seq$Sex <- "Female"
acq_seq$GINA_Step <- 3
acq_seq$BloodEOS_Abso_log10 <- median(Data.Prisma$BloodEOS_Abso_log10, na.rm = TRUE)
acq_seq$FeNO_log10 <- median(Data.Prisma$FeNO_log10, na.rm = TRUE)
acq_seq$Pre_FEV1_PORC <- median(Data.Prisma$Pre_FEV1_PORC, na.rm = TRUE)
acq_seq$Pre_FEV1_FVC_ratio <- median(Data.Prisma$Pre_FEV1_FVC_ratio, na.rm = TRUE)
acq_seq$Severe_asthma_attacks_past_12m_cat <- "No"
acq_seq$Post_FEV1_L <- median(Data.Prisma$Post_FEV1_L, na.rm = TRUE)
acq_seq$T2_hothot_med_lowlow_ordinal <- median(Data.Prisma$T2_hothot_med_lowlow_ordinal, na.rm = TRUE)

# Prediction based on the linear regression adjusted
pred <- predict(mod_prisma, newdata = acq_seq, se.fit = TRUE)

# Add prediction and confidence interval in the data frame
acq_seq$fit <- pred$fit
acq_seq$se.fit <- pred$se.fit
acq_seq$lower <- acq_seq$fit - 1.96 * acq_seq$se.fit
acq_seq$upper <- acq_seq$fit + 1.96 * acq_seq$se.fit

# Graph of the spline with confidence interval
spline_prismav2 <- ggplot(acq_seq, aes(x = acq5_moy, y = fit)) +
  geom_line(linewidth = 1.2)+
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.1, fill = "black") +
  xlab("\nMean ACQ-5") +
  ylab("Adjusted Difference \nin FEV1 postBD (L)\n") +
  scale_y_continuous(limits = c(-0.4, 0.8), breaks = seq(-0.4, 0.8, by = 0.2)) +
  theme_sjplot2() +
  theme_graph

spline_prismav2

```

```{r, echo = FALSE, fig.height = 5, fig.width = 10}

# Linear regression adjusted without the spline
mod <- lm(V2_V1_Post_FEV1_L ~  ACQ5_rescale + Post_FEV1_L + T2_hothot_med_lowlow_ordinal, data = Data.Prisma)

# Data frame with rate ratio, confidence interval and p-value
Rate_ratio_prisma <- as.data.frame(cbind(round(Confint(mod), 2), "p_value" = round(summary(mod)$coefficients[,4], 3)))[-1,]

# Forest plot
forplo(
  Rate_ratio_prisma[,1:3],
  em = "RR",
  row.labels = c("ACQ5 (per 0.5)", 'Post FEV1 V1 (L)','T2 ordinal'),
  xlim = c(-1, 1),
  title = "",
  left.align = TRUE,
  margin.left = 16,
  add.columns = as.data.frame(cbind(rep(NA, 3), format(Rate_ratio_prisma$p_value, nsmall = 3))),
  add.colnames = c("", "P-Value"),
  margin.right = 11
)

```

## BOOST {.tabset}

### Unadjusted

```{r ,echo=FALSE, fig.width= 8 , fig.height= 5}

# Linear regression with spline 3 degree of freedom
model_spline <- lm(delta ~ ns(acq_attack, df = 3), data = Data.boost)

# Create data frame for the x axis
acq_seq <- data.frame(acq_attack = seq(min(Data.boost$acq_attack, na.rm = TRUE),
                                     max(Data.boost$acq_attack, na.rm = TRUE), length.out = 100))

# Prediction based on the linear regression
pred <- predict(model_spline, newdata = acq_seq, se.fit = TRUE)

# Add prediction and confidence interval in the data frame
acq_seq <- acq_seq %>% 
  mutate(
    fit = pred$fit,
    se.fit = pred$se.fit,
    lower = fit - 1.96 * se.fit,
    upper = fit + 1.96 * se.fit
  )

# Graph of the spline with confidence interval
spline_boost.uni <-  ggplot(acq_seq, aes(x = acq_attack, y = fit)) +
  geom_line(linewidth = 1.2)+
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.1, fill = "black") +
  xlab("\nMean ACQ-5") +
  ylab("Difference \nin FEV1 postBD (L)\n") +
  scale_y_continuous(limits = c(-10, 10), breaks = seq(-10, 10, by = 0.1)) +
  scale_x_continuous(limits = c(-10, 10), breaks = seq(-10, 10, by = 1)) +
  coord_cartesian(xlim = c(2, 6), ylim = c(-0.4, 1.2)) +
  theme_sjplot2() +
  theme_graph

spline_boost.uni

```

### Adjusted v1

Adjusted for BEC, FeNO and FEV1 V1 (fev1_attack).

```{r ,echo=FALSE, fig.width= 8 , fig.height= 5}

# Linear regression adjusted with clinical parameters
mod_boost <- lm(delta ~ ns(acq_attack, df = 3) + BEC_log10 + FeNO_log10 + fev1_attack, data = Data.boost)

# Create data frame for the x axis
acq_seq <- data.frame(acq_attack = seq(min(Data.boost$acq_attack, na.rm = TRUE),
                                     max(Data.boost$acq_attack, na.rm = TRUE), length.out = 100))

# Add median of all clinical parameters as a baseline
acq_seq <- acq_seq %>% 
  mutate(
    Age_rescale10 = median(Data.boost$Age_rescale10, na.rm = TRUE),
    Sex = "Male",
    BMI_rescale5 = median(Data.boost$BMI_rescale5, na.rm = TRUE),
    BEC_log10 = median(Data.boost$BEC_log10, na.rm = TRUE),
    FeNO_log10 = median(Data.boost$FeNO_log10, na.rm = TRUE),
    Pre_FEV1_PORC_rescale = median(Data.boost$Pre_FEV1_PORC_rescale, na.rm = TRUE),
    Pre_FEV1_FVC_ratio_rescale = median(Data.boost$Pre_FEV1_FVC_ratio_rescale, na.rm = TRUE),
    Previous_asthma_attack_yesno = "No",
    fev1_attack = median(Data.boost$fev1_attack, na.rm = TRUE),
  )

# Prediction based on the linear regression adjusted
pred <- predict(mod_boost, newdata = acq_seq, se.fit = TRUE)

# Add prediction and confidence interval in the data frame
acq_seq <- acq_seq %>% 
  mutate(
    fit = pred$fit,
    se.fit = pred$se.fit,
    lower = fit - 1.96 * se.fit,
    upper = fit + 1.96 * se.fit
  )

# Graph of the spline with confidence interval
spline_boost <- ggplot(acq_seq, aes(x = acq_attack, y = fit)) +
  geom_line(linewidth = 1.2)+
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.1, fill = "black") +
  xlab("\nMean ACQ-5") +
  ylab(NULL) +
  scale_y_continuous(limits = c(-10, 10), breaks = seq(-10, 10, by = 0.2)) +
  scale_x_continuous(limits = c(-10, 10), breaks = seq(-10, 10, by = 1)) +
  coord_cartesian(xlim = c(1, 5), ylim = c(-0.4, 0.8)) +
  theme_sjplot2() +
  theme_graph

spline_boost + ylab("Adjusted Difference \nin FEV1 postBD (L)\n")

```

```{r, echo = FALSE, fig.height = 5, fig.width = 10}

# Linear regression adjusted without the spline
mod <- lm(delta ~ ACQ5_rescale  + BEC_log10 + FeNO_log10 + fev1_attack, data = Data.boost)

# Data frame with rate ratio, confidence interval and p-value
Rate_ratio_boost <- as.data.frame(cbind(round(Confint(mod), 2), "p_value" = round(summary(mod)$coefficients[,4], 3)))[2:5,]

# Forest plot
forplo(
  Rate_ratio_boost[,1:3],
  em = "RR",
  row.labels = c("ACQ5 (per 0.5)", "Blood eos Count (per 10-fold)", "FeNO (per 10-fold)", 'Post FEV1 V1 (L)'),
  xlim = c(-1, 1),
  title = "",
  left.align = TRUE,
  margin.left = 16,
  add.columns = as.data.frame(cbind(rep(NA, 4), format(Rate_ratio_boost$p_value, nsmall = 3))),
  add.colnames = c("", "P-Value"),
  margin.right = 11
)

```

## Oracle DREAM {.tabset}

### Unadjusted

```{r, echo = FALSE, fig.height = 6, fig.width = 8, warning=FALSE}

# Negative binomial regression with spline 
mod <- glm.nb(Number_severe_asthma_attacks_during_followup ~ ns(ACQ_baseline_score_mean, df = 3)*Mepo_placebo + offset(log(Follow_up_duration_days)), data = Data.oracle.dream)

# Create data frame for the x axis 
acq_seq <- data.frame(ACQ_baseline_score_mean = seq(min(Data.oracle.dream$ACQ_baseline_score_mean, na.rm = TRUE),
                                                    max(Data.oracle.dream$ACQ_baseline_score_mean, na.rm = TRUE),
                                                    length = 100))

# Baseline parameters for placebo
acq_seq_1 <- acq_seq %>% 
  mutate(
    Mepo_placebo = "Placebo",
    Follow_up_duration_days = median(Data.oracle.dream$Follow_up_duration_days, na.rm = TRUE)
  )

# Baseline parameters for mepo
acq_seq_2 <- acq_seq %>% 
  mutate(
    Mepo_placebo = "Mepo",
    Follow_up_duration_days = median(Data.oracle.dream$Follow_up_duration_days, na.rm = TRUE)
  )

# Prediction for placebo
pred1 <- predict(mod, newdata = acq_seq_1, se.fit = TRUE)
# Prediction for mepo
pred2 <- predict(mod, newdata = acq_seq_2, se.fit = TRUE)

# Add prediction and confidence interval in the data frame for placebo
acq_seq_1 <- acq_seq_1 %>% 
  mutate(
    fit_link = pred1$fit,
    se.fit_link = pred1$se.fit,
    CI_low_link = fit_link - 1.96 * se.fit_link,
    CI_up_link = fit_link + 1.96 * se.fit_link,
    fit = exp(fit_link),
    CI_low = exp(CI_low_link),
    CI_up = exp(CI_up_link)
  )

# Add prediction and confidence interval in the data frame for mepo
acq_seq_2 <- acq_seq_2 %>% 
  mutate(
    fit_link = pred2$fit,
    se.fit_link = pred2$se.fit,
    CI_low_link = fit_link - 1.96 * se.fit_link,
    CI_up_link = fit_link + 1.96 * se.fit_link,
    fit = exp(fit_link),
    CI_low = exp(CI_low_link),
    CI_up = exp(CI_up_link)
  )

# Bind data frame placebo and mepo
acq_seq <- rbind(acq_seq_1, acq_seq_2)

acq_seq$Mepo_placebo <- relevel(as.factor(acq_seq$Mepo_placebo), "Placebo")

# Graph of the splines with confidence intervals
acq_seq %>% ggplot(aes(x = ACQ_baseline_score_mean, y = fit, colour = Mepo_placebo)) + 
  geom_line(linewidth = 1.2) +
  geom_ribbon(aes(ymin = CI_low, ymax = CI_up, fill = Mepo_placebo), alpha = 0.1) +
  labs(
    title = "",
    x = "\nMean ACQ5",
    y = " Estimated annualised severe \nasthma attack rate \n"
  ) +
  theme_sjplot2() +
  scale_color_manual(values = c("Placebo" = "#00a1d5", "Mepo" = "#b24745")) +
  scale_fill_manual(values = c("Placebo" = "#00a1d5", "Mepo" = "#b24745")) +
  scale_y_continuous(limits = c(0, 100), breaks = seq(0, 15, by = 1)) +
  scale_x_continuous(limits = c(0, 6), breaks = seq(0, 15, by = 1)) +
  coord_cartesian(xlim = c(1, 5), ylim = c(0, 10)) +
  theme_graph +
  theme(
    legend.title = element_blank()
  )
  

```

### Adjusted - v1

GINA steps = 5 for all observations and no data for Allergic Rhinitis, variables weren't included in the final model.

```{r, echo = FALSE, fig.height = 6, fig.width = 8, warning = FALSE}

# Negative binomial regression with spline, adjusted with clinical parameter, interaction ACQ*Mepo_placebo, interaction Eos*Mepo_placebo, interaction FeNO*Mepo_placebo
mod_dream <- glm.nb(Number_severe_asthma_attacks_during_followup ~ ns(ACQ_baseline_score_mean, df = 3)*Mepo_placebo + Age_rescale10 + Gender_0Female_1Male + BMI_rescale5 + BEC_log10*Mepo_placebo + FeNO_log10*Mepo_placebo + FEV1_preBD_PCT_Baseline_rescale10 + Pre_FEV1_FVC_ratio_rescale10 + Smoking_0never_1ex_2current + CRSsNP + CRSwNP + offset(log(Follow_up_duration_days)), data = Data.oracle.dream)

# Create data frame for the x axis 
acq_seq <- data.frame(ACQ_baseline_score_mean = seq(min(Data.oracle.dream$ACQ_baseline_score_mean, na.rm = TRUE),
                                                    max(Data.oracle.dream$ACQ_baseline_score_mean, na.rm = TRUE),
                                                    length = 100))

# Baseline parameters for placebo
acq_seq_1 <- acq_seq %>% 
  mutate(
    Mepo_placebo = "Placebo",
    Age_rescale10 = median(Data.oracle.dream$Age_rescale10, na.rm = TRUE),
    Gender_0Female_1Male = "Male",
    BMI_rescale5 = median(Data.oracle.dream$BMI_rescale5, na.rm = TRUE),
    BEC_log10 = median(Data.oracle.dream$BEC_log10, na.rm = TRUE),
    FeNO_log10 = median(Data.oracle.dream$FeNO_log10, na.rm = TRUE),
    FEV1_preBD_PCT_Baseline_rescale10 = median(Data.oracle.dream$FEV1_preBD_PCT_Baseline_rescale10, na.rm = TRUE),
    Pre_FEV1_FVC_ratio_rescale10 = median(Data.oracle.dream$Pre_FEV1_FVC_ratio_rescale10, na.rm = TRUE),
    #Any_severe_attack_previous_12m_0no_1yes = "No",
    Smoking_0never_1ex_2current = "Never",
    #AllergicRhinitis__0no_1yes_9999notknown = "No",
    CRSsNP = "No",
    CRSwNP = "No",
    Follow_up_duration_days = median(Data.oracle.dream$Follow_up_duration_days, na.rm = TRUE),
  )

# Baseline parameters for mepo
acq_seq_2 <- acq_seq %>% 
  mutate(
    Mepo_placebo = "Mepo",
    Age_rescale10 = median(Data.oracle.dream$Age_rescale10, na.rm = TRUE),
    Gender_0Female_1Male = "Male",
    BMI_rescale5 = median(Data.oracle.dream$BMI_rescale5, na.rm = TRUE),
    BEC_log10 = median(Data.oracle.dream$BEC_log10, na.rm = TRUE),
    FeNO_log10 = median(Data.oracle.dream$FeNO_log10, na.rm = TRUE),
    FEV1_preBD_PCT_Baseline_rescale10 = median(Data.oracle.dream$FEV1_preBD_PCT_Baseline_rescale10, na.rm = TRUE),
    Pre_FEV1_FVC_ratio_rescale10 = median(Data.oracle.dream$Pre_FEV1_FVC_ratio_rescale10, na.rm = TRUE),
    #Any_severe_attack_previous_12m_0no_1yes = "No",
    Smoking_0never_1ex_2current = "Never",
    #AllergicRhinitis__0no_1yes_9999notknown = "No",
    CRSsNP = "No",
    CRSwNP = "No",
    Follow_up_duration_days = median(Data.oracle.dream$Follow_up_duration_days, na.rm = TRUE),
  )

# Prediction for placebo
pred1 <- predict(mod_dream, newdata = acq_seq_1, se.fit = TRUE)
# Prediction for mepo
pred2 <- predict(mod_dream, newdata = acq_seq_2, se.fit = TRUE)

# Add prediction and confidence interval in the data frame for placebo
acq_seq_1 <- acq_seq_1 %>% 
  mutate(
    fit_link = pred1$fit,
    se.fit_link = pred1$se.fit,
    CI_low_link = fit_link - 1.96 * se.fit_link,
    CI_up_link = fit_link + 1.96 * se.fit_link,
    fit = exp(fit_link),
    CI_low = exp(CI_low_link),
    CI_up = exp(CI_up_link)
  )

# Add prediction and confidence interval in the data frame for mepo
acq_seq_2 <- acq_seq_2 %>% 
  mutate(
    fit_link = pred2$fit,
    se.fit_link = pred2$se.fit,
    CI_low_link = fit_link - 1.96 * se.fit_link,
    CI_up_link = fit_link + 1.96 * se.fit_link,
    fit = exp(fit_link),
    CI_low = exp(CI_low_link),
    CI_up = exp(CI_up_link)
  )

# Bind data frame placebo and mepo
acq_seq <- rbind(acq_seq_1, acq_seq_2)

acq_seq$Mepo_placebo <- relevel(as.factor(acq_seq$Mepo_placebo), "Placebo")

# Graph of the splines with confidence intervals
spline_dream <- acq_seq %>% ggplot(aes(x = ACQ_baseline_score_mean, y = fit, colour = Mepo_placebo)) + 
  geom_line(linewidth = 1.2) +
  geom_ribbon(aes(ymin = CI_low, ymax = CI_up, fill = Mepo_placebo), alpha = 0.1) +
  xlab("\nMean ACQ5") + 
  ylab(" Estimated annualised severe \nasthma attack rate \n") + 
  theme_sjplot2() +
  scale_color_manual(values = c("Placebo" = "#00a1d5", "Mepo" = "#b24745")) +
  scale_fill_manual(values = c("Placebo" = "#00a1d5", "Mepo" = "#b24745")) +
  scale_y_continuous(limits = c(0, 20), breaks = seq(0, 15, by = 0.5)) +
  scale_x_continuous(limits = c(0, 6), breaks = seq(0, 15, by = 1)) +
  coord_cartesian(xlim = c(1, 5), ylim = c(0, 5)) +
  theme_graph  + 
  theme(
    legend.title = element_blank()
  )

spline_dream + labs(y = " Estimated annualised severe \nasthma attack rate \n")

```

```{r, echo = FALSE, fig.height = 6, fig.width = 10, warning=FALSE}

# Negative binomial regression, adjusted with clinical parameter, interaction ACQ*Mepo, interaction Eos*Mepo, interaction FeNO*Mepo
mod <- glm.nb(Number_severe_asthma_attacks_during_followup ~ ACQ5_rescale*Mepo_placebo + Age_rescale10 + Gender_0Female_1Male + BMI_rescale5 + BEC_log10*Mepo_placebo + FeNO_log10*Mepo_placebo + FEV1_preBD_PCT_Baseline_rescale10 + Pre_FEV1_FVC_ratio_rescale10 + Smoking_0never_1ex_2current + CRSsNP + CRSwNP + offset(log(Follow_up_duration_days)), data = Data.oracle.dream)

# Data frame with rate ratio, confidence interval and p-value
Rate_ratio_dream <- as.data.frame(cbind(round(exp(Confint(mod)), 2), "p_value" = round(summary(mod)$coefficients[,4], 3)))[-1,]

# Forest plot
forplo(
  Rate_ratio_dream[,1:3],
  em = "RR",
  row.labels = c("ACQ5 (per 0.5)", 
                 "Mepo", 
                 "Age (per 10 years)", 
                 "Sex - Female", 
                 "BMI (per 5)", 
                 "Blood eos Count (per 10-fold)", 
                 "FeNO (per 10-fold)", 
                 "FEV1% preBD (per 10% decrease)", 
                 "FEV1/FVC preBD (per 10% decrease)", 
                 "Smoking history",
                 "CRSsNP",
                 "CRSwNP",
                 "Mepo * ACQ5 (per 0.5)", 
                 "Mepo * Blood Eos (per 10-fold)",
                 'Mepo * FeNO (per 10-fold)'),
  xlim = c(0.2, 5),
  title = "",
  left.align = TRUE,
  margin.left = 16,
  add.columns = as.data.frame(cbind(rep(NA, 15), format(Rate_ratio_dream$p_value, nsmall = 3))),
  add.colnames = c("", "P-Value"),
  margin.right = 11
)

```

### Adjusted - T2 composite

GINA steps = 5 for all observations and no data for Allergic Rhinitis, variables weren't included in the final model.

Includes T2_composite (T2 Low/Low, T2-Mid, T2-High/High) treated as an ordinal variable where it replaces BEC × FeNO.

```{r, echo = FALSE, fig.height = 6, fig.width = 10, warning = FALSE}

# Negative binomial regression with spline, adjusted with clinical parameter, interaction ACQ*Mepo_placebo, interaction Eos*Mepo_placebo, interaction FeNO*Mepo_placebo
mod_dream <- glm.nb(Number_severe_asthma_attacks_during_followup ~ ns(ACQ_baseline_score_mean, df = 3)*Mepo_placebo + Age_rescale10 + Gender_0Female_1Male + BMI_rescale5 + T2_composite_ordinal*Mepo_placebo + FEV1_preBD_PCT_Baseline_rescale10 + Pre_FEV1_FVC_ratio_rescale10 + Smoking_0never_1ex_2current + CRSsNP + CRSwNP + offset(log(Follow_up_duration_days)), data = Data.oracle.dream)

# Create data frame for the x axis 
acq_seq <- data.frame(ACQ_baseline_score_mean = seq(min(Data.oracle.dream$ACQ_baseline_score_mean, na.rm = TRUE),
                                                    max(Data.oracle.dream$ACQ_baseline_score_mean, na.rm = TRUE),
                                                    length = 100))

# Baseline parameters for placebo
acq_seq_1 <- acq_seq %>% 
  mutate(
    Mepo_placebo = "Placebo",
    Age_rescale10 = median(Data.oracle.dream$Age_rescale10, na.rm = TRUE),
    Gender_0Female_1Male = "Male",
    BMI_rescale5 = median(Data.oracle.dream$BMI_rescale5, na.rm = TRUE),
    BEC_log10 = median(Data.oracle.dream$BEC_log10, na.rm = TRUE),
    FeNO_log10 = median(Data.oracle.dream$FeNO_log10, na.rm = TRUE),
    FEV1_preBD_PCT_Baseline_rescale10 = median(Data.oracle.dream$FEV1_preBD_PCT_Baseline_rescale10, na.rm = TRUE),
    Pre_FEV1_FVC_ratio_rescale10 = median(Data.oracle.dream$Pre_FEV1_FVC_ratio_rescale10, na.rm = TRUE),
    #Any_severe_attack_previous_12m_0no_1yes = "No",
    Smoking_0never_1ex_2current = "Never",
    #AllergicRhinitis__0no_1yes_9999notknown = "No",
    CRSsNP = "No",
    CRSwNP = "No",
    Follow_up_duration_days = median(Data.oracle.dream$Follow_up_duration_days, na.rm = TRUE),
    T2_composite_ordinal = median(Data.oracle.dream$T2_composite_ordinal, na.rm = TRUE)
  )

# Baseline parameters for mepo
acq_seq_2 <- acq_seq %>% 
  mutate(
    Mepo_placebo = "Mepo",
    Age_rescale10 = median(Data.oracle.dream$Age_rescale10, na.rm = TRUE),
    Gender_0Female_1Male = "Male",
    BMI_rescale5 = median(Data.oracle.dream$BMI_rescale5, na.rm = TRUE),
    BEC_log10 = median(Data.oracle.dream$BEC_log10, na.rm = TRUE),
    FeNO_log10 = median(Data.oracle.dream$FeNO_log10, na.rm = TRUE),
    FEV1_preBD_PCT_Baseline_rescale10 = median(Data.oracle.dream$FEV1_preBD_PCT_Baseline_rescale10, na.rm = TRUE),
    Pre_FEV1_FVC_ratio_rescale10 = median(Data.oracle.dream$Pre_FEV1_FVC_ratio_rescale10, na.rm = TRUE),
    #Any_severe_attack_previous_12m_0no_1yes = "No",
    Smoking_0never_1ex_2current = "Never",
    #AllergicRhinitis__0no_1yes_9999notknown = "No",
    CRSsNP = "No",
    CRSwNP = "No",
    Follow_up_duration_days = median(Data.oracle.dream$Follow_up_duration_days, na.rm = TRUE),
    T2_composite_ordinal = median(Data.oracle.dream$T2_composite_ordinal, na.rm = TRUE)
  )

# Prediction for placebo
pred1 <- predict(mod_dream, newdata = acq_seq_1, se.fit = TRUE)
# Prediction for mepo
pred2 <- predict(mod_dream, newdata = acq_seq_2, se.fit = TRUE)

# Add prediction and confidence interval in the data frame for placebo
acq_seq_1 <- acq_seq_1 %>% 
  mutate(
    fit_link = pred1$fit,
    se.fit_link = pred1$se.fit,
    CI_low_link = fit_link - 1.96 * se.fit_link,
    CI_up_link = fit_link + 1.96 * se.fit_link,
    fit = exp(fit_link),
    CI_low = exp(CI_low_link),
    CI_up = exp(CI_up_link)
  )

# Add prediction and confidence interval in the data frame for mepo
acq_seq_2 <- acq_seq_2 %>% 
  mutate(
    fit_link = pred2$fit,
    se.fit_link = pred2$se.fit,
    CI_low_link = fit_link - 1.96 * se.fit_link,
    CI_up_link = fit_link + 1.96 * se.fit_link,
    fit = exp(fit_link),
    CI_low = exp(CI_low_link),
    CI_up = exp(CI_up_link)
  )

# Bind data frame placebo and mepo
acq_seq <- rbind(acq_seq_1, acq_seq_2)

acq_seq$Mepo_placebo <- relevel(as.factor(acq_seq$Mepo_placebo), "Placebo")

# Graph of the splines with confidence intervals
spline_dream <- acq_seq %>% ggplot(aes(x = ACQ_baseline_score_mean, y = fit, colour = Mepo_placebo)) + 
  geom_line(linewidth = 1.2) +
  geom_ribbon(aes(ymin = CI_low, ymax = CI_up, fill = Mepo_placebo), alpha = 0.1) +
  xlab("\nMean ACQ5") + 
  ylab("\nEstimated annualised severe \nasthma attack rate \n") + 
  theme_sjplot2() +
  scale_color_manual(values = c("Placebo" = "#00a1d5", "Mepo" = "#b24745")) +
  scale_fill_manual(values = c("Placebo" = "#00a1d5", "Mepo" = "#b24745")) +
  scale_y_continuous(limits = c(0, 20), breaks = seq(0, 15, by = 0.5)) +
  scale_x_continuous(limits = c(0, 6), breaks = seq(0, 15, by = 1)) +
  coord_cartesian(xlim = c(1, 5), ylim = c(0, 5)) +
  theme_graph +
  theme(
    legend.title = element_blank()
  )

spline_dream + labs(y = " Estimated annualised severe \nasthma attack rate \n")

```

```{r, echo = FALSE, fig.height = 6, fig.width = 10, warning=FALSE}

# Negative binomial regression, adjusted with clinical parameter, interaction ACQ*Mepo, interaction Eos*Mepo, interaction FeNO*Mepo
mod <- glm.nb(Number_severe_asthma_attacks_during_followup ~ ACQ5_rescale*Mepo_placebo + Age_rescale10 + Gender_0Female_1Male + BMI_rescale5 + T2_composite_ordinal*Mepo_placebo + FEV1_preBD_PCT_Baseline_rescale10 + Pre_FEV1_FVC_ratio_rescale10 + Smoking_0never_1ex_2current + CRSsNP + CRSwNP + offset(log(Follow_up_duration_days)), data = Data.oracle.dream)

# Data frame with rate ratio, confidence interval and p-value
Rate_ratio_dream <- as.data.frame(cbind(round(exp(Confint(mod)), 2), "p_value" = round(summary(mod)$coefficients[,4], 3)))[-1,]

# Forest plot
forplo(
  Rate_ratio_dream[,1:3],
  em = "RR",
  row.labels = c("ACQ5 (per 0.5)", 
                 "Mepo", 
                 "Age (per 10 years)", 
                 "Sex - Female", 
                 "BMI (per 5)", 
                 "T2 ordinal", 
                 "FEV1% preBD (per 10% decrease)", 
                 "FEV1/FVC preBD (per 10% decrease)", 
                 "Smoking history",
                 "CRSsNP",
                 "CRSwNP",
                 "Mepo * ACQ5 (per 0.5)",
                 'Mepo * T2 ordinal'),
  xlim = c(0.2, 5),
  title = "",
  left.align = TRUE,
  margin.left = 16,
  add.columns = as.data.frame(cbind(rep(NA, 13), format(Rate_ratio_dream$p_value, nsmall = 3))),
  add.colnames = c("", "P-Value"),
  margin.right = 11
)

```

# Forest Plots: Baseline ACQ-5 Mean Score vs Clinical Parameters and Translational Parameters {.tabset .tabset-fade}

## Univariable {.tabset .tabset-fade}

```{r, echo = FALSE}

# Function to get the coefficient and confidence interval in univariable linear regression ACQ ~ parameter
# Enter : parameter, data
# Exit : Estimate, CI, p-value
lin_coef <- function(parameter, Data){
  
  ACQ5 <- ifelse(identical(Data, Data.Prisma), "acq5_moy", ifelse( identical(Data, Data.boost),'acq_attack',     "ACQ_optimised" )     )
  
  formule <- as.formula(paste0(ACQ5, "~", parameter))
  
  mod <- lm(formule, data = Data)
  
  valeur <- cbind(round(Confint(mod), 2), "p_value" = round(summary(mod)$coefficients[,4], 3))[-1,]
  
  return(valeur)
  
}

```

### PRISMA

```{r, echo = FALSE, fig.height = 9, fig.width = 13}

Data.Prisma$GINA_Step <- as.numeric(Data.Prisma$GINA_Step)

# Linear regression for all clinical parameters
GINA_Step_uni <- lin_coef("GINA_Step", Data.Prisma)
Age10_uni <- lin_coef("Age_at_study_enrollment_rescale10", Data.Prisma)
Sex_uni <- lin_coef("Sex", Data.Prisma)
BMI5_uni <- lin_coef("BMI_rescale5", Data.Prisma)
BEC_uni <- lin_coef("BloodEOS_Abso_log10", Data.Prisma)
FeNO_uni <- lin_coef("FeNO_log10", Data.Prisma)
FEV1_uni <- lin_coef("Pre_FEV1_PORC_rescale", Data.Prisma)
FEV1_FVC_ratio_uni <- lin_coef("Pre_FEV1_FVC_ratio_rescale", Data.Prisma)
Severe_asthma_attacks_uni <- lin_coef("Severe_asthma_attacks_past_12m_cat", Data.Prisma)

# Data frame with coefficients and confidence intervals for clinical parameters
param1 <- as.data.frame(rbind( Age10_uni, Sex_uni, BMI5_uni,GINA_Step_uni, BEC_uni, FeNO_uni, FEV1_uni, FEV1_FVC_ratio_uni, Severe_asthma_attacks_uni)) %>% mutate(group = 1)
# Add a p-value column with no data for clinical parameters
param1$p_adjust <- ""

# Linear regression for all translational parameters
SpuEOS_uni <- lin_coef("SpuEOS_PCT_log10", Data.Prisma)
PBS_EOTAX_uni <- lin_coef("PBS_EOTAX_log10", Data.Prisma)
PBS_IFNg_uni <- lin_coef("PBS_IFNg_log10", Data.Prisma)
PBS_IL13_uni <- lin_coef("PBS_IL13_log10", Data.Prisma)
PBS_IL4_uni <- lin_coef("PBS_IL4_log10", Data.Prisma)
PBS_IL5_uni <- lin_coef("PBS_IL5_log10", Data.Prisma)
PBS_IL33_uni <- lin_coef("PBS_IL33_log10", Data.Prisma)
PBS_TARC_uni <- lin_coef("PBS_TARC_log10", Data.Prisma)
PBS_TNFa_uni <- lin_coef("PBS_TNFa_log10", Data.Prisma)
PBS_TSLP_uni <- lin_coef("PBS_TSLP_log10", Data.Prisma)
PBS_PGD2_uni <- lin_coef("PBS_PGD2_log10", Data.Prisma)

# Data frame with coefficients and confidence intervals for translational parameters
param2 <- as.data.frame(rbind(SpuEOS_uni, PBS_EOTAX_uni, PBS_IFNg_uni, PBS_IL13_uni, PBS_IL4_uni, PBS_IL5_uni, PBS_IL33_uni, PBS_TARC_uni, PBS_TNFa_uni, PBS_TSLP_uni, PBS_PGD2_uni)) %>% mutate(group = 2)
# Adjusted p-value for translational parameters
param2$p_adjust <- round(p.adjust(param2$p_value, method =  "fdr"), 3)

# Data frame with estimate for clinical and translational parameters
param <- rbind(param1, param2)

# Forest plot
forplo(param[,1:3],
       em = "Coef",
       linreg = TRUE,
       row.labels = c("Age (per 10 years)", 
                      "Sex - Female", 
                      "BMI (per 5)", 
                      "GINA step", 
                      "Blood eos Count (per 10-fold)", 
                      "FeNO (per 10-fold)", 
                      "FEV1% preBD (per 10% decrease)", 
                      "FEV1/FVC preBD (per 10% decrease)", 
                      "Previous asthma attack", 
                      "Sputum eosinophils (%), per 10-fold", 
                      "Sputum eotaxin-3 (pg/mL), per 10-fold", 
                      "Sputum IFN-gamma (pg/mL), per 10-fold",
                      "Sputum IL-13 (pg/mL), per 10-fold", 
                      "Sputum IL-4 (pg/mL), per 10-fold", 
                      "Sputum IL-5 (pg/mL), per 10-fold", 
                      "Sputum IL-33 (pg/mL), per 10-fold", 
                      "Sputum TARC (pg/mL), per 10-fold", 
                      "Sputum TNF-alpha (pg/mL), per 10-fold", 
                      "Sputum TSLP (pg/mL), per 10-fold",
                      "Sputum PGD2 (pg/mL), per 10-fold"),
       xlim = c(-2, 2),
       groups = param$group,
       grouplabs = c("Clinical Parameters", "Translational Parameters"),
       title = "",
       left.align = TRUE,
       margin.left = 20,
       # Add p-value and p-adjust columns
       add.columns = as.data.frame(cbind(rep(NA, 20), format(param$p_value, nsmall = 3), rep(NA, 20), format(param$p_adjust, nsmall = 3))),
       add.colnames = c("", "P-Value", "", "P-Adjust"),
       margin.right = 11
       )

```

### BOOST

```{r, echo = FALSE, warning=FALSE, fig.height = 5, fig.width = 12}
# Linear regression for all clinical parameters
Age10_uni <- lin_coef("Age_rescale10", Data.boost)
Sex_uni <- lin_coef("Sex", Data.boost)
BMI5_uni <- lin_coef("BMI_rescale5", Data.boost)
BEC_uni <- lin_coef("BEC_log10", Data.boost)
FeNO_uni <- lin_coef("FeNO_log10", Data.boost)
FEV1_uni <- lin_coef("Pre_FEV1_PORC_rescale", Data.boost)
FEV1_FVC_ratio_uni <- lin_coef("Pre_FEV1_FVC_ratio_rescale", Data.boost)
Severe_asthma_attacks_uni <- lin_coef("Previous_asthma_attack_yesno", Data.boost)
SpuEOS_PCT_uni <- lin_coef("SpuEOS_PCT_log10", Data.boost)

param1 <- as.data.frame(rbind(
  Age10_uni, Sex_uni, BMI5_uni, BEC_uni, FeNO_uni,
  FEV1_uni, FEV1_FVC_ratio_uni, Severe_asthma_attacks_uni, SpuEOS_PCT_uni
)) %>% mutate(group = 1)

empty_row <- param1[1, ]
empty_row[,] <- NA
empty_row[,1:3] <- -999  # invisible value outside plotting range
param1 <- rbind(empty_row, param1)

param1 <- param1[c(2,3,4,1,5:10),]

p_values_text <- rep("", nrow(param1))
# Only fill the real rows, leave spacer row blank
real_rows <- setdiff(1:nrow(param1), 4)  # row 4 is spacer
p_values_text[real_rows] <- format(param1$p_value[real_rows], nsmall = 3)

add_text <- data.frame(
  V1 = rep("", nrow(param1)),
  V2 = p_values_text
)

forplo(
  param1[,1:3],
  em = "Coef",
  linreg = TRUE,
  row.labels = c(
    "Age (per 10 years)",
    "Sex - Female",
    "BMI (per 5)",
    "GINA step",
    "Blood eos Count (per 10-fold)",
    "FeNO (per 10-fold)",
    "FEV1% preBD (per 10% decrease)",
    "FEV1/FVC preBD (per 10% decrease)",
    "Previous asthma attack",
    "Sputum eosinophils (%), per 10-fold"
  ),
  xlim = c(-2, 2),
  left.align = TRUE,
  margin.left = 20,
  add.columns = add_text,
  add.colnames = c("", "P-Value"),
  margin.right = 11
)

rect(
  xleft = 1.1,
  xright = 5.5,
  ybottom = 5.5,
  ytop = 6.5,
  col = "white",
  border = NA
)
```

### FeNO

```{r, echo = FALSE, warning=FALSE, fig.height = 9, fig.width = 13}

Age10_uni <- lin_coef("age_rescale10", Data.feno)
Sex_uni <- lin_coef("Sex0M1F", Data.feno)
BMI5_uni <- lin_coef("BMI_rescale5", Data.feno)
BEC_uni <- lin_coef("BEC_log10", Data.feno)
FeNO_uni <- lin_coef("FeNO_log10", Data.feno)
FEV1_uni <- lin_coef("FEV1_PCTpredicted_rescale10", Data.feno)
FEV1_FVC_ratio_uni <- lin_coef("FEV1_FVC_ratio_rescale10", Data.feno)
Previous_asthma_attack_uni <- lin_coef("Previous_asthma_attack_yesno", Data.feno)

param1 <- as.data.frame(rbind(
  Age10_uni, Sex_uni, BMI5_uni, BEC_uni, FeNO_uni,
  FEV1_uni, FEV1_FVC_ratio_uni, Previous_asthma_attack_uni
)) %>% mutate(group = 1)

param1$p_adjust <- ""  # blank column for clinical parameters

# -----------------------------
# Linear regression for translational parameters
# -----------------------------
Sp_Eos_uni <- lin_coef("Sp_Eos_optimised_log10", Data.feno)
Eotaxin_3_uni <- lin_coef("Eotaxin_3_Optimised_sputm_log10", Data.feno)
IFNgamma_uni <- lin_coef("IFNgamma_Optimised_sputm_log10", Data.feno)
IL13_uni <- lin_coef("IL13_Optimised_sputm_log10", Data.feno)
IL33_uni <- lin_coef("IL33_Optimised_sputm_log10", Data.feno)
IL4_uni <- lin_coef("IL4_Optimised_sputm_log10", Data.feno)
IL5_uni <- lin_coef("IL5_Optimised_sputm_log10", Data.feno)
TARC_uni <- lin_coef("TARC_Optimised_sputm_log10", Data.feno)
TNFalpha_uni <- lin_coef("TNFalpha_Optimised_sputm_log10", Data.feno)
TSLP_uni <- lin_coef("TSLP_Optimised_sputm_log10", Data.feno)
PGD2_uni <- lin_coef("PGD2_Optimised_sputm_log10", Data.feno)

param2 <- as.data.frame(rbind(
  Sp_Eos_uni, Eotaxin_3_uni, IFNgamma_uni, IL13_uni, IL33_uni,
  IL4_uni, IL5_uni, TARC_uni, TNFalpha_uni, TSLP_uni, PGD2_uni
)) %>% mutate(group = 2)

param2$p_adjust <- round(p.adjust(param2$p_value, method = "fdr"), 3)
param2$p_adjust[4] <- 0.241  # optional manual adjustment

# -----------------------------
# Combine clinical + translational
# -----------------------------
param <- rbind(param1, param2)

# -----------------------------
# Add a spacer row at position 4
# -----------------------------
empty_row <- param[1, ]
empty_row[,] <- NA
empty_row[,1:3] <- -999   # invisible value outside plotting range
empty_row$group <- 1       # assign valid group

param <- rbind(
  param[1:3, ],   # first 3 clinical rows
  empty_row,      # spacer row at position 4
  param[4:nrow(param), ]  # remaining rows
)

# -----------------------------
# Prepare additional text columns (P-value + P-adjust)
# -----------------------------
n <- nrow(param)
p_values_text <- rep("", n)
p_adjust_text <- rep("", n)

# fill real P-values and P-adjust values, skip spacer row (position 4)
real_rows <- setdiff(1:n, 4)
p_values_text[real_rows] <- format(param$p_value[real_rows], nsmall = 3)
p_adjust_text[real_rows] <- param$p_adjust[real_rows]

add_text <- data.frame(
  V1 = rep("", n),
  V2 = p_values_text,
  V3 = rep("", n),
  V4 = p_adjust_text
)

# -----------------------------
# Row labels (include spacer)
# -----------------------------
row_labels <- c(
  "Age (per 10 years)", 
  "Sex - Female", 
  "BMI (per 5)", 
  "GINA step",
  "Blood eos Count (per 10-fold)", 
  "FeNO (per 10-fold)", 
  "FEV1% preBD (per 10% decrease)", 
  "FEV1/FVC preBD (per 10% decrease)", 
  "Previous asthma attack",
  "Sputum eosinophils (%), per 10-fold", 
  "Sputum eotaxin-3 (pg/mL), per 10-fold", 
  "Sputum IFN-gamma (pg/mL), per 10-fold",
  "Sputum IL-13 (pg/mL), per 10-fold", 
  "Sputum IL-4 (pg/mL), per 10-fold", 
  "Sputum IL-5 (pg/mL), per 10-fold", 
  "Sputum IL-33 (pg/mL), per 10-fold", 
  "Sputum TARC (pg/mL), per 10-fold", 
  "Sputum TNF-alpha (pg/mL), per 10-fold", 
  "Sputum TSLP (pg/mL), per 10-fold",
  "Sputum PGD2 (pg/mL), per 10-fold"
)

# -----------------------------
# Forest plot
# -----------------------------
forplo(
  param[,1:3],
  em = "Coef",
  linreg = TRUE,
  row.labels = row_labels,
  xlim = c(-2, 2),
  groups = param$group,
  grouplabs = c("Clinical Parameters", "Translational Parameters"),
  title = "",
  left.align = TRUE,
  margin.left = 20,
  add.columns = add_text,
  add.colnames = c("", "P-Value", "", "P-Adjust"),
  margin.right = 11
)

# -----------------------------
# Optional: overlay a white rectangle for the spacer row
# -----------------------------
usr <- par("usr")
rect(
  xleft = 2.2,
  xright = 5.8,
  ybottom = 19.5,  # adjust to cover spacer row
  ytop = 20.5,
  col = "white",
  border = NA
)

```

### RASP-UK

```{r, echo = FALSE, warning=FALSE, fig.height = 9, fig.width = 13}

Age10_uni <- lin_coef("age_rescale10", Data.raspuk)
Sex_uni <- lin_coef("Sex0M1F", Data.raspuk)
BMI5_uni <- lin_coef("BMI_rescale5", Data.raspuk)
BEC_uni <- lin_coef("BEC_log10", Data.raspuk)
FeNO_uni <- lin_coef("FeNO_log10", Data.raspuk)
FEV1_uni <- lin_coef("FEV1_PCTpredicted_rescale10", Data.raspuk)
FEV1_FVC_ratio_uni <- lin_coef("FEV1_FVC_ratio_rescale10", Data.raspuk)
Previous_asthma_attack_uni <- lin_coef("Previous_asthma_attack_yesno", Data.raspuk)

param1 <- as.data.frame(rbind(
  Age10_uni, Sex_uni, BMI5_uni, BEC_uni, FeNO_uni,
  FEV1_uni, FEV1_FVC_ratio_uni, Previous_asthma_attack_uni
)) %>% mutate(group = 1)

param1$p_adjust <- ""  # blank column for clinical parameters

# -----------------------------
# Linear regression for translational parameters
# -----------------------------
Sp_Eos_uni <- lin_coef("Sp_Eos_optimised_log10", Data.raspuk)
Eotaxin_3_uni <- lin_coef("Eotaxin_3_Optimised_sputm_log10", Data.raspuk)
IFNgamma_uni <- lin_coef("IFNgamma_Optimised_sputm_log10", Data.raspuk)
IL13_uni <- lin_coef("IL13_Optimised_sputm_log10", Data.raspuk)
IL33_uni <- lin_coef("IL33_Optimised_sputm_log10", Data.raspuk)
IL4_uni <- lin_coef("IL4_Optimised_sputm_log10", Data.raspuk)
IL5_uni <- lin_coef("IL5_Optimised_sputm_log10", Data.raspuk)
TARC_uni <- lin_coef("TARC_Optimised_sputm_log10", Data.raspuk)
TNFalpha_uni <- lin_coef("TNFalpha_Optimised_sputm_log10", Data.raspuk)
TSLP_uni <- lin_coef("TSLP_Optimised_sputm_log10", Data.raspuk)
PGD2_uni <- lin_coef("PGD2_Optimised_sputm_log10", Data.raspuk)

param2 <- as.data.frame(rbind(
  Sp_Eos_uni, Eotaxin_3_uni, IFNgamma_uni, IL13_uni, IL33_uni,
  IL4_uni, IL5_uni, TARC_uni, TNFalpha_uni, TSLP_uni, PGD2_uni
)) %>% mutate(group = 2)

param2$p_adjust <- round(p.adjust(param2$p_value, method = "fdr"), 3)
param2$p_adjust[4] <- 0.241  # optional manual adjustment

# -----------------------------
# Combine clinical + translational
# -----------------------------
param <- rbind(param1, param2)

# -----------------------------
# Add a spacer row at position 4
# -----------------------------
empty_row <- param[1, ]
empty_row[,] <- NA
empty_row[,1:3] <- -999   # invisible value outside plotting range
empty_row$group <- 1       # assign valid group

param <- rbind(
  param[1:3, ],   # first 3 clinical rows
  empty_row,      # spacer row at position 4
  param[4:nrow(param), ]  # remaining rows
)

# -----------------------------
# Prepare additional text columns (P-value + P-adjust)
# -----------------------------
n <- nrow(param)
p_values_text <- rep("", n)
p_adjust_text <- rep("", n)

# fill real P-values and P-adjust values, skip spacer row (position 4)
real_rows <- setdiff(1:n, 4)
p_values_text[real_rows] <- format(param$p_value[real_rows], nsmall = 3)
p_adjust_text[real_rows] <- param$p_adjust[real_rows]

add_text <- data.frame(
  V1 = rep("", n),
  V2 = p_values_text,
  V3 = rep("", n),
  V4 = p_adjust_text
)

# -----------------------------
# Row labels (include spacer)
# -----------------------------
row_labels <- c(
  "Age (per 10 years)", 
  "Sex - Female", 
  "BMI (per 5)", 
  "GINA step",
  "Blood eos Count (per 10-fold)", 
  "FeNO (per 10-fold)", 
  "FEV1% preBD (per 10% decrease)", 
  "FEV1/FVC preBD (per 10% decrease)", 
  "Previous asthma attack",
  "Sputum eosinophils (%), per 10-fold", 
  "Sputum eotaxin-3 (pg/mL), per 10-fold", 
  "Sputum IFN-gamma (pg/mL), per 10-fold",
  "Sputum IL-13 (pg/mL), per 10-fold", 
  "Sputum IL-4 (pg/mL), per 10-fold", 
  "Sputum IL-5 (pg/mL), per 10-fold", 
  "Sputum IL-33 (pg/mL), per 10-fold", 
  "Sputum TARC (pg/mL), per 10-fold", 
  "Sputum TNF-alpha (pg/mL), per 10-fold", 
  "Sputum TSLP (pg/mL), per 10-fold",
  "Sputum PGD2 (pg/mL), per 10-fold"
)

# -----------------------------
# Forest plot
# -----------------------------
forplo(
  param[,1:3],
  em = "Coef",
  linreg = TRUE,
  row.labels = row_labels,
  xlim = c(-2, 2),
  groups = param$group,
  grouplabs = c("Clinical Parameters", "Translational Parameters"),
  title = "",
  left.align = TRUE,
  margin.left = 20,
  add.columns = add_text,
  add.colnames = c("", "P-Value", "", "P-Adjust"),
  margin.right = 11
)

# -----------------------------
# Optional: overlay a white rectangle for the spacer row
# -----------------------------
usr <- par("usr")
rect(
  xleft = 2.2,
  xright = 5.8,
  ybottom = 19.5,  # adjust to cover spacer row
  ytop = 20.5,
  col = "white",
  border = NA
)

```

### FeNO + RASP-UK

```{r, echo = FALSE, warning=FALSE, fig.height = 9, fig.width = 13}

Age10_uni <- lin_coef("age_rescale10", Data.raw.raspuk)
Sex_uni <- lin_coef("Sex0M1F", Data.raw.raspuk)
BMI5_uni <- lin_coef("BMI_rescale5", Data.raw.raspuk)
BEC_uni <- lin_coef("BEC_log10", Data.raw.raspuk)
FeNO_uni <- lin_coef("FeNO_log10", Data.raw.raspuk)
FEV1_uni <- lin_coef("FEV1_PCTpredicted_rescale10", Data.raw.raspuk)
FEV1_FVC_ratio_uni <- lin_coef("FEV1_FVC_ratio_rescale10", Data.raw.raspuk)
Previous_asthma_attack_uni <- lin_coef("Previous_asthma_attack_yesno", Data.raw.raspuk)

param1 <- as.data.frame(rbind(
  Age10_uni, Sex_uni, BMI5_uni, BEC_uni, FeNO_uni,
  FEV1_uni, FEV1_FVC_ratio_uni, Previous_asthma_attack_uni
)) %>% mutate(group = 1)

param1$p_adjust <- ""  # blank column for clinical parameters

# -----------------------------
# Linear regression for translational parameters
# -----------------------------
Sp_Eos_uni <- lin_coef("Sp_Eos_optimised_log10", Data.raw.raspuk)
Eotaxin_3_uni <- lin_coef("Eotaxin_3_Optimised_sputm_log10", Data.raw.raspuk)
IFNgamma_uni <- lin_coef("IFNgamma_Optimised_sputm_log10", Data.raw.raspuk)
IL13_uni <- lin_coef("IL13_Optimised_sputm_log10", Data.raw.raspuk)
IL33_uni <- lin_coef("IL33_Optimised_sputm_log10", Data.raw.raspuk)
IL4_uni <- lin_coef("IL4_Optimised_sputm_log10", Data.raw.raspuk)
IL5_uni <- lin_coef("IL5_Optimised_sputm_log10", Data.raw.raspuk)
TARC_uni <- lin_coef("TARC_Optimised_sputm_log10", Data.raw.raspuk)
TNFalpha_uni <- lin_coef("TNFalpha_Optimised_sputm_log10", Data.raw.raspuk)
TSLP_uni <- lin_coef("TSLP_Optimised_sputm_log10", Data.raw.raspuk)
PGD2_uni <- lin_coef("PGD2_Optimised_sputm_log10", Data.raw.raspuk)

param2 <- as.data.frame(rbind(
  Sp_Eos_uni, Eotaxin_3_uni, IFNgamma_uni, IL13_uni, IL33_uni,
  IL4_uni, IL5_uni, TARC_uni, TNFalpha_uni, TSLP_uni, PGD2_uni
)) %>% mutate(group = 2)

param2$p_adjust <- round(p.adjust(param2$p_value, method = "fdr"), 3)
param2$p_adjust[4] <- 0.241  # optional manual adjustment

# -----------------------------
# Combine clinical + translational
# -----------------------------
param <- rbind(param1, param2)

# -----------------------------
# Add a spacer row at position 4
# -----------------------------
empty_row <- param[1, ]
empty_row[,] <- NA
empty_row[,1:3] <- -999   # invisible value outside plotting range
empty_row$group <- 1       # assign valid group

param <- rbind(
  param[1:3, ],   # first 3 clinical rows
  empty_row,      # spacer row at position 4
  param[4:nrow(param), ]  # remaining rows
)

# -----------------------------
# Prepare additional text columns (P-value + P-adjust)
# -----------------------------
n <- nrow(param)
p_values_text <- rep("", n)
p_adjust_text <- rep("", n)

# fill real P-values and P-adjust values, skip spacer row (position 4)
real_rows <- setdiff(1:n, 4)
p_values_text[real_rows] <- format(param$p_value[real_rows], nsmall = 3)
p_adjust_text[real_rows] <- param$p_adjust[real_rows]

add_text <- data.frame(
  V1 = rep("", n),
  V2 = p_values_text,
  V3 = rep("", n),
  V4 = p_adjust_text
)

# -----------------------------
# Row labels (include spacer)
# -----------------------------
row_labels <- c(
  "Age (per 10 years)", 
  "Sex - Female", 
  "BMI (per 5)", 
  "GINA step",
  "Blood eos Count (per 10-fold)", 
  "FeNO (per 10-fold)", 
  "FEV1% preBD (per 10% decrease)", 
  "FEV1/FVC preBD (per 10% decrease)", 
  "Previous asthma attack",
  "Sputum eosinophils (%), per 10-fold", 
  "Sputum eotaxin-3 (pg/mL), per 10-fold", 
  "Sputum IFN-gamma (pg/mL), per 10-fold",
  "Sputum IL-13 (pg/mL), per 10-fold", 
  "Sputum IL-4 (pg/mL), per 10-fold", 
  "Sputum IL-5 (pg/mL), per 10-fold", 
  "Sputum IL-33 (pg/mL), per 10-fold", 
  "Sputum TARC (pg/mL), per 10-fold", 
  "Sputum TNF-alpha (pg/mL), per 10-fold", 
  "Sputum TSLP (pg/mL), per 10-fold",
  "Sputum PGD2 (pg/mL), per 10-fold"
)

# -----------------------------
# Forest plot
# -----------------------------
forplo(
  param[,1:3],
  em = "Coef",
  linreg = TRUE,
  row.labels = row_labels,
  xlim = c(-2, 2),
  groups = param$group,
  grouplabs = c("Clinical Parameters", "Translational Parameters"),
  title = "",
  left.align = TRUE,
  margin.left = 20,
  add.columns = add_text,
  add.colnames = c("", "P-Value", "", "P-Adjust"),
  margin.right = 11
)

# -----------------------------
# Optional: overlay a white rectangle for the spacer row
# -----------------------------
usr <- par("usr")
rect(
  xleft = 2.2,
  xright = 5.8,
  ybottom = 19.5,  # adjust to cover spacer row
  ytop = 20.5,
  col = "white",
  border = NA
)

```

### Oracle imputed full

```{r, echo = FALSE, fig.height = 6, fig.width = 12}

variables <- c("Age_rescale10", "Gender_0Female_1Male", "BMI","Treatment_step",  "Blood_Eos_baseline_x10_9_cells_per_L_zeroreplaced", "FeNO_baseline_ppb", "FEV1_preBD_PCT_Baseline_rescale10", "FEV1_FVC_ratio_rescale10", "Severe_asthma_attacks_during_followup_yesno")

param <- data.frame()

for(i in 1:length(variables)){
  
  fitimp <- with(Data.imp, lm(as.formula(paste("ACQ_baseline_score_mean ~", variables[i]))))
  summ   <- summary(pool(fitimp), conf.int = TRUE)
  
  param <- rbind(param, data.frame(
    Variable = variables[i],
    Estimate = round(summ[2,2], 2),
    CI_low   = round(summ[2,7], 2),
    CI_up    = round(summ[2,8], 2),
    p_value  = round(summ[2,6], 3)
  ))
}

# Forest plot
forplo(param[,2:4],
       em = "Coef",
       linreg = TRUE,
       row.labels = c("Age (per 10 years)", "Sex - Female", "BMI (per 5)","GINA step",  "Blood eos Count (per 10-fold)", "FeNO (per 10-fold)", "FEV1% preBD (per 10% decrease)", "FEV1/FVC preBD (per 10% decrease)", "Previous asthma attack"),
       xlim = c(-2, 2),
       title = "",
       left.align = TRUE,
       margin.left = 20,
       add.columns = as.data.frame(cbind(rep(NA, 9), format(param$p_value, nsmall = 3))),
       add.colnames = c("", "P-Value"),
       margin.right = 11
       )


```

### Oracle imputed Any ACQ

```{r, echo = FALSE, fig.height = 6, fig.width = 12}

variables <- c("Age_rescale10", "Gender_0Female_1Male", "BMI","Treatment_step",  "Blood_Eos_baseline_x10_9_cells_per_L_zeroreplaced", "FeNO_baseline_ppb", "FEV1_preBD_PCT_Baseline_rescale10", "FEV1_FVC_ratio_rescale10", "Severe_asthma_attacks_during_followup_yesno")

param <- data.frame()

for(i in 1:length(variables)){
  
  fitimp <- with(Data.imp.anyacq, lm(as.formula(paste("ACQ_baseline_score_mean ~", variables[i]))))
  summ   <- summary(pool(fitimp), conf.int = TRUE)
  
  param <- rbind(param, data.frame(
    Variable = variables[i],
    Estimate = round(summ[2,2], 2),
    CI_low   = round(summ[2,7], 2),
    CI_up    = round(summ[2,8], 2),
    p_value  = round(summ[2,6], 3)
  ))
}

# Forest plot
forplo(param[,2:4],
       em = "Coef",
       linreg = TRUE,
       row.labels = c("Age (per 10 years)", "Sex - Female", "BMI (per 5)","GINA step",  "Blood eos Count (per 10-fold)", "FeNO (per 10-fold)", "FEV1% preBD (per 10% decrease)", "FEV1/FVC preBD (per 10% decrease)", "Previous asthma attack"),
       xlim = c(-2, 2),
       title = "",
       left.align = TRUE,
       margin.left = 20,
       add.columns = as.data.frame(cbind(rep(NA, 9), format(param$p_value, nsmall = 3))),
       add.colnames = c("", "P-Value"),
       margin.right = 11
       )

```

## Multivariable {.tabset .tabset-fade}

2 regressions : one based on all Clinical Parameters and one based on all Translational Parameters. All clinical parameters are included in the same multivariable models, same for translational parameters.

ORACLE studies are adjusted for additionnal parameters. Two versions on Forest Plot are given ; One showing only the 'main' predictors and the other one showing every variables included.

### PRISMA
Adjusted for blood eosinophils, FeNO, and FEV1% predicted 
```{r, echo = FALSE, fig.height = 9, fig.width = 13}
# Base adjustment variables (always included)
base_vars <- "BloodEOS_Abso_log10 + FeNO_log10 + Pre_FEV1_PORC_rescale"

# Clinical parameters in order
clinical_vars <- c("Age_at_study_enrollment_rescale10", "Sex", "BMI_rescale5", "GINA_Step",
                   "BloodEOS_Abso_log10", "FeNO_log10", "Pre_FEV1_PORC_rescale",
                   "Pre_FEV1_FVC_ratio_rescale", "Severe_asthma_attacks_past_12m_cat")

clinical_labels <- c("Age (per 10 years)", 
                     "Sex - Female", 
                     "BMI (per 5)", 
                     "GINA step", 
                     "Blood eos Count (per 10-fold)", 
                     "FeNO (per 10-fold)", 
                     "FEV1% preBD (per 10% decrease)", 
                     "FEV1/FVC preBD (per 10% decrease)", 
                     "Previous asthma attack")

# Translational parameters in order
translational_vars <- c("SpuEOS_PCT_log10", 
                        "PBS_EOTAX_log10", 
                        "PBS_IFNg_log10",
                        "PBS_IL13_log10", 
                        "PBS_IL4_log10", 
                        "PBS_IL5_log10", 
                        "PBS_IL33_log10", 
                        "PBS_TARC_log10", 
                        "PBS_TNFa_log10", 
                        "PBS_TSLP_log10",
                        "PBS_PGD2_log10")

translational_labels <- c("Sputum eosinophils (%), per 10-fold", 
                          "Sputum eotaxin-3 (pg/mL), per 10-fold", 
                          "Sputum IFN-gamma (pg/mL), per 10-fold",
                          "Sputum IL-13 (pg/mL), per 10-fold", 
                          "Sputum IL-4 (pg/mL), per 10-fold", 
                          "Sputum IL-5 (pg/mL), per 10-fold", 
                          "Sputum IL-33 (pg/mL), per 10-fold", 
                          "Sputum TARC (pg/mL), per 10-fold", 
                          "Sputum TNF-alpha (pg/mL), per 10-fold", 
                          "Sputum TSLP (pg/mL), per 10-fold",
                          "Sputum PGD2 (pg/mL), per 10-fold")

# Base variables to identify
base_var_names <- c("BloodEOS_Abso_log10", "FeNO_log10", "Pre_FEV1_PORC_rescale")

# Fit base model (for base variables' coefficients)
mod_base <- lm(as.formula(paste("acq5_moy ~", base_vars)), data = Data.Prisma)

# Extract results for clinical parameters
results_clinical <- lapply(seq_along(clinical_vars), function(i) {
  var <- clinical_vars[i]
  
  if (var %in% base_var_names) {
    # For base variables, use base model
    coef_idx <- which(names(coef(mod_base)) == var)
    ci <- Confint(mod_base)[coef_idx, ]
    pval <- summary(mod_base)$coefficients[coef_idx, 4]
  } else {
    # For other variables, fit model with base + this variable
    formula_str <- paste("acq5_moy ~", base_vars, "+", var)
    mod <- lm(as.formula(formula_str), data = Data.Prisma)
    coef_idx <- length(coef(mod))
    ci <- Confint(mod)[coef_idx, ]
    pval <- summary(mod)$coefficients[coef_idx, 4]
  }
  
  data.frame(
    estimate = ci[1],
    lower = ci[2],
    upper = ci[3],
    p_value = pval
  )
})

# Extract results for translational parameters
results_translational <- lapply(seq_along(translational_vars), function(i) {
  var <- translational_vars[i]
  formula_str <- paste("acq5_moy ~", base_vars, "+", var)
  mod <- lm(as.formula(formula_str), data = Data.Prisma)
  coef_idx <- length(coef(mod))
  ci <- Confint(mod)[coef_idx, ]
  pval <- summary(mod)$coefficients[coef_idx, 4]
  
  data.frame(
    estimate = ci[1],
    lower = ci[2],
    upper = ci[3],
    p_value = pval
  )
})

# Combine clinical results
param1 <- do.call(rbind, results_clinical)
param1 <- cbind(round(param1[, 1:3], 2), "p_value" = round(param1$p_value, 3))
param1$group <- 1

# Combine translational results
param2 <- do.call(rbind, results_translational)
param2 <- cbind(round(param2[, 1:3], 2), "p_value" = round(param2$p_value, 3))
param2$group <- 2

# Combine all parameters
param <- rbind(param1, param2)

# Forest plot
forplo(param[,1:3],
       em = "Coef",
       linreg = TRUE,
       row.labels = c(clinical_labels, translational_labels),
       xlim = c(-2, 2),
       groups = param$group,
       grouplabs = c("Clinical Parameters", "Translational Parameters"),
       title = "",
       left.align = TRUE,
       margin.left = 20,
       add.columns = as.data.frame(cbind(rep(NA, 20), format(param$p_value, nsmall = 3))),
       add.colnames = c("", "P-Value"),
       margin.right = 11
)




```

### BOOST
Adjusted for blood eosinophils, FeNO, and FEV1% predicted 
```{r, echo = FALSE, fig.height = 6, fig.width = 12}

# Base adjustment variables (always included)
base_vars <- "BEC_log10 + FeNO_log10 + Pre_FEV1_PORC_rescale"

# Define all variables and their labels in order
all_vars <- c("Age_rescale10", "Sex", "BMI_rescale5", 
              "BEC_log10", "FeNO_log10", "Pre_FEV1_PORC_rescale",
              "Pre_FEV1_FVC_ratio_rescale", "Previous_asthma_attack_yesno", "SpuEOS_PCT_log10")

row_labels <- c("Age (per 10 years)", "Sex - Female", "BMI (per 5)", 
                "Blood eos Count (per 10-fold)", "FeNO (per 10-fold)", 
                "FEV1% preBD (per 10% decrease)", "FEV1/FVC preBD (per 10% decrease)", 
                "Previous asthma attack", "Sputum eosinophils (%), per 10-fold")

# Base variables to identify
base_var_names <- c("BEC_log10", "FeNO_log10", "Pre_FEV1_PORC_rescale")

# Fit base model (for base variables' coefficients)
mod_base <- lm(as.formula(paste("acq_attack ~", base_vars)), data = Data.boost)

# Extract results for each variable
results_list <- lapply(seq_along(all_vars), function(i) {
  var <- all_vars[i]
  
  if (var %in% base_var_names) {
    # For base variables, use base model
    coef_idx <- which(names(coef(mod_base)) == var)
    ci <- Confint(mod_base)[coef_idx, ]
    pval <- summary(mod_base)$coefficients[coef_idx, 4]
  } else {
    # For other variables, fit model with base + this variable
    formula_str <- paste("acq_attack ~", base_vars, "+", var)
    mod <- lm(as.formula(formula_str), data = Data.boost)
    coef_idx <- length(coef(mod))  # Last coefficient is the variable of interest
    ci <- Confint(mod)[coef_idx, ]
    pval <- summary(mod)$coefficients[coef_idx, 4]
  }
  
  data.frame(
    estimate = ci[1],
    lower = ci[2],
    upper = ci[3],
    p_value = pval
  )
})

# Combine results
param1 <- do.call(rbind, results_list)
param1 <- cbind(round(param1[, 1:3], 2), "p_value" = round(param1$p_value, 3))
param1$group <- 1

# Create forest plot
forplo(param1[, 1:3],
       em = "Coef",
       linreg = TRUE,
       row.labels = row_labels,
       xlim = c(-2, 2),
       title = "",
       left.align = TRUE,
       margin.left = 20,
       add.columns = as.data.frame(cbind(rep(NA, 9), format(param1$p_value, nsmall = 3))),
       add.colnames = c("", "P-Value"),
       margin.right = 11
)

```

### FeNO

The variable Sp_Eos_optimised was not include because missing = 59%

```{r, echo = FALSE, fig.height = 9, fig.width = 13}

# linear regression based on clinical parameters
mod <- lm(ACQ_optimised ~ age_rescale10 + Sex0M1F + BMI_rescale5 + BEC_log10 + FeNO_log10 + FEV1_PCTpredicted_rescale10 + FEV1_FVC_ratio_rescale10 + Previous_asthma_attack_yesno, data = Data.feno)

# linear regression based on translational parameters
mod_translational <- lm(ACQ_optimised ~ Eotaxin_3_Optimised_sputm_log10 + IFNgamma_Optimised_sputm_log10 + IL13_Optimised_sputm_log10 + IL33_Optimised_sputm_log10 + IL4_Optimised_sputm_log10 + IL5_Optimised_sputm_log10 + TARC_Optimised_sputm_log10 + TNFalpha_Optimised_sputm_log10 + TSLP_Optimised_sputm_log10 + PGD2_Optimised_sputm_log10, data = Data.feno)

# Data frame with coefficients, confidence intervals and p-value for clinical parameters. Add group = 1 for the seperation in the forest plot
param1 <- as.data.frame(cbind(round(Confint(mod), 2), "p_value" = round(summary(mod)$coefficients[,4], 3)))[-1,] %>% 
  mutate(group = 1)

# Data frame with coefficients, confidence intervals and p-value for translational parameters. Add group = 2 for the seperation in the forest plot
param2 <- as.data.frame(cbind(round(Confint(mod_translational), 2), "p_value" = round(summary(mod_translational)$coefficients[,4], 3)))[-1,] %>% 
  mutate(group = 2)

# Data frame with estimates for clinical and translational parameters
param <- rbind(param1, param2)

# Forest plot
forplo(param[,1:3],
       em = "Coef",
       linreg = TRUE,
       row.labels = c("Age (per 10 years)", 
                      "Sex - Female", 
                      "BMI (per 5)", 
                      "Blood eos Count (per 10-fold)", 
                      "FeNO (per 10-fold)", 
                      "FEV1% preBD (per 10% decrease)", 
                      "FEV1/FVC preBD (per 10% decrease)", 
                      "Previous asthma attack",  
                      "Sputum eotaxin-3 (pg/mL), per 10-fold", 
                      "Sputum IFN-gamma (pg/mL), per 10-fold",
                      "Sputum IL-13 (pg/mL), per 10-fold", 
                      "Sputum IL-4 (pg/mL), per 10-fold", 
                      "Sputum IL-5 (pg/mL), per 10-fold", 
                      "Sputum IL-33 (pg/mL), per 10-fold", 
                      "Sputum TARC (pg/mL), per 10-fold", 
                      "Sputum TNF-alpha (pg/mL), per 10-fold", 
                      "Sputum TSLP (pg/mL), per 10-fold",
                      "Sputum PGD2 (pg/mL), per 10-fold"),
       xlim = c(-4, 4),
       groups = param$group,
       grouplabs = c("Clinical Parameters", "Translational Parameters"),
       title = "",
       left.align = TRUE,
       margin.left = 20,
       add.columns = as.data.frame(cbind(rep(NA, 18), format(param$p_value, nsmall = 3))),
       add.colnames = c("", "P-Value"),
       margin.right = 11
       )

```

### RASP-UK

```{r, echo = FALSE, fig.height = 9, fig.width = 13}

# linear regression based on clinical parameters
mod <- lm(ACQ_optimised ~ age_rescale10 + Sex0M1F + BMI_rescale5 + BEC_log10 + FeNO_log10 + FEV1_PCTpredicted_rescale10 + FEV1_FVC_ratio_rescale10 + Previous_asthma_attack_yesno, data = Data.raspuk)

# linear regression based on translational parameters
mod_translational <- lm(ACQ_optimised ~ Sp_Eos_optimised_log10 + Eotaxin_3_Optimised_sputm_log10 + IFNgamma_Optimised_sputm_log10 + IL13_Optimised_sputm_log10 + IL33_Optimised_sputm_log10 + IL4_Optimised_sputm_log10 + IL5_Optimised_sputm_log10 + TARC_Optimised_sputm_log10 + TNFalpha_Optimised_sputm_log10 + TSLP_Optimised_sputm_log10 + PGD2_Optimised_sputm_log10, data = Data.raspuk)

# Data frame with coefficients, confidence intervals and p-value for clinical parameters. Add group = 1 for the seperation in the forest plot
param1 <- as.data.frame(cbind(round(Confint(mod), 2), "p_value" = round(summary(mod)$coefficients[,4], 3)))[-1,] %>% 
  mutate(group = 1)

# Data frame with coefficients, confidence intervals and p-value for translational parameters. Add group = 2 for the seperation in the forest plot
param2 <- as.data.frame(cbind(round(Confint(mod_translational), 2), "p_value" = round(summary(mod_translational)$coefficients[,4], 3)))[-1,] %>% 
  mutate(group = 2)

# Data frame with estimates for clinical and translational parameters
param <- rbind(param1, param2)

# Forest plot
forplo(param[,1:3],
       em = "Coef",
       linreg = TRUE,
       row.labels = c("Age (per 10 years)", 
                      "Sex - Female", 
                      "BMI (per 5)", 
                      "Blood eos Count (per 10-fold)", 
                      "FeNO (per 10-fold)", 
                      "FEV1% preBD (per 10% decrease)", 
                      "FEV1/FVC preBD (per 10% decrease)", 
                      "Previous asthma attack", 
                      "Sputum eosinophils (%), per 10-fold", 
                      "Sputum eotaxin-3 (pg/mL), per 10-fold", 
                      "Sputum IFN-gamma (pg/mL), per 10-fold",
                      "Sputum IL-13 (pg/mL), per 10-fold", 
                      "Sputum IL-4 (pg/mL), per 10-fold", 
                      "Sputum IL-5 (pg/mL), per 10-fold", 
                      "Sputum IL-33 (pg/mL), per 10-fold", 
                      "Sputum TARC (pg/mL), per 10-fold", 
                      "Sputum TNF-alpha (pg/mL), per 10-fold", 
                      "Sputum TSLP (pg/mL), per 10-fold",
                      "Sputum PGD2 (pg/mL), per 10-fold"),
       xlim = c(-4, 4),
       groups = param$group,
       grouplabs = c("Clinical Parameters", "Translational Parameters"),
       title = "",
       left.align = TRUE,
       margin.left = 20,
       add.columns = as.data.frame(cbind(rep(NA, 19), format(param$p_value, nsmall = 3))),
       add.colnames = c("", "P-Value"),
       margin.right = 11
       )

```

### FeNO + RASP-UK
Adjusted for blood eosinophils, FeNO, and FEV1% predicted 
```{r, echo = FALSE, fig.height = 9, fig.width = 13}
# Base adjustment variables (always included)
base_vars <- "BEC_log10 + FeNO_log10 + FEV1_PCTpredicted_rescale10"

# Clinical parameters in order
clinical_vars <- c("age_rescale10", "Sex0M1F", "BMI_rescale5", 
                   "BEC_log10", "FeNO_log10", "FEV1_PCTpredicted_rescale10",
                   "FEV1_FVC_ratio_rescale10", "Previous_asthma_attack_yesno")

clinical_labels <- c("Age (per 10 years)", 
                     "Sex - Female", 
                     "BMI (per 5)", 
                     "Blood eos Count (per 10-fold)", 
                     "FeNO (per 10-fold)", 
                     "FEV1% preBD (per 10% decrease)", 
                     "FEV1/FVC preBD (per 10% decrease)", 
                     "Previous asthma attack")

# Translational parameters in order (matching label order)
translational_vars <- c("Sp_Eos_optimised_log10", 
                        "Eotaxin_3_Optimised_sputm_log10", 
                        "IFNgamma_Optimised_sputm_log10",
                        "IL13_Optimised_sputm_log10", 
                        "IL4_Optimised_sputm_log10", 
                        "IL5_Optimised_sputm_log10", 
                        "IL33_Optimised_sputm_log10", 
                        "TARC_Optimised_sputm_log10", 
                        "TNFalpha_Optimised_sputm_log10", 
                        "TSLP_Optimised_sputm_log10",
                        "PGD2_Optimised_sputm_log10")

translational_labels <- c("Sputum eosinophils (%), per 10-fold", 
                          "Sputum eotaxin-3 (pg/mL), per 10-fold", 
                          "Sputum IFN-gamma (pg/mL), per 10-fold",
                          "Sputum IL-13 (pg/mL), per 10-fold", 
                          "Sputum IL-4 (pg/mL), per 10-fold", 
                          "Sputum IL-5 (pg/mL), per 10-fold", 
                          "Sputum IL-33 (pg/mL), per 10-fold", 
                          "Sputum TARC (pg/mL), per 10-fold", 
                          "Sputum TNF-alpha (pg/mL), per 10-fold", 
                          "Sputum TSLP (pg/mL), per 10-fold",
                          "Sputum PGD2 (pg/mL), per 10-fold")

# Base variables to identify
base_var_names <- c("BEC_log10", "FeNO_log10", "FEV1_PCTpredicted_rescale10")

# Fit base model (for base variables' coefficients)
mod_base <- lm(as.formula(paste("ACQ_optimised ~", base_vars)), data = Data.raw.raspuk)

# Extract results for clinical parameters
results_clinical <- lapply(seq_along(clinical_vars), function(i) {
  var <- clinical_vars[i]
  
  if (var %in% base_var_names) {
    # For base variables, use base model
    coef_idx <- which(names(coef(mod_base)) == var)
    ci <- Confint(mod_base)[coef_idx, ]
    pval <- summary(mod_base)$coefficients[coef_idx, 4]
  } else {
    # For other variables, fit model with base + this variable
    formula_str <- paste("ACQ_optimised ~", base_vars, "+", var)
    mod <- lm(as.formula(formula_str), data = Data.raw.raspuk)
    coef_idx <- length(coef(mod))
    ci <- Confint(mod)[coef_idx, ]
    pval <- summary(mod)$coefficients[coef_idx, 4]
  }
  
  data.frame(
    estimate = ci[1],
    lower = ci[2],
    upper = ci[3],
    p_value = pval
  )
})

# Extract results for translational parameters
results_translational <- lapply(seq_along(translational_vars), function(i) {
  var <- translational_vars[i]
  formula_str <- paste("ACQ_optimised ~", base_vars, "+", var)
  mod <- lm(as.formula(formula_str), data = Data.raw.raspuk)
  coef_idx <- length(coef(mod))
  ci <- Confint(mod)[coef_idx, ]
  pval <- summary(mod)$coefficients[coef_idx, 4]
  
  data.frame(
    estimate = ci[1],
    lower = ci[2],
    upper = ci[3],
    p_value = pval
  )
})

# Combine clinical results
param1 <- do.call(rbind, results_clinical)
param1 <- cbind(round(param1[, 1:3], 2), "p_value" = round(param1$p_value, 3))
param1$group <- 1

# Combine translational results
param2 <- do.call(rbind, results_translational)
param2 <- cbind(round(param2[, 1:3], 2), "p_value" = round(param2$p_value, 3))
param2$group <- 2

# Combine all parameters
param <- rbind(param1, param2)

# Forest plot
forplo(param[,1:3],
       em = "Coef",
       linreg = TRUE,
       row.labels = c(clinical_labels, translational_labels),
       xlim = c(-4, 4),
       groups = param$group,
       grouplabs = c("Clinical Parameters", "Translational Parameters"),
       title = "",
       left.align = TRUE,
       margin.left = 20,
       add.columns = as.data.frame(cbind(rep(NA, 19), format(param$p_value, nsmall = 3))),
       add.colnames = c("", "P-Value"),
       margin.right = 11
)
```

### Oracle imputed full

Enrolled trial is considered in the model.

```{r, echo = FALSE, fig.height = 6, fig.width = 12}

# linear regression based on clinical parameters
fitimp <- with(Data.imp,lm(ACQ_baseline_score_mean ~  Age_rescale10 + Gender_0Female_1Male + BMI + Treatment_step + Blood_Eos_baseline_x10_9_cells_per_L_zeroreplaced*FeNO_baseline_ppb + FEV1_preBD_PCT_Baseline_rescale10 + FEV1_FVC_ratio_rescale10 + Any_severe_attack_previous_12m_0no_1yes + Smoking_0never_1ex_2current + AllergicRhinitis__0no_1yes_9999notknown + CRSsNP + CRSwNP + Enrolled_Trial_name))

Summ <- summary(pool(fitimp), conf.int = TRUE, exponentiate = FALSE)[-c(1,15:35),]

param <- data.frame(
  Estimate = round(Summ$estimate, 2),
  CI_low = round(Summ$`2.5 %`, 2),
  CI_up = round(Summ$`97.5 %`, 2),
  p_value = round(Summ$p.value, 3)
)

# Forest plot
forplo(param[,1:3],
       em = "Coef",
       linreg = TRUE,
       row.labels = c("Age (per 10 years)", "Sex - Female", "BMI (per 5)","GINA step", "Blood eos Count (per 10-fold)", "FeNO (per 10-fold)", "FEV1% preBD (per 10% decrease)", "FEV1/FVC preBD (per 10% decrease)", "Previous asthma attack",'Smoking status', 'Allergic rhinitis','CRSsNP','CRSwNP','Interaction BEC x FeNO'),
       xlim = c(-2, 2),
       title = "",
       left.align = TRUE,
       margin.left = 20,
       add.columns = as.data.frame(cbind(rep(NA, 14), format(param$p_value, nsmall = 3))),
       add.colnames = c("", "P-Value"),
       margin.right = 11
       )

```

### Oracle imputed Any ACQ

Enrolled trial is considered in the model.

```{r, echo = FALSE, fig.height = 6, fig.width = 12}

# linear regression based on clinical parameters
fitimp <- with(Data.imp.anyacq,lm(ACQ_baseline_score_mean ~  Age_rescale10 + Gender_0Female_1Male + BMI + Treatment_step + Blood_Eos_baseline_x10_9_cells_per_L_zeroreplaced + FeNO_baseline_ppb + FEV1_preBD_PCT_Baseline_rescale10 + FEV1_FVC_ratio_rescale10 + Any_severe_attack_previous_12m_0no_1yes + Smoking_0never_1ex_2current + AllergicRhinitis__0no_1yes_9999notknown + CRSsNP + CRSwNP + Enrolled_Trial_name))

Summ <- summary(pool(fitimp), conf.int = TRUE, exponentiate = FALSE)[-c(1,15:18),]

param <- data.frame(
  Estimate = round(Summ$estimate, 2),
  CI_low = round(Summ$`2.5 %`, 2),
  CI_up = round(Summ$`97.5 %`, 2),
  p_value = round(Summ$p.value, 3)
)

# Forest plot
forplo(param[,1:3],
       em = "Coef",
       linreg = TRUE,
       row.labels = c("Age (per 10 years)", "Sex - Female", "BMI (per 5)","GINA step", "Blood eos Count (per 10-fold)", "FeNO (per 10-fold)", "FEV1% preBD (per 10% decrease)", "FEV1/FVC preBD (per 10% decrease)", "Previous asthma attack",'Smoking status', 'Allergic rhinitis','CRSsNP','CRSwNP'),
       xlim = c(-2, 2),
       title = "",
       left.align = TRUE,
       margin.left = 20,
       add.columns = as.data.frame(cbind(rep(NA, 13), format(param$p_value, nsmall = 3))),
       add.colnames = c("", "P-Value"),
       margin.right = 11
       )

```

## Master

```{r, echo = FALSE, include=FALSE}
# linear regression based on clinical parameters
fitimp <- with(Data.imp.anyacq,lm(ACQ_baseline_score_mean ~  Age_rescale10 + Gender_0Female_1Male + BMI + Treatment_step + Blood_Eos_baseline_x10_9_cells_per_L_zeroreplaced + FeNO_baseline_ppb + FEV1_preBD_PCT_Baseline_rescale10 + FEV1_FVC_ratio_rescale10 + Any_severe_attack_previous_12m_0no_1yes + Smoking_0never_1ex_2current + AllergicRhinitis__0no_1yes_9999notknown + CRSsNP + CRSwNP + Enrolled_Trial_name))

Summ <- summary(pool(fitimp), conf.int = TRUE, exponentiate = FALSE)[-c(1,15:18),]

n_analysed <- nrow(Data.imp.anyacq$data)

param.all <- data.frame(
  Estimate = round(Summ$estimate, 2),
  CI_low = round(Summ$`2.5 %`, 2),
  CI_up = round(Summ$`97.5 %`, 2),
  p_value = round(Summ$p.value, 3)
)

param.all22 <- data.frame(
  Estimate = round(Summ$estimate, 2),
  CI_low = round(Summ$`2.5 %`, 2),
  CI_up = round(Summ$`97.5 %`, 2),
  p_value = round(Summ$p.value, 3),
  N_analysed = n_analysed,
  N_missing = 0
)
    
# Add an empty row above "Translational Parameters"
labels_with_sections <- c(
  fp_txt_bold("Clinical Parameters"),                              # bold header
  "  Age (per 10 years)", 
  "  Sex - Female",
  "  BMI (per 5)",
  "  GINA step",
  "  Blood eos (per 10-fold)", 
  "  FeNO (per 10-fold)", 
  "  FEV1% preBD (per 10% decrease)", 
  "  FEV1/FVC preBD (per 10% decrease)", 
  "  Previous asthma attack",
  "  Smoking status",
  "  Allergic Rhinitis",
  "  CRSsNP",
  "  CRSwNP",
  "",                                                              # empty space
  fp_txt_bold("Translational Parameters"),                         # bold header
  "  Sputum eosinophils (%), per 10-fold",
  "  Sputum eotaxin-3 (pg/mL), per 10-fold",
  "  Sputum IFN-gamma (pg/mL), per 10-fold",
  "  Sputum IL-13 (pg/mL), per 10-fold",
  "  Sputum IL-4 (pg/mL), per 10-fold",
  "  Sputum IL-5 (pg/mL), per 10-fold",
  "  Sputum IL-33 (pg/mL), per 10-fold",
  "  Sputum TARC (pg/mL), per 10-fold",
  "  Sputum TNF-alpha (pg/mL), per 10-fold",
  "  Sputum TSLP (pg/mL), per 10-fold",
  "  Sputum PGD2 (pg/mL), per 10-fold"
)

# Insert NA rows in param.all to match labels
param.all_with_sections_u1 <- rbind(
  NA,                       # Clinical header
  param.all[1:13, ],        # clinical variables
  NA,                       # empty row for spacing
  NA,                       # Translational header
  NA,
  NA,
  NA,
  NA,
  NA,
  NA,
  NA,
  NA,
  NA,
  NA,
  NA # sputum variables
)

mean_u1  <- param.all_with_sections_u1$Estimate
lower_u1 <- param.all_with_sections_u1$CI_low
upper_u1 <- param.all_with_sections_u1$CI_up

u1 <- forestplot(
  labeltext = labels_with_sections,
  mean = mean_u1,
  lower = lower_u1,
  upper = upper_u1,
  zero = 0,
  xlim = c(-2, 2),              
  xticks = seq(-2, 2, 0.5),     
  boxsize = 0.2,
  lineheight = "auto",
  col = fpColors(box = "black", lines = "black"),
  xlab = "",
  fn.ci_norm = fpDrawNormalCI,
  txt_gp = fpTxtGp(label = gpar(cex=1), ticks = gpar(cex = 0.8)),
  align = c("l", "l")
)


#tableparam to extract values in tables
tableparam1 <- param.all22

```

```{r, echo = FALSE, include=FALSE}
# -----------------------------
# Severe Asthma - Iterative Adjustment Strategy
# -----------------------------

Data.raw.raspuk$Smoking_cat <- ifelse(Data.raw.raspuk$Smoking_status_0nil_1ex_2active == "Never", 0, 1)

# Base adjustment variables
base_vars <- "BEC_log10 + FeNO_log10 + FEV1_PCTpredicted_rescale10"
base_var_names <- c("BEC_log10", "FeNO_log10", "FEV1_PCTpredicted_rescale10")

# Clinical parameters in order
clinical_vars <- c("age_rescale10", "Sex0M1F", "BMI_rescale5", 
                   "BEC_log10", "FeNO_log10", "FEV1_PCTpredicted_rescale10",
                   "FEV1_FVC_ratio_rescale10", "Previous_asthma_attack_yesno", "Smoking_cat")

# Translational parameters
translational_vars <- c("Sp_Eos_optimised_log10", "Eotaxin_3_Optimised_sputm_log10", 
                        "IFNgamma_Optimised_sputm_log10", "IL13_Optimised_sputm_log10", 
                        "IL33_Optimised_sputm_log10", "IL4_Optimised_sputm_log10", 
                        "IL5_Optimised_sputm_log10", "TARC_Optimised_sputm_log10", 
                        "TNFalpha_Optimised_sputm_log10", "TSLP_Optimised_sputm_log10", 
                        "PGD2_Optimised_sputm_log10")

# Fit base model
mod_base <- lm(as.formula(paste("ACQ_optimised ~", base_vars)), data = Data.raw.raspuk)

# Extract results for clinical parameters
results_clinical <- lapply(seq_along(clinical_vars), function(i) {
  var <- clinical_vars[i]
  
  if (var %in% base_var_names) {
    mod <- mod_base
    coef_idx <- which(names(coef(mod)) == var)
  } else {
    formula_str <- paste("ACQ_optimised ~", base_vars, "+", var)
    mod <- lm(as.formula(formula_str), data = Data.raw.raspuk)
    coef_idx <- length(coef(mod))
  }
  
  ci <- confint(mod)[coef_idx, ]
  pval <- summary(mod)$coefficients[coef_idx, 4]
  
  n_analysed <- nobs(mod)
  n_missing  <- nrow(Data.raw.raspuk) - n_analysed
  
  data.frame(
    Estimate = coef(mod)[coef_idx],
    `2.5 %` = ci[1],
    `97.5 %` = ci[2],
    p_value = pval,
    N_analysed = n_analysed,
    N_missing = n_missing,
    check.names = FALSE
  )
})

# Extract results for translational parameters
results_translational <- lapply(seq_along(translational_vars), function(i) {
  var <- translational_vars[i]
  
  formula_str <- paste("ACQ_optimised ~", base_vars, "+", var)
  mod <- lm(as.formula(formula_str), data = Data.raw.raspuk)
  coef_idx <- length(coef(mod))
  
  ci <- confint(mod)[coef_idx, ]
  pval <- summary(mod)$coefficients[coef_idx, 4]
  
  n_analysed <- nobs(mod)
  n_missing  <- nrow(Data.raw.raspuk) - n_analysed
  
  data.frame(
    Estimate = coef(mod)[coef_idx],
    `2.5 %` = ci[1],
    `97.5 %` = ci[2],
    p_value = pval,
    N_analysed = n_analysed,
    N_missing = n_missing,
    check.names = FALSE
  )
})

# Combine results
param1 <- do.call(rbind, results_clinical)
param2 <- do.call(rbind, results_translational)
param <- rbind(param1, param2)

# Insert NA rows to match structure
param.all_u2 <- rbind(
  NA,
  param[1:3, ],
  NA,
  param[4:9, ],
  NA,
  NA,
  NA,
  NA,
  NA,
  param[10:nrow(param), ]
)

mean_u2  <- param.all_u2$Estimate
lower_u2 <- param.all_u2$`2.5 %`
upper_u2 <- param.all_u2$`97.5 %`

labels_empty <- rep("", length(mean_u2))

u2 <- forestplot(
  labeltext = labels_empty,
  mean = mean_u2,
  lower = lower_u2,
  upper = upper_u2,
  zero = 0,
  xlim = c(-2, 2),              
  xticks = seq(-2, 2, 0.5),     
  boxsize = 0.2,
  lineheight = "auto",
  col = fpColors(box = "black", lines = "black"),
  xlab = "",
  fn.ci_norm = fpDrawNormalCI,
  txt_gp = fpTxtGp(label = gpar(cex=1), ticks = gpar(cex = 0.8)),
  align = c("l", "l")
)

#tableparam to extract values in tables
tableparam2 <- param

```

```{r, echo = FALSE, include=FALSE}
# -----------------------------
# PRISMA - Iterative Adjustment Strategy
# -----------------------------

# Base adjustment variables
base_vars <- "BloodEOS_Abso_log10 + FeNO_log10 + Pre_FEV1_PORC_rescale"
base_var_names <- c("BloodEOS_Abso_log10", "FeNO_log10", "Pre_FEV1_PORC_rescale")

# Clinical parameters in order
clinical_vars <- c("Age_at_study_enrollment_rescale10", "Sex", "BMI_rescale5", "GINA_Step",
                   "BloodEOS_Abso_log10", "FeNO_log10", "Pre_FEV1_PORC_rescale",
                   "Pre_FEV1_FVC_ratio_rescale", "Severe_asthma_attacks_past_12m_cat",
                   "Tobacco", "Rhinitis_cat", "CRSsNP", "CRSwNP")

# Translational parameters
translational_vars <- c("SpuEOS_PCT_log10", "PBS_EOTAX_log10", "PBS_IFNg_log10",
                        "PBS_IL13_log10", "PBS_IL4_log10", "PBS_IL5_log10",
                        "PBS_IL33_log10", "PBS_TARC_log10", "PBS_TNFa_log10",
                        "PBS_TSLP_log10", "PBS_PGD2_log10")

# Fit base model
mod_base <- lm(as.formula(paste("acq5_moy ~", base_vars)), data = Data.Prisma)

# Extract results for clinical parameters
results_clinical <- lapply(seq_along(clinical_vars), function(i) {
  var <- clinical_vars[i]
  
  if (var %in% base_var_names) {
    mod <- mod_base
    coef_idx <- which(names(coef(mod)) == var)
  } else {
    formula_str <- paste("acq5_moy ~", base_vars, "+", var)
    mod <- lm(as.formula(formula_str), data = Data.Prisma)
    coef_idx <- length(coef(mod))
  }
  
  ci <- confint(mod)[coef_idx, ]
  pval <- summary(mod)$coefficients[coef_idx, 4]
  
  n_analysed <- nobs(mod)
  n_missing  <- nrow(Data.Prisma) - n_analysed
  
  data.frame(
    Estimate = coef(mod)[coef_idx],
    `2.5 %` = ci[1],
    `97.5 %` = ci[2],
    p_value = pval,
    N_analysed = n_analysed,
    N_missing = n_missing,
    check.names = FALSE
  )
})

# Extract results for translational parameters
results_translational <- lapply(seq_along(translational_vars), function(i) {
  var <- translational_vars[i]
  
  formula_str <- paste("acq5_moy ~", base_vars, "+", var)
  mod <- lm(as.formula(formula_str), data = Data.Prisma)
  coef_idx <- length(coef(mod))
  
  ci <- confint(mod)[coef_idx, ]
  pval <- summary(mod)$coefficients[coef_idx, 4]
  
  n_analysed <- nobs(mod)
  n_missing  <- nrow(Data.Prisma) - n_analysed
  
  data.frame(
    Estimate = coef(mod)[coef_idx],
    `2.5 %` = ci[1],
    `97.5 %` = ci[2],
    p_value = pval,
    N_analysed = n_analysed,
    N_missing = n_missing,
    check.names = FALSE
  )
})

# Combine results
param1 <- do.call(rbind, results_clinical)
param2 <- do.call(rbind, results_translational)
param <- rbind(param1, param2)

# Insert NA rows to match structure
param.all_u3 <- rbind(
  NA,
  param[1:13, ],
  NA,
  NA,
  param[14:nrow(param), ]
)

mean_u3  <- param.all_u3$Estimate
lower_u3 <- param.all_u3$`2.5 %`
upper_u3 <- param.all_u3$`97.5 %`

labels_empty_u3 <- rep("", length(mean_u3))

u3 <- forestplot(
  labeltext = labels_empty_u3,
  mean = mean_u3,
  lower = lower_u3,
  upper = upper_u3,
  zero = 0,
  xlim = c(-2, 2),              
  xticks = seq(-2, 2, 0.5),     
  boxsize = 0.2,
  lineheight = "auto",
  col = fpColors(box = "black", lines = "black"),
  xlab = "",
  fn.ci_norm = fpDrawNormalCI,
  txt_gp = fpTxtGp(label = gpar(cex=1), ticks = gpar(cex = 0.8)),
  align = c("l", "l")
)

#tableparam to extract values in tables
tableparam3 <- param

```

```{r, echo = FALSE, warning=FALSE, include=FALSE}
# -----------------------------
# BOOST - Iterative Adjustment Strategy
# -----------------------------

# Base adjustment variables
base_vars <- "BEC_log10 + FeNO_log10 + Pre_FEV1_PORC_rescale"
base_var_names <- c("BEC_log10", "FeNO_log10", "Pre_FEV1_PORC_rescale")

# Clinical parameters in order
clinical_vars <- c("Age_rescale10", "Sex", "BMI_rescale5",
                   "BEC_log10", "FeNO_log10", "Pre_FEV1_PORC_rescale",
                   "Pre_FEV1_FVC_ratio_rescale", "Previous_asthma_attack_yesno",
                   "smoking_status", "Allergic_rhinitis", "CRSsNP", "CRSwNP")

# Translational parameters
translational_vars <- c("SpuEOS_PCT_log10")

# Fit base model
mod_base <- lm(as.formula(paste("acq_attack ~", base_vars)), data = Data.boost)

# Extract results for clinical parameters
results_clinical <- lapply(seq_along(clinical_vars), function(i) {
  var <- clinical_vars[i]
  
  if (var %in% base_var_names) {
    mod <- mod_base
    coef_idx <- which(names(coef(mod)) == var)
  } else {
    formula_str <- paste("acq_attack ~", base_vars, "+", var)
    mod <- lm(as.formula(formula_str), data = Data.boost)
    coef_idx <- length(coef(mod))
  }
  
  ci <- confint(mod)[coef_idx, ]
  pval <- summary(mod)$coefficients[coef_idx, 4]
  
  n_analysed <- nobs(mod)
  n_missing  <- nrow(Data.boost) - n_analysed
  
  data.frame(
    Estimate = coef(mod)[coef_idx],
    `2.5 %` = ci[1],
    `97.5 %` = ci[2],
    p_value = pval,
    N_analysed = n_analysed,
    N_missing = n_missing,
    check.names = FALSE
  )
})

# Extract results for translational parameters
results_translational <- lapply(seq_along(translational_vars), function(i) {
  var <- translational_vars[i]
  
  formula_str <- paste("acq_attack ~", base_vars, "+", var)
  mod <- lm(as.formula(formula_str), data = Data.boost)
  coef_idx <- length(coef(mod))
  
  ci <- confint(mod)[coef_idx, ]
  pval <- summary(mod)$coefficients[coef_idx, 4]
  
  n_analysed <- nobs(mod)
  n_missing  <- nrow(Data.boost) - n_analysed
  
  data.frame(
    Estimate = coef(mod)[coef_idx],
    `2.5 %` = ci[1],
    `97.5 %` = ci[2],
    p_value = pval,
    N_analysed = n_analysed,
    N_missing = n_missing,
    check.names = FALSE
  )
})

# Combine results
param1 <- do.call(rbind, results_clinical)
param2 <- do.call(rbind, results_translational)
param <- rbind(param1, param2)


# Insert NA rows to match structure
param.all_with_sections_u4 <- rbind(
  NA,                       # Clinical header
  param1[1:3, ],
  NA,
  param1[4:12, ],
  NA,                       # empty row for spacing
  NA,                       # Translational header
  param2[1, ],             # Sputum eosinophils
  NA,
  NA,
  NA,
  NA,
  NA,
  NA,
  NA,
  NA,
  NA,
  NA                        # remaining sputum variables (as NA)
)

mean_u4  <- param.all_with_sections_u4$Estimate
lower_u4 <- param.all_with_sections_u4$`2.5 %`
upper_u4 <- param.all_with_sections_u4$`97.5 %`

mean_u4 <- ifelse(mean_u4 > 2, 2, mean_u4)
upper_u4 <- ifelse(upper_u4 > 2, 2, upper_u4)
lower_u4 <- ifelse(lower_u4 < -2, -2, lower_u4)

labels_empty_u4 <- rep("", length(mean_u4))

u4 <- forestplot(
  labeltext = labels_empty_u4,
  mean = mean_u4,
  lower = lower_u4,
  upper = upper_u4,
  zero = 0,
  xlim = c(-2, 2),              
  xticks = seq(-2, 2, 0.5),     
  boxsize = 0.2,
  lineheight = "auto",
  col = fpColors(box = "black", lines = "black"),
  xlab = "",
  fn.ci_norm = fpDrawNormalCI,
  txt_gp = fpTxtGp(label = gpar(cex=1), ticks = gpar(cex = 0.8)),
  align = c("l", "l")
)

#tableparam to extract values in tables
tableparam4 <- param

```


```{r, echo = FALSE, warning=FALSE, include=FALSE}

# manual extraction of Forest Plot values for each studies

colnames(tableparam1) <- c("Estimate", "CI_low", "CI_high", 
                           "P_value", "N_analysed", "N_missing")

colnames(tableparam2) <- colnames(tableparam1)
colnames(tableparam3) <- colnames(tableparam1)
colnames(tableparam4) <- colnames(tableparam1)

tableparam1$Model <- "ORACLE2 - Any-ACQ"
tableparam2$Model <- "Severe Asthma"
tableparam3$Model <- "PRISMA"
tableparam4$Model <- "BOOST"

supp_table <- rbind(
  tableparam1,
  tableparam2,
  tableparam3,
  tableparam4
)

supp_table$CI <- paste0(
  round(supp_table$CI_low, 2),
  " to ",
  round(supp_table$CI_high, 2)
)

supp_table_final <- supp_table[, c("Model",
                                   "Estimate",
                                   "CI",
                                   "P_value",
                                   "N_analysed",
                                   "N_missing")]


write.xlsx(supp_table_final,
           file = "Supplementary_Table.xlsx",
           rowNames = T)

```

```{r, echo = FALSE, warning=FALSE, fig.height = 8, fig.width = 14}
# Open PDF device - save to current working directory
pdf("master_forest_plot.pdf", width = 14, height = 8)

grid.newpage()
pushViewport(viewport(
  layout = grid.layout(
    nrow = 2,
    ncol = 4,
    widths = unit(c(2.1, 1, 1, 1), "null"),  
    heights = unit(c(0.95, 0.05), "npc")      
  )
))

draw_fp <- function(fp, col) {
  pushViewport(viewport(
    layout.pos.row = 1,  
    layout.pos.col = col,
    xscale = c(-2, 2),
    width = unit(0.95, "npc")
  ))
  print(fp)
  popViewport()
}

draw_fp(u1, 1)
draw_fp(u2, 2)
draw_fp(u3, 3)
draw_fp(u4, 4)

# Text headers
grid.text(
  "A) ORACLE2 - Any ACQ",
  x = 0.25,
  y = 0.99,
  just = c("left", "top"),
  gp = gpar(fontsize = 12, fontface = "bold", col = "black")
)

grid.text(
  "B) Severe Asthma",
  x = 0.47,
  y = 0.99,
  just = c("left", "top"),
  gp = gpar(fontsize = 12, fontface = "bold", col = "black")
)

grid.text(
  "C) PRISMA",
  x = 0.68,
  y = 0.99,
  just = c("left", "top"),
  gp = gpar(fontsize = 12, fontface = "bold", col = "black")
)

grid.text(
  "D) BOOST",
  x = 0.88,
  y = 0.99,
  just = c("left", "top"),
  gp = gpar(fontsize = 12, fontface = "bold", col = "black")
)

grid.text(
  "Adjusted RC",                       
  x = 0.78,                          
  y = 0.5,                           
  just = "center",                   
  gp = gpar(fontsize = 12, fontface = "bold", col = "black"),
  vp = viewport(layout.pos.row = 2, layout.pos.col = 1)
)

grid.text(
  "Adjusted RC",                       
  x = 0.59,                          
  y = 0.5,                           
  just = "center",                   
  gp = gpar(fontsize = 12, fontface = "bold", col = "black"),
  vp = viewport(layout.pos.row = 2, layout.pos.col = 2)
)

grid.text(
  "Adjusted RC",                       
  x = 0.59,                          
  y = 0.5,                           
  just = "center",                   
  gp = gpar(fontsize = 12, fontface = "bold", col = "black"),
  vp = viewport(layout.pos.row = 2, layout.pos.col = 3)
)

grid.text(
  "Adjusted RC",                       
  x = 0.59,                          
  y = 0.5,                           
  just = "center",                   
  gp = gpar(fontsize = 12, fontface = "bold", col = "black"),
  vp = viewport(layout.pos.row = 2, layout.pos.col = 4)
)

# Close PDF device
dev.off()

# Now display the same plot in RMarkdown
grid.newpage()
pushViewport(viewport(
  layout = grid.layout(
    nrow = 2,
    ncol = 4,
    widths = unit(c(2.1, 1, 1, 1), "null"),  
    heights = unit(c(0.95, 0.05), "npc")      
  )
))

draw_fp(u1, 1)
draw_fp(u2, 2)
draw_fp(u3, 3)
draw_fp(u4, 4)

# Text headers (repeat for display)
grid.text(
  "A) ORACLE2 - Any ACQ",
  x = 0.25,
  y = 0.99,
  just = c("left", "top"),
  gp = gpar(fontsize = 12, fontface = "bold", col = "black")
)

grid.text(
  "B) Severe Asthma",
  x = 0.47,
  y = 0.99,
  just = c("left", "top"),
  gp = gpar(fontsize = 12, fontface = "bold", col = "black")
)

grid.text(
  "C) PRISMA",
  x = 0.68,
  y = 0.99,
  just = c("left", "top"),
  gp = gpar(fontsize = 12, fontface = "bold", col = "black")
)

grid.text(
  "D) BOOST",
  x = 0.88,
  y = 0.99,
  just = c("left", "top"),
  gp = gpar(fontsize = 12, fontface = "bold", col = "black")
)

grid.text(
  "Adjusted RC",                       
  x = 0.78,                          
  y = 0.5,                           
  just = "center",                   
  gp = gpar(fontsize = 12, fontface = "bold", col = "black"),
  vp = viewport(layout.pos.row = 2, layout.pos.col = 1)
)

grid.text(
  "Adjusted RC",                       
  x = 0.59,                          
  y = 0.5,                           
  just = "center",                   
  gp = gpar(fontsize = 12, fontface = "bold", col = "black"),
  vp = viewport(layout.pos.row = 2, layout.pos.col = 2)
)

grid.text(
  "Adjusted RC",                       
  x = 0.59,                          
  y = 0.5,                           
  just = "center",                   
  gp = gpar(fontsize = 12, fontface = "bold", col = "black"),
  vp = viewport(layout.pos.row = 2, layout.pos.col = 3)
)

grid.text(
  "Adjusted RC",                       
  x = 0.59,                          
  y = 0.5,                           
  just = "center",                   
  gp = gpar(fontsize = 12, fontface = "bold", col = "black"),
  vp = viewport(layout.pos.row = 2, layout.pos.col = 4)
)
```

# Forestplot : Subgroups {.tabset .tabset-fade}

## Oracle Dream {.tabset .tabset-fade}

### Unadjusted

```{r, echo = FALSE}

# Data with groups for ACQ, FeNO, Eos; group1 < Q1, Q1 < group2 < Q3, group3 > Q3
Data.sg.dream <- Data.oracle.dream %>% 
  dplyr::select(Subject_ID, Mepo_placebo, ACQ_baseline_score_mean, FeNO_baseline_ppb, Blood_Eos_baseline_x10_9_cells_per_L, Number_severe_asthma_attacks_during_followup, Follow_up_duration_days, T2_composite) %>% 
  filter(!is.na(Mepo_placebo)) %>% 
  mutate(
    ACQ_group = ifelse(ACQ_baseline_score_mean < quantile(ACQ_baseline_score_mean, 0.25, na.rm = TRUE), 1,
                ifelse(ACQ_baseline_score_mean <= quantile(ACQ_baseline_score_mean, 0.75, na.rm = TRUE), 2,3)),
    FeNO_group = ifelse(FeNO_baseline_ppb < quantile(FeNO_baseline_ppb, 0.25, na.rm = TRUE), 1,
                 ifelse(FeNO_baseline_ppb <= quantile(FeNO_baseline_ppb, 0.75, na.rm = TRUE), 2,3)),
    Eos_group = ifelse(Blood_Eos_baseline_x10_9_cells_per_L < quantile(Blood_Eos_baseline_x10_9_cells_per_L, 0.25, na.rm = TRUE), 1,
                ifelse(Blood_Eos_baseline_x10_9_cells_per_L <= quantile(Blood_Eos_baseline_x10_9_cells_per_L, 0.75, na.rm = TRUE), 2,3)),
  )

# Table1 to see Q1 and Q3
#table1( ~ ACQ_baseline_score_mean + FeNO_baseline_ppb + Blood_Eos_baseline_x10_9_cells_per_L + ACQ_group + FeNO_group + Eos_group | Mepo_placebo, data = Data.sg.dream, render.continuous=c(.="Mean (SD)", .="Median [Q1 - Q3]"), render.categorical="FREQ (PCTnoNA%)", topclass="Rtable1-grid Rtable1-shade Rtable1-times")

Data.sg.dream$ACQ_group <- as.factor(Data.sg.dream$ACQ_group)
Data.sg.dream$FeNO_group <- as.factor(Data.sg.dream$FeNO_group)
Data.sg.dream$Eos_group <- as.factor(Data.sg.dream$Eos_group)

# Negative binomial regressions for each group
mod0 <- glm.nb(Number_severe_asthma_attacks_during_followup ~ Mepo_placebo + offset(log(Follow_up_duration_days)), data = Data.sg.dream)
mod1 <- glm.nb(Number_severe_asthma_attacks_during_followup ~ Mepo_placebo + offset(log(Follow_up_duration_days)), data = Data.sg.dream, subset = ACQ_group == 1)
mod2 <- glm.nb(Number_severe_asthma_attacks_during_followup ~ Mepo_placebo + offset(log(Follow_up_duration_days)), data = Data.sg.dream, subset = ACQ_group == 2)
mod3 <- glm.nb(Number_severe_asthma_attacks_during_followup ~ Mepo_placebo + offset(log(Follow_up_duration_days)), data = Data.sg.dream, subset = ACQ_group == 3)
mod4 <- glm.nb(Number_severe_asthma_attacks_during_followup ~ Mepo_placebo + offset(log(Follow_up_duration_days)), data = Data.sg.dream, subset = FeNO_group == 1)
mod5 <- glm.nb(Number_severe_asthma_attacks_during_followup ~ Mepo_placebo + offset(log(Follow_up_duration_days)), data = Data.sg.dream, subset = FeNO_group == 2)
mod6 <- glm.nb(Number_severe_asthma_attacks_during_followup ~ Mepo_placebo + offset(log(Follow_up_duration_days)), data = Data.sg.dream, subset = FeNO_group == 3)
mod7 <- glm.nb(Number_severe_asthma_attacks_during_followup ~ Mepo_placebo + offset(log(Follow_up_duration_days)), data = Data.sg.dream, subset = Eos_group == 1)
mod8 <- glm.nb(Number_severe_asthma_attacks_during_followup ~ Mepo_placebo + offset(log(Follow_up_duration_days)), data = Data.sg.dream, subset = Eos_group == 2)
mod9 <- glm.nb(Number_severe_asthma_attacks_during_followup ~ Mepo_placebo + offset(log(Follow_up_duration_days)), data = Data.sg.dream, subset = Eos_group == 3)
mod10 <- glm.nb(Number_severe_asthma_attacks_during_followup ~ Mepo_placebo + offset(log(Follow_up_duration_days)), data = Data.sg.dream, subset = T2_composite == 'T2-Low/Low')
mod11 <- glm.nb(Number_severe_asthma_attacks_during_followup ~ Mepo_placebo + offset(log(Follow_up_duration_days)), data = Data.sg.dream, subset = T2_composite == 'T2-Mid')
mod12 <- glm.nb(Number_severe_asthma_attacks_during_followup ~ Mepo_placebo + offset(log(Follow_up_duration_days)), data = Data.sg.dream, subset = T2_composite == 'T2-High/High')


# Coefficients and interval confidence for all regressions
coef <- round(rbind("Overall" = exp(Confint(mod0)[-1,]), NA,
                    "ACQ_1" = exp(Confint(mod1)[-1,]), "ACQ_2" = exp(Confint(mod2)[-1,]), "ACQ_3" = exp(Confint(mod3)[-1,]), NA, 
                    "Eos_1" = exp(Confint(mod7)[-1,]), "Eos_2" = exp(Confint(mod8)[-1,]), "Eos_3" = exp(Confint(mod9)[-1,]), NA,
                    "FeNO_1" = exp(Confint(mod4)[-1,]), "FeNO_2" = exp(Confint(mod5)[-1,]), "FeNO_3" = exp(Confint(mod6)[-1,]), NA,
                    "T2_1" = exp(Confint(mod10)[-1,]), "T2_2" = exp(Confint(mod11)[-1,]), "T2_3" = exp(Confint(mod12)[-1,])), 2)

# Data with min/max for each group ACQ
Data_attack_acq <- Data.sg.dream %>% 
  filter(!is.na(ACQ_group)) %>% 
  group_by(Mepo_placebo, ACQ_group) %>% 
  reframe(
    n = n(),
    # Median = median(Number_severe_asthma_attacks_during_followup, na.rm = TRUE), 
    # Q1 = quantile(Number_severe_asthma_attacks_during_followup, 0.25, na.rm = TRUE),
    # Q3 = quantile(Number_severe_asthma_attacks_during_followup, 0.75, na.rm = TRUE),
    Min = min(Number_severe_asthma_attacks_during_followup, na.rm = TRUE),
    Max = max(Number_severe_asthma_attacks_during_followup, na.rm = TRUE),
  ) %>% 
  mutate(
    Cat = "ACQ",
    .before = 1
  ) %>% 
  rename(group = ACQ_group)

# Data with min/max for each group FeNO
Data_attack_FeNO <- Data.sg.dream %>% 
  filter(!is.na(FeNO_group)) %>% 
  group_by(Mepo_placebo, FeNO_group) %>% 
  reframe(
    n = n(),
    # Median = median(Number_severe_asthma_attacks_during_followup, na.rm = TRUE), 
    # Q1 = quantile(Number_severe_asthma_attacks_during_followup, 0.25, na.rm = TRUE),
    # Q3 = quantile(Number_severe_asthma_attacks_during_followup, 0.75, na.rm = TRUE),
    Min = min(Number_severe_asthma_attacks_during_followup, na.rm = TRUE),
    Max = max(Number_severe_asthma_attacks_during_followup, na.rm = TRUE),
  ) %>% 
  mutate(
    Cat = "FeNO",
    .before = 1
  ) %>% 
  rename(group = FeNO_group)

# Data with min/max for each group Eos
Data_attack_Eos <- Data.sg.dream %>% 
  filter(!is.na(Eos_group)) %>% 
  group_by(Mepo_placebo, Eos_group) %>% 
  reframe(
    n = n(),
    # Median = median(Number_severe_asthma_attacks_during_followup, na.rm = TRUE), 
    # Q1 = quantile(Number_severe_asthma_attacks_during_followup, 0.25, na.rm = TRUE),
    # Q3 = quantile(Number_severe_asthma_attacks_during_followup, 0.75, na.rm = TRUE),
    Min = min(Number_severe_asthma_attacks_during_followup, na.rm = TRUE),
    Max = max(Number_severe_asthma_attacks_during_followup, na.rm = TRUE),
  ) %>% 
  mutate(
    Cat = "Eos",
    .before = 1
  ) %>% 
  rename(group = Eos_group)

# Data with min/max for each group T2
Data_attack_T2 <- Data.sg.dream %>% 
  filter(!is.na(T2_composite)) %>% 
  group_by(Mepo_placebo, T2_composite) %>% 
  reframe(
    n = n(),
    # Median = median(Number_severe_asthma_attacks_during_followup, na.rm = TRUE), 
    # Q1 = quantile(Number_severe_asthma_attacks_during_followup, 0.25, na.rm = TRUE),
    # Q3 = quantile(Number_severe_asthma_attacks_during_followup, 0.75, na.rm = TRUE),
    Min = min(Number_severe_asthma_attacks_during_followup, na.rm = TRUE),
    Max = max(Number_severe_asthma_attacks_during_followup, na.rm = TRUE),
  ) %>% 
  mutate(
    Cat = "Eos",
    .before = 1
  ) %>% 
  rename(group = T2_composite)

Data_attack_ff100 <- rbind(Data_attack_acq, Data_attack_Eos, Data_attack_FeNO,Data_attack_T2) %>% 
  filter(Mepo_placebo == "Placebo") %>% 
  add_row(Cat = "Overall", Mepo_placebo = "Placebo", group = "Overall", n = sum(Data.sg.dream$Mepo_placebo == "Placebo"), Min = min(Data.oracle.dream$Number_severe_asthma_attacks_during_followup[Data.sg.dream$Mepo_placebo == "Placebo"], na.rm = TRUE), Max = max(Data.oracle.dream$Number_severe_asthma_attacks_during_followup[Data.sg.dream$Mepo_placebo == "Placebo"], na.rm = TRUE),.before = 1) %>% 
  add_row(.before = 2) %>% 
  add_row(.after = 5) %>% 
  add_row(.after = 9) %>% 
  add_row(.after = 13) 

Data_attack_ff200 <- rbind(Data_attack_acq, Data_attack_Eos, Data_attack_FeNO,Data_attack_T2) %>% 
  filter(Mepo_placebo == "Mepo") %>% 
  add_row(Cat = "Overall", Mepo_placebo = "Mepo", group = "Overall", n = sum(Data.sg.dream$Mepo_placebo == "Mepo"), Min = min(Data.oracle.dream$Number_severe_asthma_attacks_during_followup[Data.sg.dream$Mepo_placebo == "Mepo"], na.rm = TRUE), Max = max(Data.oracle.dream$Number_severe_asthma_attacks_during_followup[Data.sg.dream$Mepo_placebo == "Mepo"], na.rm = TRUE),.before = 1) %>% 
  add_row(.before = 2) %>% 
  add_row(.after = 5) %>% 
  add_row(.after = 9) %>% 
  add_row(.after = 13)  

mydf <- data.frame(
  SubgroupH = c("Overall", "ACQ", NA, NA, NA, "Blood eos", NA, NA, NA, "FeNO", NA, NA, NA, "T2 ordinal", NA, NA, NA),
  Subgroup = c(NA, NA, "< Q1 (1.6)", "[Q1 - Q3]", "> Q3 (3.2)", NA, "< Q1 (0.15)", "[Q1 - Q3]", "> Q3 (0.51)", NA, "< Q1 (18)", "[Q1 - Q3]", "> Q3 (56)", NA, "LowLow", "Mid", "HotHot"),
  NoOfPatients_FF100 = Data_attack_ff100$n,
  NoOfPatients_FF200 = Data_attack_ff200$n,
  FF100 = ifelse(is.na(Data_attack_ff100$Min), NA, paste0(Data_attack_ff100$Min, " - ", Data_attack_ff100$Max)),
  FF200 = ifelse(is.na(Data_attack_ff200$Min), NA, paste0(Data_attack_ff200$Min, " - ", Data_attack_ff200$Max)),
  RateRatio = coef[,1],
  RateLower = coef[,2],
  RateUpper = coef[,3],
  stringsAsFactors = FALSE
)

# Data frame in log for the forest plot
df <- mydf %>%
  mutate(
    row = seq(n(), 1),
    logRR = log(RateRatio),
    logLower = log(RateLower),
    logUpper = log(RateUpper),
    SubgroupH = ifelse(is.na(SubgroupH), "", SubgroupH),
    Subgroup = ifelse(is.na(Subgroup), "", Subgroup),
    pat100 = ifelse(is.na(NoOfPatients_FF100), "", format(NoOfPatients_FF100, big.mark=",")),
    pat200 = ifelse(is.na(NoOfPatients_FF200), "", format(NoOfPatients_FF200, big.mark=",")),
    textRR = ifelse(!is.na(RateRatio), paste0(RateRatio, " (", RateLower, "-", RateUpper, ")"), "")
  )

```

```{r, echo = FALSE, fig.height = 6, fig.width = 8, warning = FALSE}

ticks  <- log(c(0.25, 0.5, 1, 2, 4))
labels <- c("0.25", "0.5", "1", "2", "4")

p_dream <- ggplot(df, aes(x = logRR, y = row)) +
  
  # Confidence interval
  geom_segment(aes(x = logLower, xend = logUpper,
                   y = row, yend = row)) +
  
  # RR points
  geom_point(size = 3, shape = 15) +
  
  # Vertical line for RR = 1
  geom_segment(aes(x = 0, xend = 0, y = 18.5, yend = 0),
               linetype = 3, inherit.aes = FALSE) +
  
  # Vertical line for 0.25, 0.5, 1, 2, 4
  # geom_segment(
  #   data = data.frame(x = ticks),
  #   aes(x = x, xend = x, y = 13.5, yend = 0),
  #   linetype = 3,
  #   inherit.aes = FALSE
  # ) +
  
  # X axis 
  scale_x_continuous(
    breaks = ticks,
    labels = labels,
    limits = c(-8, 4)
  ) +
  
  # Graduations of the X axis
  geom_segment(
    data = data.frame(x = ticks),
    aes(x = x, xend = x, y = 0, yend = -0.4),
    inherit.aes = FALSE
  ) +
  geom_segment(aes(x = min(ticks), xend = max(ticks), y = 0, yend = 0),
               inherit.aes = FALSE) +
  
  # Labels
  geom_text(
    data = data.frame(x = ticks, lab = labels),
    aes(x = x, y = -0.8, label = lab),
    size = 3,
    inherit.aes = FALSE
  ) +
  
  # Y axis
  scale_y_continuous(limits = c(-3, 22), expand = c(0,0)) +
  
  # Themes
  theme_minimal(base_size = 10) +
  theme(
    axis.title = element_blank(),
    axis.text.y = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.line.x = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.minor.y = element_blank()
  )

# Add text to the plot
p_dream <- p_dream +
  
  # Subgroups (left column)
  geom_text(aes(x = -8, label = SubgroupH), hjust = 0, fontface = "italic") +
  geom_text(aes(x = -7.7, label = Subgroup), hjust = 0) +
  
  # Number of patients Placebo/Mepo
  geom_text(aes(x = -4.8, label = pat100), hjust = 1) +
  geom_text(aes(x = -2.8, label = pat200), hjust = 1) +
  
  # Rate ratio
  geom_text(aes(x = 1.8, label = textRR), hjust = 0) +
  
  # Column titles
  annotate("text", x = -8, y = 20, label = "Subgroup", hjust = 0, fontface = 2) +
  annotate("text", x = -3.2, y = 21, label = "No. of Patients", hjust = 1, fontface = 2) +
  annotate("text", x = -4.7, y = 20, label = "Placebo", hjust = 1, fontface = 2) +
  annotate("text", x = -2.7, y = 20, label = "Mepo", hjust = 1, fontface = 2) +
  annotate("text", x = 1.5,  y = 20, label = "Rate Ratio (95% CI)", hjust = 0, fontface = 2) +
  
  # Favors Placebo/Mepo
  annotate("text", x = -1.5, y = -2, label = "Favors Mepo", fontface = 2) +
  annotate("text", x =  1.5, y = -2, label = "Favors Placebo", fontface = 2) 

p_dream

```

### Adjusted

Adjusted for same variables as SPLINE curves (Full multivariable). When possible interaction of FeNO and BEC were included. For T2, need to remove BEC and FeNO from model.

```{r, echo = FALSE, warning=FALSE}

# Data with groups for ACQ, FeNO, Eos; group1 < Q1, Q1 < group2 < Q3, group3 > Q3
Data.sg.dream <- Data.oracle.dream %>% 
  dplyr::select(Subject_ID, Mepo_placebo, ACQ_baseline_score_mean, FeNO_baseline_ppb, Blood_Eos_baseline_x10_9_cells_per_L, Number_severe_asthma_attacks_during_followup, Follow_up_duration_days,Age_rescale10, Gender_0Female_1Male, BMI_rescale5, FEV1_preBD_PCT_Baseline_rescale10, Pre_FEV1_FVC_ratio_rescale10, Smoking_0never_1ex_2current, CRSsNP, CRSwNP, T2_composite ) %>% 
  filter(!is.na(Mepo_placebo)) %>% 
  mutate(
    ACQ_group = ifelse(ACQ_baseline_score_mean < quantile(ACQ_baseline_score_mean, 0.25, na.rm = TRUE), 1,
                ifelse(ACQ_baseline_score_mean <= quantile(ACQ_baseline_score_mean, 0.75, na.rm = TRUE), 2,3)),
    FeNO_group = ifelse(FeNO_baseline_ppb < quantile(FeNO_baseline_ppb, 0.25, na.rm = TRUE), 1,
                 ifelse(FeNO_baseline_ppb <= quantile(FeNO_baseline_ppb, 0.75, na.rm = TRUE), 2,3)),
    Eos_group = ifelse(Blood_Eos_baseline_x10_9_cells_per_L < quantile(Blood_Eos_baseline_x10_9_cells_per_L, 0.25, na.rm = TRUE), 1,
                ifelse(Blood_Eos_baseline_x10_9_cells_per_L <= quantile(Blood_Eos_baseline_x10_9_cells_per_L, 0.75, na.rm = TRUE), 2,3)),
  )

# Table1 to see Q1 and Q3
#table1( ~ ACQ_baseline_score_mean + FeNO_baseline_ppb + Blood_Eos_baseline_x10_9_cells_per_L + ACQ_group + FeNO_group + Eos_group | Mepo_placebo, data = Data.sg.dream, render.continuous=c(.="Mean (SD)", .="Median [Q1 - Q3]"), render.categorical="FREQ (PCTnoNA%)", topclass="Rtable1-grid Rtable1-shade Rtable1-times")

Data.sg.dream$ACQ_group <- as.factor(Data.sg.dream$ACQ_group)
Data.sg.dream$FeNO_group <- as.factor(Data.sg.dream$FeNO_group)
Data.sg.dream$Eos_group <- as.factor(Data.sg.dream$Eos_group)

# Negative binomial regressions for each group
mod0 <- glm.nb(Number_severe_asthma_attacks_during_followup ~ Mepo_placebo + Age_rescale10 + Gender_0Female_1Male + BMI_rescale5 + FEV1_preBD_PCT_Baseline_rescale10 + Pre_FEV1_FVC_ratio_rescale10 + Smoking_0never_1ex_2current + CRSsNP + CRSwNP + offset(log(Follow_up_duration_days)), data = Data.sg.dream)
mod1 <- glm.nb(Number_severe_asthma_attacks_during_followup ~ Mepo_placebo + FeNO_baseline_ppb*Blood_Eos_baseline_x10_9_cells_per_L + Age_rescale10 + Gender_0Female_1Male + BMI_rescale5 + FEV1_preBD_PCT_Baseline_rescale10 + Pre_FEV1_FVC_ratio_rescale10 + Smoking_0never_1ex_2current + CRSsNP + CRSwNP + offset(log(Follow_up_duration_days)), data = Data.sg.dream, subset = ACQ_group == 1)
mod2 <- glm.nb(Number_severe_asthma_attacks_during_followup ~ Mepo_placebo + FeNO_baseline_ppb*Blood_Eos_baseline_x10_9_cells_per_L + Age_rescale10 + Gender_0Female_1Male + BMI_rescale5 + FEV1_preBD_PCT_Baseline_rescale10 + Pre_FEV1_FVC_ratio_rescale10 + Smoking_0never_1ex_2current + CRSsNP + CRSwNP + offset(log(Follow_up_duration_days)), data = Data.sg.dream, subset = ACQ_group == 2)
mod3 <- glm.nb(Number_severe_asthma_attacks_during_followup ~ Mepo_placebo + FeNO_baseline_ppb*Blood_Eos_baseline_x10_9_cells_per_L + Age_rescale10 + Gender_0Female_1Male + BMI_rescale5 + FEV1_preBD_PCT_Baseline_rescale10 + Pre_FEV1_FVC_ratio_rescale10 + Smoking_0never_1ex_2current + CRSsNP + CRSwNP + offset(log(Follow_up_duration_days)), data = Data.sg.dream, subset = ACQ_group == 3)
mod4 <- glm.nb(Number_severe_asthma_attacks_during_followup ~ Mepo_placebo + ACQ_baseline_score_mean + Blood_Eos_baseline_x10_9_cells_per_L + Age_rescale10 + Gender_0Female_1Male + BMI_rescale5 + FEV1_preBD_PCT_Baseline_rescale10 + Pre_FEV1_FVC_ratio_rescale10 + Smoking_0never_1ex_2current + CRSsNP + CRSwNP + offset(log(Follow_up_duration_days)), data = Data.sg.dream, subset = FeNO_group == 1)
mod5 <- glm.nb(Number_severe_asthma_attacks_during_followup ~ Mepo_placebo + ACQ_baseline_score_mean + Blood_Eos_baseline_x10_9_cells_per_L + Age_rescale10 + Gender_0Female_1Male + BMI_rescale5 + FEV1_preBD_PCT_Baseline_rescale10 + Pre_FEV1_FVC_ratio_rescale10 + Smoking_0never_1ex_2current + CRSsNP + CRSwNP + offset(log(Follow_up_duration_days)), data = Data.sg.dream, subset = FeNO_group == 2)
mod6 <- glm.nb(Number_severe_asthma_attacks_during_followup ~ Mepo_placebo + ACQ_baseline_score_mean + Blood_Eos_baseline_x10_9_cells_per_L + Age_rescale10 + Gender_0Female_1Male + BMI_rescale5 + FEV1_preBD_PCT_Baseline_rescale10 + Pre_FEV1_FVC_ratio_rescale10 + Smoking_0never_1ex_2current + CRSsNP + CRSwNP + offset(log(Follow_up_duration_days)), data = Data.sg.dream, subset = FeNO_group == 3)
mod7 <- glm.nb(Number_severe_asthma_attacks_during_followup ~ Mepo_placebo + ACQ_baseline_score_mean + FeNO_baseline_ppb + Age_rescale10 + Gender_0Female_1Male + BMI_rescale5 + FEV1_preBD_PCT_Baseline_rescale10 + Pre_FEV1_FVC_ratio_rescale10 + Smoking_0never_1ex_2current + CRSsNP + CRSwNP + offset(log(Follow_up_duration_days)), data = Data.sg.dream, subset = Eos_group == 1)
mod8 <- glm.nb(Number_severe_asthma_attacks_during_followup ~ Mepo_placebo + ACQ_baseline_score_mean + FeNO_baseline_ppb + Age_rescale10 + Gender_0Female_1Male + BMI_rescale5 + FEV1_preBD_PCT_Baseline_rescale10 + Pre_FEV1_FVC_ratio_rescale10 + Smoking_0never_1ex_2current + CRSsNP + CRSwNP + offset(log(Follow_up_duration_days)), data = Data.sg.dream, subset = Eos_group == 2)
mod9 <- glm.nb(Number_severe_asthma_attacks_during_followup ~ Mepo_placebo + ACQ_baseline_score_mean + FeNO_baseline_ppb + Age_rescale10 + Gender_0Female_1Male + BMI_rescale5 + FEV1_preBD_PCT_Baseline_rescale10 + Pre_FEV1_FVC_ratio_rescale10 + Smoking_0never_1ex_2current + CRSsNP + CRSwNP + offset(log(Follow_up_duration_days)), data = Data.sg.dream, subset = Eos_group == 3)
mod10 <- glm.nb(Number_severe_asthma_attacks_during_followup ~ Mepo_placebo + ACQ_baseline_score_mean  + Age_rescale10 + Gender_0Female_1Male + BMI_rescale5 + FEV1_preBD_PCT_Baseline_rescale10 + Pre_FEV1_FVC_ratio_rescale10 + Smoking_0never_1ex_2current + CRSsNP + offset(log(Follow_up_duration_days)), data = Data.sg.dream, subset = T2_composite == 'T2-Low/Low')
mod11 <- glm.nb(Number_severe_asthma_attacks_during_followup ~ Mepo_placebo + ACQ_baseline_score_mean  + Age_rescale10 + Gender_0Female_1Male + BMI_rescale5 + FEV1_preBD_PCT_Baseline_rescale10 + Pre_FEV1_FVC_ratio_rescale10 + Smoking_0never_1ex_2current + CRSsNP + CRSwNP + offset(log(Follow_up_duration_days)), data = Data.sg.dream, subset = T2_composite == 'T2-Mid')
mod12 <- glm.nb(Number_severe_asthma_attacks_during_followup ~ Mepo_placebo + ACQ_baseline_score_mean  + Age_rescale10 + Gender_0Female_1Male + BMI_rescale5 + FEV1_preBD_PCT_Baseline_rescale10 + Pre_FEV1_FVC_ratio_rescale10 + Smoking_0never_1ex_2current + CRSsNP + CRSwNP + offset(log(Follow_up_duration_days)), data = Data.sg.dream, subset = T2_composite == 'T2-High/High')


# Coefficients and interval confidence for all regressions
coef <- round(rbind("Overall" = exp(Confint(mod0)[2,]), NA,
                    "ACQ_1" = exp(Confint(mod1)[2,]), "ACQ_2" = exp(Confint(mod2)[2,]), "ACQ_3" = exp(Confint(mod3)[2,]), NA, 
                    "Eos_1" = exp(Confint(mod7)[2,]), "Eos_2" = exp(Confint(mod8)[2,]), "Eos_3" = exp(Confint(mod9)[2,]), NA,
                    "FeNO_1" = exp(Confint(mod4)[2,]), "FeNO_2" = exp(Confint(mod5)[2,]), "FeNO_3" = exp(Confint(mod6)[2,]), NA,
                    "T2_1" = exp(Confint(mod10)[2,]), "T2_2" = exp(Confint(mod11)[2,]), "T2_3" = exp(Confint(mod12)[2,])), 2)


# Data with min/max for each group ACQ
Data_attack_acq <- Data.sg.dream %>% 
  filter(!is.na(ACQ_group)) %>% 
  group_by(Mepo_placebo, ACQ_group) %>% 
  reframe(
    n = n(),
    AbsRisk = sum(Number_severe_asthma_attacks_during_followup, na.rm = TRUE) /
              (sum(Follow_up_duration_days, na.rm = TRUE) / 365.25),
    # Median = median(Number_severe_asthma_attacks_during_followup, na.rm = TRUE), 
    # Q1 = quantile(Number_severe_asthma_attacks_during_followup, 0.25, na.rm = TRUE),
    # Q3 = quantile(Number_severe_asthma_attacks_during_followup, 0.75, na.rm = TRUE),
    Min = min(Number_severe_asthma_attacks_during_followup, na.rm = TRUE),
    Max = max(Number_severe_asthma_attacks_during_followup, na.rm = TRUE),
  ) %>% 
  mutate(
    Cat = "ACQ",
    .before = 1
  ) %>% 
  rename(group = ACQ_group)

# Data with min/max for each group FeNO
Data_attack_FeNO <- Data.sg.dream %>% 
  filter(!is.na(FeNO_group)) %>% 
  group_by(Mepo_placebo, FeNO_group) %>% 
  reframe(
    n = n(),
    AbsRisk = sum(Number_severe_asthma_attacks_during_followup, na.rm = TRUE) /
              (sum(Follow_up_duration_days, na.rm = TRUE) / 365.25),
    # Median = median(Number_severe_asthma_attacks_during_followup, na.rm = TRUE), 
    # Q1 = quantile(Number_severe_asthma_attacks_during_followup, 0.25, na.rm = TRUE),
    # Q3 = quantile(Number_severe_asthma_attacks_during_followup, 0.75, na.rm = TRUE),
    Min = min(Number_severe_asthma_attacks_during_followup, na.rm = TRUE),
    Max = max(Number_severe_asthma_attacks_during_followup, na.rm = TRUE),
  ) %>% 
  mutate(
    Cat = "FeNO",
    .before = 1
  ) %>% 
  rename(group = FeNO_group)

# Data with min/max for each group Eos
Data_attack_Eos <- Data.sg.dream %>% 
  filter(!is.na(Eos_group)) %>% 
  group_by(Mepo_placebo, Eos_group) %>% 
  reframe(
    n = n(),
    AbsRisk = sum(Number_severe_asthma_attacks_during_followup, na.rm = TRUE) /
              (sum(Follow_up_duration_days, na.rm = TRUE) / 365.25),
    # Median = median(Number_severe_asthma_attacks_during_followup, na.rm = TRUE), 
    # Q1 = quantile(Number_severe_asthma_attacks_during_followup, 0.25, na.rm = TRUE),
    # Q3 = quantile(Number_severe_asthma_attacks_during_followup, 0.75, na.rm = TRUE),
    Min = min(Number_severe_asthma_attacks_during_followup, na.rm = TRUE),
    Max = max(Number_severe_asthma_attacks_during_followup, na.rm = TRUE),
  ) %>% 
  mutate(
    Cat = "Eos",
    .before = 1
  ) %>% 
  rename(group = Eos_group)


# Data with min/max for each group T2
Data_attack_T2 <- Data.sg.dream %>% 
  filter(!is.na(T2_composite)) %>% 
  group_by(Mepo_placebo, T2_composite) %>% 
  reframe(
    n = n(),
    AbsRisk = sum(Number_severe_asthma_attacks_during_followup, na.rm = TRUE) /
              (sum(Follow_up_duration_days, na.rm = TRUE) / 365.25),
    # Median = median(Number_severe_asthma_attacks_during_followup, na.rm = TRUE), 
    # Q1 = quantile(Number_severe_asthma_attacks_during_followup, 0.25, na.rm = TRUE),
    # Q3 = quantile(Number_severe_asthma_attacks_during_followup, 0.75, na.rm = TRUE),
    Min = min(Number_severe_asthma_attacks_during_followup, na.rm = TRUE),
    Max = max(Number_severe_asthma_attacks_during_followup, na.rm = TRUE),
  ) %>% 
  mutate(
    Cat = "Eos",
    .before = 1
  ) %>% 
  rename(group = T2_composite)


Data_attack_placebo <- rbind(Data_attack_acq, Data_attack_Eos, Data_attack_FeNO,Data_attack_T2) %>% 
  filter(Mepo_placebo == "Placebo") %>% 
  add_row(Cat = "Overall", Mepo_placebo = "Placebo", group = "Overall", n = sum(Data.sg.dream$Mepo_placebo == "Placebo"), Min = min(Data.oracle.dream$Number_severe_asthma_attacks_during_followup[Data.sg.dream$Mepo_placebo == "Placebo"], na.rm = TRUE), Max = max(Data.oracle.dream$Number_severe_asthma_attacks_during_followup[Data.sg.dream$Mepo_placebo == "Placebo"], na.rm = TRUE),.before = 1) %>% 
  add_row(.before = 2) %>% 
  add_row(.after = 5) %>% 
  add_row(.after = 9) %>% 
  add_row(.after = 13) 

Data_attack_mepo <- rbind(Data_attack_acq, Data_attack_Eos, Data_attack_FeNO,Data_attack_T2) %>% 
  filter(Mepo_placebo == "Mepo") %>% 
  add_row(Cat = "Overall", Mepo_placebo = "Mepo", group = "Overall", n = sum(Data.sg.dream$Mepo_placebo == "Mepo"), Min = min(Data.oracle.dream$Number_severe_asthma_attacks_during_followup[Data.sg.dream$Mepo_placebo == "Mepo"], na.rm = TRUE), Max = max(Data.oracle.dream$Number_severe_asthma_attacks_during_followup[Data.sg.dream$Mepo_placebo == "Mepo"], na.rm = TRUE),.before = 1) %>% 
  add_row(.before = 2) %>% 
  add_row(.after = 5) %>% 
  add_row(.after = 9) %>% 
  add_row(.after = 13) 

mydf <- data.frame(
  SubgroupH = c("Overall", "ACQ", NA, NA, NA, "Blood eos", NA, NA, NA, "FeNO", NA, NA, NA, "T2 ordinal", NA, NA, NA),
  Subgroup = c(NA, NA, "< Q1 (1.6)", "[Q1 - Q3]", "> Q3 (3.2)", NA, "< Q1 (0.15)", "[Q1 - Q3]", "> Q3 (0.51)", NA, "< Q1 (18)", "[Q1 - Q3]", "> Q3 (56)", NA, "LowLow", "Mid", "HotHot"),
  NoOfPatients_placebo = Data_attack_placebo$n,
  NoOfPatients_mepo = Data_attack_mepo$n,
  Placebo = ifelse(is.na(Data_attack_placebo$AbsRisk), NA, round(Data_attack_placebo$AbsRisk,2) ),
  Mepo = ifelse(is.na(Data_attack_mepo$AbsRisk), NA, round(Data_attack_mepo$AbsRisk,2) ),
  RateRatio = coef[,1],
  RateLower = coef[,2],
  RateUpper = coef[,3],
  stringsAsFactors = FALSE
)

# Data frame in log for the forest plot
df <- mydf %>%
  mutate(
    row = seq(n(), 1),
    logRR = log(RateRatio),
    logLower = log(RateLower),
    logUpper = log(RateUpper),
    SubgroupH = ifelse(is.na(SubgroupH), "", SubgroupH),
    Subgroup = ifelse(is.na(Subgroup), "", Subgroup),
    pat100 = ifelse(is.na(NoOfPatients_placebo), "", format(NoOfPatients_placebo, big.mark=",")),
    pat200 = ifelse(is.na(NoOfPatients_mepo), "", format(NoOfPatients_mepo, big.mark=",")),
    textRR = ifelse(!is.na(RateRatio), paste0(RateRatio, " (", RateLower, "-", RateUpper, ")"), "")
  )

df$Placebo <- ifelse(df$SubgroupH == 'Overall', 2.11, df$Placebo)
df$Mepo <- ifelse(df$SubgroupH == 'Overall', 1.19, df$Mepo)

df <- df %>%
  mutate(
    lab_placebo = ifelse(
      is.na(pat100) | is.na(Placebo),
      "",
      paste0(pat100, " (", sprintf("%.2f", Placebo), ")")
    ),
    lab_mepo = ifelse(
      is.na(pat200) | is.na(Mepo),
      "",
      paste0(pat200, " (", sprintf("%.2f", Mepo), ")")
    )
  )


```

```{r, echo = FALSE, fig.height = 6, fig.width = 9, warning = FALSE}


ticks  <- log(c(0.25, 0.5, 1, 2, 4))
labels <- c("0.25", "0.5", "1", "2", "4")

p_dream_adjusted <- ggplot(df, aes(x = logRR, y = row)) +
  
  # Confidence interval
  geom_segment(aes(x = logLower, xend = logUpper,
                   y = row, yend = row)) +
  
  # RR points
  geom_point(size = 3, shape = 15) +
  
  # Vertical line for RR = 1
  geom_segment(aes(x = 0, xend = 0, y = 18.5, yend = 0),
               linetype = 3, inherit.aes = FALSE) +
  
  # Vertical line for 0.25, 0.5, 1, 2, 4
  # geom_segment(
  #   data = data.frame(x = ticks),
  #   aes(x = x, xend = x, y = 13.5, yend = 0),
  #   linetype = 3,
  #   inherit.aes = FALSE
  # ) +
  
  # X axis 
  scale_x_continuous(
    breaks = ticks,
    labels = labels,
    limits = c(-8, 4)
  ) +
  
  # Graduations of the X axis
  geom_segment(
    data = data.frame(x = ticks),
    aes(x = x, xend = x, y = -0.1, yend = -0.4),
    inherit.aes = FALSE
  ) +
  geom_segment(aes(x = min(ticks), xend = max(ticks), y = -0.1, yend = -0.1),
               inherit.aes = FALSE) +
  
  # Labels
  geom_text(
    data = data.frame(x = ticks, lab = labels),
    aes(x = x, y = -0.8, label = lab),
    size = 3,
    inherit.aes = FALSE
  ) +
  
  # Y axis
  scale_y_continuous(limits = c(-3, 22), expand = c(0,0)) +
  
  # Themes
  theme_minimal(base_size = 10) +
  theme(
    axis.title = element_blank(),
    axis.text.y = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.line.x = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.minor.y = element_blank()
  )

# Add text to the plot
p_dream_adjusted <- p_dream_adjusted +
  
  # Subgroups (left column)
  geom_text(aes(x = -8, label = SubgroupH), hjust = 0, fontface = "italic") +
  geom_text(aes(x = -7.7, label = Subgroup), hjust = 0) +
  
  geom_text(aes(x = -4.8, label = lab_placebo), hjust = 1) +
  geom_text(aes(x = -2.8, label = lab_mepo), hjust = 1) +
  # Rate ratio
  geom_text(aes(x = 1.8, label = textRR), hjust = 0) +
  
  # Column titles
  annotate("text", x = -8, y = 20, label = "Subgroup", hjust = 0, fontface = 2) +
  annotate("text", x = -2.6, y = 21, label = "No. of Patients (Estimated rate)", hjust = 1, fontface = 2) +
  annotate("text", x = -4.7, y = 20, label = "Placebo", hjust = 1, fontface = 2) +
  annotate("text", x = -2.7, y = 20, label = "Mepo", hjust = 1, fontface = 2) +
  annotate("text", x = 1.5,  y = 20, label = "        Adjusted\nRate Ratio (95% CI)", hjust = 0, fontface = 2) +
  
  annotate("text", x = -1.5, y = -2, label = "Favors Mepo", fontface = 2) +
  annotate("text", x =  1.5, y = -2, label = "Favors Placebo", fontface = 2) 

p_dream_adjusted

```

## PRISMA {.tabset .tabset-fade}

### Unadjusted

```{r, echo = FALSE}

Data.sg.Prisma <- Data.Prisma %>% 
  dplyr::select(ID, ACQ5_rescale,acq5_moy, FeNO_log10, BloodEOS_Abso_log10, V2_V1_Post_FEV1_L, GINA_Step, Age_at_study_enrollment, Sex, BMI, Pre_FEV1_PORC, Pre_FEV1_FVC_ratio, Severe_asthma_attacks_past_12m_cat, V1_Pre_FEV1_L_12m)

Q25 <- quantile(Data.sg.Prisma$acq5_moy, 0.25, na.rm = TRUE)
Q75 <- quantile(Data.sg.Prisma$acq5_moy, 0.75, na.rm = TRUE)
IQR_acq5 <- Q75 - Q25
Data.sg.Prisma$acq5_IQR <- Data.sg.Prisma$acq5_moy / IQR_acq5

# Linear regressions for each group
mod0 <- lm(V2_V1_Post_FEV1_L ~ acq5_IQR, data = Data.sg.Prisma)
mod1 <- lm(V2_V1_Post_FEV1_L ~ FeNO_log10, data = Data.sg.Prisma)
mod2 <- lm(V2_V1_Post_FEV1_L ~ BloodEOS_Abso_log10, data = Data.sg.Prisma)

# Coefficients and interval confidence for all regressions
coef <- round(rbind("Overall" = Confint(mod0)[-1,], "ACQ_2" = Confint(mod2)[-1,],
                    "ACQ_1" = Confint(mod1)[-1,]), 2)

mydf <- data.frame(
  SubgroupH = c("ACQ5 (per QIR)", 'Blood eos (per 10-fold)', "FeNO (per 10-fold)"),
  RateRatio = coef[,1],
  RateLower = coef[,2],
  RateUpper = coef[,3],
  stringsAsFactors = FALSE
)

# Data frame in log for the forest plot
df <- mydf %>%
  mutate(
    row = seq(n(), 1),
    logRR = RateRatio,
    logLower = RateLower,
    logUpper = RateUpper,
    SubgroupH = ifelse(is.na(SubgroupH), "", SubgroupH),
    textRR = ifelse(!is.na(RateRatio), paste0(RateRatio, " (", RateLower, "-", RateUpper, ")"), "")
  )

```

```{r, echo = FALSE, fig.height = 4, fig.width = 8, warning = FALSE}

ticks  <- seq(-0.5, 1, by = 0.25)
labels <- ticks

p_prisma <- ggplot(df, aes(x = logRR, y = row)) +
  
  # Confidence interval
  geom_segment(aes(x = logLower, xend = logUpper,
                   y = row, yend = row)) +
  
  # MD points
  geom_point(size = 3, shape = 15) +
  
  # Vertical line for MD = 0
  geom_segment(aes(x = 0, xend = 0, y = 4, yend = 0),
               linetype = 3, inherit.aes = FALSE) +
  
  # Graduations of the X axis
  geom_segment(
    data = data.frame(x = ticks),
    aes(x = x, xend = x, y = 0, yend = -0.4),
    inherit.aes = FALSE
  ) +
  geom_segment(aes(x = min(ticks), xend = max(ticks), y = 0, yend = 0),
               inherit.aes = FALSE) +
  
  # Labels
  geom_text(
    data = data.frame(x = ticks, lab = labels),
    aes(x = x, y = -0.6, label = lab),
    size = 3,
    inherit.aes = FALSE
  ) +
  
  # Y axis
  scale_y_continuous(limits = c(-2, 5), expand = c(0,0))+
  scale_x_continuous(limits = c(-2.5, 3)) +
  
  # Themes
  theme_minimal(base_size = 10) +
  theme(
    axis.title = element_blank(),
    axis.text.y = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.line.x = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.minor.y = element_blank()
  )

# Add text to the plot
p_prisma <- p_prisma +
  
    # title
  geom_text(aes(x = -2.5, label = SubgroupH), hjust = 0, fontface = "italic") +
  
  # Mean Diff
  geom_text(aes(x = 1.5, label = textRR), hjust = 0) +
  
  # Column titles
  annotate("text", x = -2.5, y = 4.5, label = "Variable", hjust = 0, fontface = 2) +
  annotate("text", x = 1.5,  y = 4.5, label = "Mean difference (95% CI)", hjust = 0, fontface = 2) +
  
  annotate("text", x = -0.5, y = -1.5, label = "Favors Less Improvement", fontface = 2) +
  annotate("text", x =  1.2, y = -1.5, label = "Favors Improvement", fontface = 2) 

p_prisma

```

### Adjusted - v1

One model adjusted for BEC, FeNO and FEV1 V1 (Post_FEV1_L).

```{r, echo = FALSE}

Data.sg.Prisma <- Data.Prisma %>% 
  dplyr::select(ID, ACQ5_rescale,acq5_moy, FeNO_log10,FeNO, BloodEOS_Abso_log10,BloodEOS_Abso, 
                V2_V1_Post_FEV1_L, GINA_Step, Age_at_study_enrollment, 
                Sex, BMI, Pre_FEV1_PORC, Pre_FEV1_FVC_ratio, 
                Severe_asthma_attacks_past_12m_cat, Post_FEV1_L,T2_hothot_med_lowlow_ordinal)

Q25 <- quantile(Data.sg.Prisma$acq5_moy, 0.25, na.rm = TRUE)
Q75 <- quantile(Data.sg.Prisma$acq5_moy, 0.75, na.rm = TRUE)
IQR_acq5 <- Q75 - Q25
Data.sg.Prisma$acq5_IQR <- Data.sg.Prisma$acq5_moy / IQR_acq5

Q25 <- quantile(Data.sg.Prisma$FeNO, 0.25, na.rm = TRUE)
Q75 <- quantile(Data.sg.Prisma$FeNO, 0.75, na.rm = TRUE)
IQR_FeNO <- Q75 - Q25
Data.sg.Prisma$FeNO_IQR <- Data.sg.Prisma$FeNO / IQR_FeNO

Q25 <- quantile(Data.sg.Prisma$BloodEOS_Abso, 0.25, na.rm = TRUE)
Q75 <- quantile(Data.sg.Prisma$BloodEOS_Abso, 0.75, na.rm = TRUE)
IQR_BloodEOS_Abso <- Q75 - Q25
Data.sg.Prisma$BloodEOS_Abso_IQR <- Data.sg.Prisma$BloodEOS_Abso / IQR_BloodEOS_Abso

# Fit multivariable linear regression
mod_multi <- lm(V2_V1_Post_FEV1_L ~ acq5_IQR + BloodEOS_Abso_IQR + FeNO_IQR + Post_FEV1_L,
                data = Data.sg.Prisma)

mod_multi2 <- lm(V2_V1_Post_FEV1_L ~ acq5_IQR + T2_hothot_med_lowlow_ordinal + Post_FEV1_L,
                data = Data.sg.Prisma)

# Extract coefficients and 95% CIs for the predictors only
coef <- round(
  Confint(mod_multi)[c(2:4), ], 2
)

coef2 <- round(
  Confint(mod_multi2)[c(3), ], 2
)

coef <- rbind(coef,coef2)

df <- data.frame(
  SubgroupH = c("ACQ5 (per IQR 1.6)", 
                "Blood eos (per IQR 0.26)", 
                "FeNO (per IQR 26)",
                "T2 ordinal"),
  logRR    = coef[,1],
  logLower = coef[,2],
  logUpper = coef[,3],
  row      = 4:1
)

df$textRR <- paste0(df$logRR, " (", df$logLower, ", ", df$logUpper, ")")

```

```{r, echo = FALSE, fig.height = 4, fig.width = 8, warning = FALSE}

ticks  <- seq(-0.5, 1, by = 0.25)
labels <- ticks

p_prisma_multi <- ggplot(df, aes(x = logRR, y = row)) +
  
  # Confidence interval
  geom_segment(aes(x = logLower, xend = logUpper,
                   y = row, yend = row)) +
  
  # MD points
  geom_point(size = 3, shape = 15) +
  
  # Vertical line for MD = 0
  geom_segment(aes(x = 0, xend = 0, y = 4.7, yend = 0),
               linetype = 3, inherit.aes = FALSE) +
  
  # Graduations of the X axis
  geom_segment(
    data = data.frame(x = ticks),
    aes(x = x, xend = x, y = -0.1, yend = -0.4),
    inherit.aes = FALSE
  ) +
  geom_segment(aes(x = min(ticks), xend = max(ticks), y = -0.1, yend = -0.1),
               inherit.aes = FALSE) +
  
  # Labels
  geom_text(
    data = data.frame(x = ticks, lab = labels),
    aes(x = x, y = -0.6, label = lab),
    size = 3,
    inherit.aes = FALSE
  ) +
  
  # Y axis
  scale_y_continuous(limits = c(-2, 5.5), expand = c(0,0))+
  scale_x_continuous(limits = c(-2.5, 3)) +
  
  # Themes
  theme_minimal(base_size = 10) +
  theme(
    axis.title = element_blank(),
    axis.text.y = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.line.x = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.minor.y = element_blank()
  )

# Add text to the plot
p_prisma_multi <- p_prisma_multi +
  
    # title
  geom_text(aes(x = -2.5, label = SubgroupH), hjust = 0, fontface = "italic") +
  
  # Mean Diff
  geom_text(aes(x = 1.5, label = textRR), hjust = 0) +
  
  # Column titles
  annotate("text", x = -2.5, y = 5, label = "Variable", hjust = 0, fontface = 2) +
  annotate("text", x = 1.5,  y = 5, label = "Mean difference (95% CI)", hjust = 0, fontface = 2) +
  
  annotate("text", x = -1, y = -1.5, label = "Favors Less Improvement", fontface = 2) +
  annotate("text", x =  1.5, y = -1.5, label = "Favors Improvement", fontface = 2) 

p_prisma_multi

```

## BOOST {.tabset .tabset-fade}

### Unadjusted

```{r, echo = FALSE}

Data.sg.boost <- Data.boost %>% 
  dplyr::select(study_id, ACQ5_rescale,acq_attack, FeNO_log10,feno_attack, BEC_log10,blood_eos_attack, delta,Age_rescale10,Sex , BMI_rescale5,Pre_FEV1_PORC_rescale, Pre_FEV1_FVC_ratio_rescale, Previous_asthma_attack_yesno, fev1_attack  )

Q25 <- quantile(Data.sg.boost$acq_attack, 0.25, na.rm = TRUE)
Q75 <- quantile(Data.sg.boost$acq_attack, 0.75, na.rm = TRUE)
IQR_acq5 <- Q75 - Q25
Data.sg.boost$acq5_IQR <- Data.sg.boost$acq_attack / IQR_acq5

Q25 <- quantile(Data.sg.boost$feno_attack, 0.25, na.rm = TRUE)
Q75 <- quantile(Data.sg.boost$feno_attack, 0.75, na.rm = TRUE)
IQR_FeNO <- Q75 - Q25
Data.sg.boost$FeNO_IQR <- Data.sg.boost$feno_attack / IQR_FeNO

Q25 <- quantile(Data.sg.boost$blood_eos_attack, 0.25, na.rm = TRUE)
Q75 <- quantile(Data.sg.boost$blood_eos_attack, 0.75, na.rm = TRUE)
IQR_BloodEOS_Abso <- Q75 - Q25
Data.sg.boost$BloodEOS_Abso_IQR <- Data.sg.boost$blood_eos_attack / IQR_BloodEOS_Abso

# Linear regressions for each group
mod0 <- lm(delta ~ acq5_IQR, data = Data.sg.boost)
mod1 <- lm(delta ~ FeNO_log10, data = Data.sg.boost)
mod2 <- lm(delta ~ BEC_log10, data = Data.sg.boost)

# Coefficients and interval confidence for all regressions
coef <- round(rbind("Overall" = Confint(mod0)[-1,], "ACQ_2" = Confint(mod2)[-1,],
                    "ACQ_1" = Confint(mod1)[-1,]), 2)

mydf <- data.frame(
  SubgroupH = c("ACQ5 (per IQR)", 'Blood eos (per 10-fold)', "FeNO (per 10-fold)"),
  RateRatio = coef[,1],
  RateLower = coef[,2],
  RateUpper = coef[,3],
  stringsAsFactors = FALSE
)

# Data frame in log for the forest plot
df <- mydf %>%
  mutate(
    row = seq(n(), 1),
    logRR = RateRatio,
    logLower = RateLower,
    logUpper = RateUpper,
    SubgroupH = ifelse(is.na(SubgroupH), "", SubgroupH),
    textRR = ifelse(!is.na(RateRatio), paste0(RateRatio, " (", RateLower, "-", RateUpper, ")"), "")
  )

```

```{r, echo = FALSE, fig.height = 4, fig.width = 8, warning = FALSE}


ticks  <- seq(-0.5, 1, by = 0.25)
labels <- ticks

p_boost <- ggplot(df, aes(x = logRR, y = row)) +
  
  # Confidence interval
  geom_segment(aes(x = logLower, xend = logUpper,
                   y = row, yend = row)) +
  
  # MD points
  geom_point(size = 3, shape = 15) +
  
  # Vertical line for MD = 0
  geom_segment(aes(x = 0, xend = 0, y = 4, yend = 0),
               linetype = 3, inherit.aes = FALSE) +
  
  # Graduations of the X axis
  geom_segment(
    data = data.frame(x = ticks),
    aes(x = x, xend = x, y = 0, yend = -0.4),
    inherit.aes = FALSE
  ) +
  geom_segment(aes(x = min(ticks), xend = max(ticks), y = 0, yend = 0),
               inherit.aes = FALSE) +
  
  # Labels
  geom_text(
    data = data.frame(x = ticks, lab = labels),
    aes(x = x, y = -0.6, label = lab),
    size = 3,
    inherit.aes = FALSE
  ) +
  
  # Y axis
  scale_y_continuous(limits = c(-2, 5), expand = c(0,0))+
  scale_x_continuous(limits = c(-2.5, 3)) +
  
  # Themes
  theme_minimal(base_size = 10) +
  theme(
    axis.title = element_blank(),
    axis.text.y = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.line.x = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.minor.y = element_blank()
  )

# Add text to the plot
p_boost <- p_boost +
  
    # title
  geom_text(aes(x = -2.5, label = SubgroupH), hjust = 0, fontface = "italic") +
  
  # Mean Diff
  geom_text(aes(x = 1.5, label = textRR), hjust = 0) +
  
  # Column titles
  annotate("text", x = -2.5, y = 4.5, label = "Variable", hjust = 0, fontface = 2) +
  annotate("text", x = 1.5,  y = 4.5, label = "Mean difference (95% CI)", hjust = 0, fontface = 2) +
  
  annotate("text", x = -0.5, y = -1.5, label = "Favors Less Improvement", fontface = 2) +
  annotate("text", x =  1.2, y = -1.5, label = "Favors Improvement", fontface = 2) 

p_boost


```

### Adjusted - v1

```{r, echo = FALSE}

# Fit multivariable linear regression
mod_multi <- lm(delta ~ acq5_IQR + BloodEOS_Abso_IQR + FeNO_IQR  + fev1_attack, data = Data.sg.boost)

# Extract coefficients and 95% CIs for the predictors only
coef <- round(
  Confint(mod_multi)[c(2:4), ], 2
)

df <- data.frame(
  SubgroupH = c("ACQ5 (per IQR 1.4)", 
                "Blood eos (per IQR 0.06)", 
                "FeNO (per IQR 62.5)"),
  logRR    = coef[,1],
  logLower = coef[,2],
  logUpper = coef[,3],
  row      = 3:1
)

df$textRR <- paste0(df$logRR, " (", df$logLower, ", ", df$logUpper, ")")

df$logLower <- ifelse( df$logLower < -0.5, -0.5, df$logLower              )
df$logUpper <- ifelse( df$logUpper > 1, 1, df$logUpper              )

```

```{r, echo = FALSE, fig.height = 4, fig.width = 8, warning = FALSE}

ticks  <- seq(-0.5, 1, by = 0.25)
labels <- ticks

p_boost_multi <- ggplot(df, aes(x = logRR, y = row)) +
  
  # Confidence interval
  geom_segment(aes(x = logLower, xend = logUpper,
                   y = row, yend = row)) +
  
  # MD points
  geom_point(size = 3, shape = 15) +
  
  # Vertical line for MD = 0
  geom_segment(aes(x = 0, xend = 0, y = 4, yend = 0),
               linetype = 3, inherit.aes = FALSE) +
  
  # Graduations of the X axis
  geom_segment(
    data = data.frame(x = ticks),
    aes(x = x, xend = x, y = -0.1, yend = -0.4),
    inherit.aes = FALSE
  ) +
  geom_segment(aes(x = min(ticks), xend = max(ticks), y = -0.1, yend = -0.1),
               inherit.aes = FALSE) +
  
  # Labels
  geom_text(
    data = data.frame(x = ticks, lab = labels),
    aes(x = x, y = -0.6, label = lab),
    size = 3,
    inherit.aes = FALSE
  ) +
  
  # Y axis
  scale_y_continuous(limits = c(-2, 5), expand = c(0,0))+
  scale_x_continuous(limits = c(-2.5, 3)) +
  
  # Themes
  theme_minimal(base_size = 10) +
  theme(
    axis.title = element_blank(),
    axis.text.y = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.line.x = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.minor.y = element_blank()
  )

# Add text to the plot
p_boost_multi <- p_boost_multi +
  
    # title
  geom_text(aes(x = -2.5, label = SubgroupH), hjust = 0, fontface = "italic") +
  
  # Mean Diff
  geom_text(aes(x = 1.5, label = textRR), hjust = 0) +
  
  # Column titles
  annotate("text", x = -2.5, y = 4.5, label = "Variable", hjust = 0, fontface = 2) +
  annotate("text", x = 1.5,  y = 4.5, label = "Mean difference (95% CI)", hjust = 0, fontface = 2) +
  
  annotate("text", x = -1, y = -1.5, label = "Favors Less Improvement", fontface = 2) +
  annotate("text", x =  1.5, y = -1.5, label = "Favors Improvement", fontface = 2) 

p_boost_multi

```

# Density plots {.tabset .tabset-fade}

## Oracle Dream

```{r, echo = FALSE, fig.height = 4, fig.width = 8}

p_density_dream <- Data.oracle.dream %>% 
  filter(!is.na(ACQ_baseline_score_mean)) %>% 
  ggplot(aes(x = ACQ_baseline_score_mean, colour = Mepo_placebo, fill = Mepo_placebo)) +
  geom_density(alpha = 0.5) +
  scale_color_manual(values = c("#00a1d5", "#b24745")) +
  scale_fill_manual(values = c("#00a1d5", "#b24745")) +
  scale_x_continuous(limits = c(0, 10), breaks = seq(0, 5, by = 1), expand = c(0, 0)) +
  coord_cartesian(xlim = c(1, 5), expand = FALSE) +
  xlab("\nBaseline ACQ-5 Score") +
  ylab("\nProbability density \n") +
  theme_classic() +
  theme(
    axis.text.x = element_text(size = 12),
    axis.title = element_text(face = "bold", size = 12),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank(),
    legend.title = element_blank()
  )

p_density_dream + ylab("Probability density \n")

```

## Prisma

```{r, echo = FALSE, fig.height = 4, fig.width = 8}

p_density_prisma <- Data.Prisma %>% 
  filter(!is.na(acq5_moy)) %>% 
  ggplot(aes(x = acq5_moy)) +
  geom_density(fill = "lightblue", colour = "lightblue", alpha = 0.5) +
  scale_x_continuous(limits = c(0, 10), breaks = seq(0, 5, by = 1), expand = c(0, 0)) +
  coord_cartesian(xlim = c(1, 5), expand = FALSE) +
  labs(
    x = "\nBaseline ACQ-5 Score", 
    y = "\n Probability density \n"
  ) +
  theme_classic() +
  theme(
    axis.text.x = element_text(size = 12),
    axis.title = element_text(face = "bold", size = 12),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank(),
    legend.title = element_blank()
  )
p_density_prisma

```

## Boost

```{r, echo = FALSE, fig.height = 4, fig.width = 8}

p_density_boost <- Data.boost %>% 
  filter(!is.na(acq_attack)) %>% 
  ggplot(aes(x = acq_attack)) +
  geom_density(fill = "lightblue", colour = "lightblue", alpha = 0.5) +
  scale_x_continuous(limits = c(0, 10), breaks = seq(0, 5, by = 1), expand = c(0, 0)) +
  coord_cartesian(xlim = c(1, 5), expand = FALSE) +
  xlab("\nBaseline ACQ-5 Score") +
  ylab(NULL) +
  theme_classic() +
  theme(
    axis.text.x = element_text(size = 12),
    axis.title = element_text(face = "bold", size = 12),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank(),
    legend.title = element_blank()
  )

p_density_boost + ylab("Probability density \n")

```

# Stacked figures {.tabset .tabset-fade}

## Dream

```{r, echo = FALSE, fig.height = 14, fig.width = 8, warning = FALSE}

top_dream <- plot_grid(
  NULL,
  spline_dream,
  NULL,
  p_density_dream,
  ncol = 1,
  rel_heights = c(0.2, 1, 0.2, 0.5),
  labels = c("", "a", "", "b"),
  align = "v"
)


bot_dream <- plot_grid(
  p_dream_adjusted,
  NULL,
  ncol = 1,
  rel_heights = c(1, 0.1),
  labels = c("c", "")
)

plot_grid(
  top_dream,
  NULL,
  bot_dream,
  ncol = 1,
  rel_heights = c(1, 0.15, 1.2)
)


u_dream <- plot_grid(
  top_dream,
  NULL,
  bot_dream,
  ncol = 1,
  rel_heights = c(1, 0.15, 1.2)
)

ggsave(
  filename = "plot_DREAM.pdf",
  plot = u_dream,   
  device = "pdf",
  width = 8,
  height = 14,
  units = "in"
)

```

## Prisma and boost (Adjusted for BEC, FeNO and FEV1 post L V1 in one model) - v1

```{r, echo = FALSE, fig.height = 12, fig.width = 12, warning = FALSE}

top4 <- plot_grid(
  spline_prisma, 
  spline_boost,
  NULL,
  NULL,
  p_density_prisma,
  p_density_boost,
  ncol = 2,
  labels = c("a", "b", "", "", "c", "d"),
  rel_heights = c(1, 0.2, 0.5),
  align = "v"
)

bot2 <- plot_grid(
  p_prisma_multi,
  p_boost_multi,
  NULL,
  NULL,
  ncol = 2,
  labels = c("e", "f", "", ""),
  rel_heights = c(1, 0.1),
  align = "h"
)

plot_grid(
  top4,
  NULL,
  bot2,
  ncol = 1,
  rel_heights = c(1, 0.1, 0.4)
)

u <- plot_grid(
  top4,
  NULL,
  bot2,
  ncol = 1,
  rel_heights = c(1, 0.1, 0.4)
)

ggsave(
  filename = "plot_PRISMA_BOOSTv2.pdf",
  plot = u,   
  device = "pdf",
  width = 14,
  height = 14,
  units = "in"
)

```
